### 2015-12-31 paper1.tex
initial commit

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.

\title{\LARGE \bf
  Longitudinal Trajectory Planning and Tracking for Fixed-Path Autonomous Driving
}

\author{
  Robert G. Cofield $^{1}$,
  Rakesh Gupta $^{2}$,
  Ambarish Goswami $^{3}$,
  and
  David Bevly $^{4}$
  \thanks{
    * This work is supported by Honda Research Institute USA.
  }
  \thanks{
    $^{1}$ Robert G. Cofield is a graduate researcher at Auburn University's GPS \& Vehicle Dynamics Laboratory.
  }
  \thanks{
    $^{2}$ Rakesh Gupta ...
  }
  \thanks{
    $^{3}$ Ambarish Goswami ... 
  }
  \thanks{
    $^{4}$ David Bevly ...
  }
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

When operating an autonomous ground vehicle on roadways, it is common to employ the path-velocity decomposition (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
Given a desired path, we present a novel method of planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
This method first segments the path using an arbitrary set of kinematic or dynamic constaints.
Each segement is then planned sequentially by choosing a piecewise constant jerk profile.
The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous throughout the entire path.
The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control.
This method is shown to be efficient and reliable for use in online planning with a test vehicle.

\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

asdf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2015-12-31 paper1.tex
started outlining

@@
 
-When operating an autonomous ground vehicle on roadways, it is common to employ the path-velocity decomposition (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
+When planning trajectories for an autonomous ground vehicle operating on roadways, it is common to employ the path-velocity decomposition (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
 Given a desired path, we present a novel method of planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
-This method first segments the path using an arbitrary set of kinematic or dynamic constaints.
+This method first segments the path using an arbitrary set of kinematic or dynamic constaints which may be functions of the desired path.
 Each segement is then planned sequentially by choosing a piecewise constant jerk profile.
@@
 
+... autonomous vehicles ... \emph{The 1st sentence is the hardest}.
+One such task is computing a tentative reference trajectory plan for travelling from one point to another.
+While the set of possible paths between any two points is infinite, it is typically heavily restricted by consideration of constraints such as traversability and legally available travel lanes, among others.
+The intended speed, acceleration, and higher derivates of position with respect to time are also subject to constraints.
+These may include legal speed limits, engine dynamics, desired sideslip limits, tractions availability, and rollover concerns.
+
+% intersection of path-planning constraints and speed-planning constraints is vehicle dynamics
+
+% review of literature for a few paragraphs]
+%  - searching planners (OMPL stuff)
+%  - robot joint planning --> constant jerk intervals
+
+% We will use the path-velocity decomposition
+% mention using curvilinear CS
+%
+
+% tracking
+Once a trajectory plan is formulated, it must be utilized online in order for a reference signal to be sent to the controllers which govern steering, braking, and engine speed.
+This is the trajectory tracking problem.
+
+% mention using same cuvilinear cs scheme to project onto path
+
+\section{Trajectory Planning}
+
 asdf
+
+\section{Trajectory Tracking}
+
+asdf
+
+\section{Implementation}
+
+% How much do we really want to put here?
+
+% Show tracking performance? It's really more a function of the
 

### 2016-01-11 paper1.tex
added all the images and diagrams from HRI presentation

@@
 The intended speed, acceleration, and higher derivates of position with respect to time are also subject to constraints.
-These may include legal speed limits, engine dynamics, desired sideslip limits, tractions availability, and rollover concerns.
+These may include legal speed limits, engine dynamics, desired sideslip limits, traction availability, and rollover concerns.
 
@@
 
-asdf
+For the purposes of this work, the objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
+Additionally, prescribed initial and end conditions for position, speed, and acceleration must also be honored.
 

### 2016-01-11 dlc_schematic.py

@@ -1,0 +1,95 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+import g1fitting
+import its_utils
+import sys
+from numpy import ceil, rad2deg, pi
+import numpy as np
+import matplotlib.pyplot as plt
+
+## settings
+
+obs_dist = 10.
+lat_offset = 3.
+res = 0.1
+
+color_pri = 'g'
+color_alt = 'c'
+color_sn = 'b'
+
+## fixed points
+
+x1 = 0.
+y1 = 0.
+x2 = obs_dist
+y2 = lat_offset
+x3 = 2.*obs_dist
+y3 = 0.
+Ls = np.mean((x2-x1,x3-x2))
+
+nleg = ceil(Ls/res)
+x00 = list(np.linspace(x1-Ls, x1, nleg))
+y00 = [y1]*nleg
+x44 = list(np.linspace(x3, x3+Ls, nleg))
+y44 = [y3]*nleg
+
+## sinusoidal thing I created
+
+w1 = y2-y1
+w2 = y3-y2
+L1 = x2-x1
+L2 = x3-x2
+pth_sn1 = its_utils.lateral_transition_cosine(L1, w1, res)
+pth_sn2 = its_utils.lateral_transition_cosine(L2, w2, res, x2, y2)
+pth_sn = zip(x00, y00)[:-1] + pth_sn1[:-1] + pth_sn2[:-1] + zip(x44,y44)
+
+## Examine curvature
+
+[R_sn, k_sn, t_sn, dt_sn] = its_utils.circular_curvature(pth_sn)
+s_sn = its_utils.convert_path_to_1D(pth_sn)
+
+## Plot
+
+# path
+fig1 = plt.figure(facecolor='black')
+ax1 = fig1.add_subplot(1,1,1, axisbg='k')
+ax1.grid(True, color='w')
+ax1.hold(True)
+ax1.plot([p[0] for p in pth_sn], [0.]*len(pth_sn), color_pri, linewidth=8, label='Orig. Path')
+ax1.plot([p[0] for p in pth_sn], [lat_offset]*len(pth_sn), color_alt, linewidth=8, label='Alt. Path')
+ax1.plot([p[0] for p in pth_sn], [p[1] for p in pth_sn], color_sn, linewidth=8, label='DLC Path')
+# # w arrow
+# ax1.annotate(
+#   '', xy=(0, lat_offset), xycoords='data',
+#   xytext=(0, 0), textcoords='data',
[…]

### 2016-01-11 general_segment_solving.m

@@ -1,0 +1,151 @@
+%%
+% clear all; close all; clc
+% 
+% j = [1 0 -1 0 -1 0 1];
+% t = [0 1 2 3 4 5 6 7];
+% jpp = mkpp(t,j);
+% app = fnint(jpp, 0);
+% vpp = fnint(app, 0);
+% spp = fnint(vpp, 0);
+% 
+% tt = linspace(t(1), t(end), 500);
+% ss = fnval(spp, tt);
+% vv = fnval(vpp, tt);
+% aa = fnval(app, tt);
+% jj = fnval(jpp, tt);
+% 
+% f = figure
+% subplot(4,1,1)
+% plot(tt,ss, 'linewidth', 6)
+% grid on
+% whitebg('k')
+% ylabel('Pos', 'FontSize', 16, 'FontWeight', 'bold')
+% title('7 Phase Profile')
+% subplot(4,1,2)
+% plot(tt,vv, 'linewidth', 6)
+% grid on
+% whitebg('k')
+% ylabel('Speed', 'FontSize', 16, 'FontWeight', 'bold')
+% subplot(4,1,3)
+% plot(tt,aa, 'linewidth', 6)
+% grid on
+% ylabel('Accel', 'FontSize', 16, 'FontWeight', 'bold')
+% whitebg('k')
+% subplot(4,1,4)
+% plot(tt,jj, 'linewidth', 6)
+% grid on
+% whitebg('k')
+% ylabel('Jerk', 'FontSize', 16, 'FontWeight', 'bold')
+% xlabel('Time', 'FontSize', 16, 'FontWeight', 'bold')
+
+%%
+clear all; close all; clc
+
+j = [1 0 -1 0 -1 0 1];
+t = [0 .5 1.5 2.5 4 5 6.9+[0 1]];
+jpp = mkpp(t,j);
+app = fnint(jpp, .5);
+vpp = fnint(app, 1);
+spp = fnint(vpp, 0);
+
+tt = linspace(t(1), t(end), 500);
+ss = fnval(spp, tt);
+vv = fnval(vpp, tt);
+aa = fnval(app, tt);
+jj = fnval(jpp, tt);
+
+t_ = [t(1), t(end)];
+
+f = figure;
+p1 = subplot(2,2,1);
+hold on
+h = plot(tt,ss, 'linewidth', 6);
+plot(t_, [15 15], '--c', 'linewidth', 3)
[…]

### 2016-01-11 reverse_lookup_plot.m

@@ -1,0 +1,30 @@
+%%
+clear all; close all; clc
+
+j = [1 0 -1 0 -1 0 1];
+t = [0 1 2 3 4 5 6 7];
+jpp = mkpp(t,j);
+app = fnint(jpp, 0);
+vpp = fnint(app, 0);
+spp = fnint(vpp, 0);
+
+tt = linspace(t(1), t(end), 500);
+ss = fnval(spp, tt);
+vv = fnval(vpp, tt);
+aa = fnval(app, tt);
+jj = fnval(jpp, tt);
+
+f = figure
+plot(tt,ss, 'linewidth', 6)
+grid on
+xlabel('t', 'FontSize', 16, 'FontWeight', 'bold')
+ylabel('s(t)', 'FontSize', 16, 'FontWeight', 'bold')
+whitebg('k')
+
+f = figure
+plot(ss,tt, 'linewidth', 6)
+grid on
+xlabel('s', 'FontSize', 16, 'FontWeight', 'bold')
+ylabel('t(s)', 'FontSize', 16, 'FontWeight', 'bold')
+whitebg('k')
+

### 2016-01-11 show_segmentation.py

@@ -1,0 +1,155 @@
+#!/usr/bin/env python
+import sys, os
+from its_utils import *
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from numpy import sqrt
+from itertools import groupby
+
+
+def group(iterable, predicate = lambda x: bool(x)):
+  """groups an iterable by the result of running predicate on each member.
+  groups are not necessarily distinct.
+  @return idx the first index of each group
+  @return val the value of each group
+  """
+  groups = [(0, predicate(iterable[0]))]
+  for idx in range(1, len(iterable)):
+    res = predicate(iterable[idx])
+    if groups[-1][1] == res:
+      continue
+    else:
+      groups.append((idx, res))
+  return zip(*groups)
+
+
+dat = []
+with open(sys.argv[1], 'r') as path_csv:
+  for line in path_csv:
+    for v in line.split(','):
+      dat.append(float(v))
+ugly_path = zip([d - dat[0] for d in dat[0::3]],
+                [d - dat[1] for d in dat[1::3]])
+
+################################################################################
+
+# # fig1 = plt.figure()
+# # ax = fig1.add_subplot(1,1,1)
+# # ax.set_title('Ugly path indices')
+# # ax.hold(True)
+# # for i in range(len(ugly_path)):
+# #   ax.plot(ugly_path[i][0], ugly_path[i][1], 'b.')
+# #   ax.annotate(str(i), xy=ugly_path[i])
+# # ax.grid(True)
+
+# path = pretty_path(ugly_path[15:80], 10, 3, 0.1, 250)
+# s = convert_path_to_1D(path)
+# _, k, _, _ = circular_curvature(path)
+
+
+# fig = plt.figure(facecolor='black')
+# ax = fig.add_subplot(1,1,1, axisbg='k')
+# ax.hold(True)
+# for i in range(len(path)):
+#   if abs(k[i]) < 0.07:
+#     ax.plot(path[i][0], path[i][1], 'b.', markersize=8)
+# for i in range(len(path)):
+#   if abs(k[i]) >= 0.07:
+#     ax.plot(path[i][0],path[i][1], 'r.', markersize=8)
+# 
+# ax.grid(True, color='w')
[…]

### 2016-01-11 show_three_phase.m

@@ -1,0 +1,44 @@
+clear all; close all; clc
+
+j = [-1 0 1];
+t = [0 1 2 3];
+jpp = mkpp(t,j);
+app = fnint(jpp, 0);
+vpp = fnint(app, 2);
+spp = fnint(vpp, 0);
+
+tt = linspace(t(1), t(end), 500);
+ss = fnval(spp, tt);
+vv = fnval(vpp, tt);
+aa = fnval(app, tt);
+jj = fnval(jpp, tt);
+
+f = figure
+subplot(4,1,1)
+plot(tt,ss, 'linewidth', 6)
+grid on
+whitebg('k')
+ylabel('Pos', 'FontSize', 16, 'FontWeight', 'bold')
+subplot(4,1,2)
+plot(tt,vv, 'linewidth', 6)
+grid on
+whitebg('k')
+ylabel('Speed', 'FontSize', 16, 'FontWeight', 'bold')
+subplot(4,1,3)
+plot(tt,aa, 'linewidth', 6)
+grid on
+ylabel('Accel', 'FontSize', 16, 'FontWeight', 'bold')
+whitebg('k')
+subplot(4,1,4)
+plot(tt,jj, 'linewidth', 6)
+grid on
+whitebg('k')
+ylabel('Jerk', 'FontSize', 16, 'FontWeight', 'bold')
+xlabel('Time', 'FontSize', 16, 'FontWeight', 'bold')
+
+f = figure
+plot(ss,vv, 'linewidth', 6)
+grid on
+whitebg('k')
+xlabel('Pos', 'FontSize', 16, 'FontWeight', 'bold')
+ylabel('Speed', 'FontSize', 16, 'FontWeight', 'bold')

### 2016-01-11 stop_parameters.m

@@ -1,0 +1,87 @@
+clear all; close all; clc
+
+j = [-1 0 1];
+t = [0 1 2 3];
+jpp = mkpp(t,j);
+app = fnint(jpp, 0);
+vpp = fnint(app, 2);
+spp = fnint(vpp, 0);
+
+tt = linspace(t(1), t(end), 500);
+ss = fnval(spp, tt);
+vv = fnval(vpp, tt);
+aa = fnval(app, tt);
+jj = fnval(jpp, tt);
+
+t_ = [t(1), t(end)];
+
+
+f = figure;
+
+p1 = subplot(2,2,1);
+hold on
+h = plot(tt,ss, 'linewidth', 6);
+plot(t_, [ss(end), ss(end)], '--g', 'linewidth', 3)
+text(0.05*t(end), 0.95*ss(end), 'L', 'Color', 'g', 'FontSize', 16, 'VerticalAlignment', 'top', 'FontWeight', 'bold')
+grid on
+whitebg('k')
+uistack(h, 'top')
+ylabel('Distance', 'FontSize', 16, 'FontWeight', 'bold')
+xlabel('Time', 'FontSize', 16, 'FontWeight', 'bold')
+
+p2 = subplot(2,2,2);
+hold on
+h = plot(tt,vv, 'w', 'linewidth', 6);
+plot(t_, [vv(1), vv(1)], '--c', 'linewidth', 3)
+text(0.95*t(end), 0.95*vv(1), 'v_i', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'top', 'HorizontalAlignment', 'right', 'FontWeight', 'bold')
+grid on
+whitebg('k')
+uistack(h, 'top')
+ylabel('Speed', 'FontSize', 16, 'FontWeight', 'bold')
+xlabel('Time', 'FontSize', 16, 'FontWeight', 'bold')
+
+p3 = subplot(2,2,3);
+hold on
+h = plot(tt,aa, 'w', 'linewidth', 6);
+plot(t_, [min(aa), min(aa)], '--c', 'linewidth', 3)
+text(0.05*t(end), 1.05*min(aa), 'a_p^-', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'top', 'HorizontalAlignment', 'left', 'FontWeight', 'bold')
+plot(t_, 1.5*[min(aa), min(aa)], '--r', 'linewidth', 3)
[…]

### 2016-01-12 paper1.tex
started making a little headway

@@
 
+\usepackage[pdftex]{graphicx}
+\usepackage[hypcap]{caption}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 Once a trajectory plan is formulated, it must be utilized online in order for a reference signal to be sent to the controllers which govern steering, braking, and engine speed.
-This is the trajectory tracking problem.
+This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
 
@@
 \section{Trajectory Planning}
+\label{sec:trajectoryplanning}
 
 For the purposes of this work, the objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
-Additionally, prescribed initial and end conditions for position, speed, and acceleration must also be honored.
+It is assumed that an independent path planner will produce this information.
+Prescribed initial and end conditions for position, speed, and acceleration must also be honored. 
+This means that the desired trajectory plan is one dimensional, so using curvilinear coordinates will significantly simplify the solution.
 
-\section{Trajectory Tracking}
+% Need some lead in here?
+
+% \subsection{Simple Case}
+
+Consider, as a starting point, the case in which a vehicle is at rest and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+Figure \ref{fig:7phaseprofile} shows this trajectory over time.
+The problem is defined by the vector of known values $\bf{G} = [v_i, v_f, L, v_m]$, being initial speed, final speed, the total distance, and the speed limit, respectively.
+If the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, \Delta t_2, ..., \Delta t_M]$.
+Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
+The user-defined parameters now become $\bar{a}_m = [a^+_m , a^-_m]$ and $\bar{j}_m = [j^+_m , j^-_m]$.
+There are, however multiple solutions for $\Delta t_2$, $\Delta t_4$, and $\Delta t_6$.
+The proper solution must be chosen by eliminating negative and complex solutions.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/7phaseprofile.png}
+  \caption{7 phase constant jerk profile along curvilinear distance axis.}
+  \label{fig:7phaseprofile}
+\end{figure}
+
+It is obvious that there are many cases in which this solution will fail.
+For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\bar{a}_m$ and $\bar{j}_m$.
+The authors then propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\bf{G}$.
+The other profiles are explained below in section \ref{sec:jerkprofiles}.
+A path is then a set of $i$ segments, each of which is defined by a set of parameters $G_i$ and can be solved indivually in sequence.
+Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
+
+\subsection{Jerk Profiles}
+\label{sec:jerkprofiles}
 
@@
 
-\section{Implementation}
+\subsection{Path Segmentation}
+\label{sec:pathsegmentation}
 
-% How much do we really want to put here?
+asdf
 
-% Show tracking performance? It's really more a function of the
+\section{Trajectory Tracking}
+\label{sec:trajectorytracking}
+
+asdf
 

### 2016-01-12 paper1.tex
fleshed out each jerk profile

@@
 
+Paragraphs here
+
+Paragraphs here
+
 % tracking
@@
 For the purposes of this work, the objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
-It is assumed that an independent path planner will produce this information.
+Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
+It is assumed that the path planner operates independently from the trajectory planner and tracker.
 Prescribed initial and end conditions for position, speed, and acceleration must also be honored. 
@@
 % Need some lead in here?
-
 % \subsection{Simple Case}
-
-Consider, as a starting point, the case in which a vehicle is at rest and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
-Figure \ref{fig:7phaseprofile} shows this trajectory over time.
-The problem is defined by the vector of known values $\bf{G} = [v_i, v_f, L, v_m]$, being initial speed, final speed, the total distance, and the speed limit, respectively.
-If the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, \Delta t_2, ..., \Delta t_M]$.
-Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
-The user-defined parameters now become $\bar{a}_m = [a^+_m , a^-_m]$ and $\bar{j}_m = [j^+_m , j^-_m]$.
-There are, however multiple solutions for $\Delta t_2$, $\Delta t_4$, and $\Delta t_6$.
-The proper solution must be chosen by eliminating negative and complex solutions.
 
@@
 
+Consider, as a starting point, the case in which a vehicle is at rest and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+Figure \ref{fig:7phaseprofile} shows this trajectory over time.
+The problem is defined by the vector of known values $\bar{G}  = [v_i, v_f, L, v_m]$, being initial speed, final speed, the total distance, and the speed limit, respectively.
+If the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
+The user-defined parameters now become $\bar{a}_m = [a^+_m , a^-_m]$ and $\bar{j}_m = [j^+_m , j^-_m]$.
+There are, however, multiple solutions for $\Delta t_2$, $\Delta t_4$, and $\Delta t_6$.
+The proper solution must be chosen by eliminating negative and complex solutions.
+
 It is obvious that there are many cases in which this solution will fail.
@@
 The other profiles are explained below in section \ref{sec:jerkprofiles}.
-A path is then a set of $i$ segments, each of which is defined by a set of parameters $G_i$ and can be solved indivually in sequence.
+A path is then a set of $N$ segments, each of which is defined by a set of parameters $G_i, i = 1 ... N$ and can be solved indivually in sequence.
 Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
@@
 
-asdf
+There are a total of five potential jerk profiles which have been created to cover all feasible values in $\bar{G}$. All impossible scenarios are automatically ruled out.
+
+\subsubsection{6 Phase}
+\label{sec:6phase}
+
+The case mentioned above wherein the speed limit is too high for a solution to exist for a 7 phase profile.
+The middle travel period is removed, so that the vehicle increases speed and then immediately decreases speed.
+The peak speed in this case has a solution, but cannot be manipulated directly.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{graphics/6phase_v(s).png}
+  \caption{6 phase profile, displayed as speed vs. position along path}
+  \label{fig:6phaseprofile}
+\end{figure}
+
+\subsubsection{4 Phase}
+\label{sec:4phase}
+
+When $v_i < v_m \simeq v_f$, the vehicle should increase speed, then maintain that speed for the distance remaining in the segment.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{graphics/4phase_v(s).png}
+  \caption{4 phase profile, displayed as speed vs. position along path.}
+  \label{fig:4phaseprofile}
+\end{figure}
+
+\subsubsection{Reversed 4 Phase}
+\label{sec:reversed4phase}
+
+When $v_i \simeq v_m > v_f$, it would be nonsensical to immediately brake. The initial speed is maintained for a period before applying the brakes.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{graphics/4Rphase_all_derivatives.png}
+  \caption{Reversed 4 phase profile, displayed as speed vs. position along path.}
+  \label{fig:4rphaseprofile}
+\end{figure}
+
+\subsubsection{3 Phase}
+\label{sec:3phase}
+
+A change from one speed to another.
+This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
+As such, this profile is chosen for a segment in which it is impossible to completely arrive at the end speed given the initial speed.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.6\columnwidth]{graphics/3phase_v(s).png}
+  \caption{3 phase profile, displayed as speed vs. position along path.}
+  \label{fig:3phaseprofile}
+\end{figure}
 
@@
 
+% make sure to incorporate extensibility to other constraints
 asdf

### 2016-01-13 paper1.tex
added bibliography

@@
 \overrideIEEEmargins                                      % Needed to meet printer requirements.
+
+\usepackage[pdftex]{graphicx}
+\usepackage[hypcap]{caption}
+\usepackage[noadjust]{cite}
 
@@
 }
-
-\usepackage[pdftex]{graphicx}
-\usepackage[hypcap]{caption}
 
@@
 
+\subsection{Jerk Profiles}
+\label{sec:jerkprofiles}
+
 Consider, as a starting point, the case in which a vehicle is at rest and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
@@
 The other profiles are explained below in section \ref{sec:jerkprofiles}.
-A path is then a set of $N$ segments, each of which is defined by a set of parameters $G_i, i = 1 ... N$ and can be solved indivually in sequence.
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $G_k, k = 1 ... N_s$ and can be solved indivually in sequence.
 Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
-
-\subsection{Jerk Profiles}
-\label{sec:jerkprofiles}
 
@@
 
-The case mentioned above wherein the speed limit is too high for a solution to exist for a 7 phase profile.
-The middle travel period is removed, so that the vehicle increases speed and then immediately decreases speed.
+In the case mentioned above, wherein the speed limit is too high for a solution to exist for a 7 phase profile, The middle travel period is removed, so that the vehicle increases speed and then immediately decreases speed.
 The peak speed in this case has a solution, but cannot be manipulated directly.
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{graphics/6phase_v(s).png}
+  \includegraphics[width=0.7\columnwidth]{graphics/6phase_v(s).png}
   \caption{6 phase profile, displayed as speed vs. position along path}
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{graphics/4phase_v(s).png}
+  \includegraphics[width=0.7\columnwidth]{graphics/4phase_v(s).png}
   \caption{4 phase profile, displayed as speed vs. position along path.}
@@
 
-When $v_i \simeq v_m > v_f$, it would be nonsensical to immediately brake. The initial speed is maintained for a period before applying the brakes.
+When $v_i \simeq v_m > v_f$, it would be nonsensical to immediately brake.
+The initial speed is maintained for a period before applying the brakes.
 
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{graphics/4Rphase_all_derivatives.png}
+  \includegraphics[width=0.7\columnwidth]{graphics/4Rphase_all_derivatives.png}
   \caption{Reversed 4 phase profile, displayed as speed vs. position along path.}
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{graphics/3phase_v(s).png}
+  \includegraphics[width=0.7\columnwidth]{graphics/3phase_v(s).png}
   \caption{3 phase profile, displayed as speed vs. position along path.}
@@
 \end{figure}
+
+To decide which profile to employ, a cascade approach is used.
+Analytic values of time intervals are evaluated for each profile by plugging in values of $\bar{G}$, $\bar{a}_m$, and $\bar{j}_m$ until all time intervals are positive and real.
+The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 $\rightarrow$ 4R $\rightarrow$ 3.
 
@@
 % make sure to incorporate extensibility to other constraints
-asdf
+The goal of the segmentation process is to divide a given path into multiple portions where each has a uniform upper limit on speed, acceleration, and braking.
+After this is done, the process described above in Section \ref{sec:jerkprofiles} may be applied to each of the segments.
+
+To begin, 5 values must be computed for each point in the planned path consisting of $N_p$ points: $v_{m,k}$, $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ for $k = 1, ...
+N_p$ .
+Each value is a minimax, i.e., a minimum of several maxima computed from various constraints.
+For instance, ...
+
+For the purposes of this work, two constraints are used for $v_m$: legal speed limit (set constant for the whole path) and lateral acceleration limit.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
+For each path point, the value $v_{m,k}(a_{B,y})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+The value $\kappa_k$ denotes the path curvature at each point.
+For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using recommendations from \cite{Maurya2012}
+An arbitrary set of constraints may
 
@@
 
+\bibliography{HRI_internship_trajectory_planner_articles}
+\bibliographystyle{IEEEtran}
+
 \end{document}

### 2016-01-13 paper1.tex
got the bib to work

@@
 \usepackage[pdftex]{graphicx}
-\usepackage[hypcap]{caption}
+% \usepackage[hypcap]{caption}
 \usepackage[noadjust]{cite}
@@
 The value $\kappa_k$ denotes the path curvature at each point.
-For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using recommendations from \cite{Maurya2012}
+For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using recommendations from ~\cite{Maurya2012}
 An arbitrary set of constraints may
@@
 
+\section*{Acknowledgments}
+
+Thanks.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\bibliographystyle{IEEEtran}
 \bibliography{HRI_internship_trajectory_planner_articles}
-\bibliographystyle{IEEEtran}
 

### 2016-01-13 paper1.tex
removed urls in citations fleshed out the segmentation section added two figures to segmentation section changing notation (WIP)

@@
 \usepackage[pdftex]{graphicx}
-% \usepackage[hypcap]{caption}
+% \usepackage
 \usepackage[noadjust]{cite}
@@
 Figure \ref{fig:7phaseprofile} shows this trajectory over time.
-The problem is defined by the vector of known values $\bar{G}  = [v_i, v_f, L, v_m]$, being initial speed, final speed, the total distance, and the speed limit, respectively.
-If the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+The problem is defined by the vector of known values $\bf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
+If the magnitude of the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
 Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
-The user-defined parameters now become $\bar{a}_m = [a^+_m , a^-_m]$ and $\bar{j}_m = [j^+_m , j^-_m]$.
+The remaining parameters yet to be set now become $\bf{a}_m = [a^+_m , a^-_m]$ and $\bf{j}_m = [j^+_m , j^-_m]$.
 There are, however, multiple solutions for $\Delta t_2$, $\Delta t_4$, and $\Delta t_6$.
@@
 It is obvious that there are many cases in which this solution will fail.
-For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\bar{a}_m$ and $\bar{j}_m$.
-The authors then propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\bf{G}$.
+For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\bf{a}_m$ and $\bf{j}_m$.
+The authors then propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\bf{b}$.
 The other profiles are explained below in section \ref{sec:jerkprofiles}.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $G_k, k = 1 ... N_s$ and can be solved indivually in sequence.
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\bf{b}_k, k = 1 ... N_s$ and can be solved indivually in sequence.
 Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
 
-There are a total of five potential jerk profiles which have been created to cover all feasible values in $\bar{G}$. All impossible scenarios are automatically ruled out.
+There are a total of five potential jerk profiles which have been created to cover all feasible values in $\bf{b}$. All impossible scenarios are automatically ruled out.
 
@@
 
-In the case mentioned above, wherein the speed limit is too high for a solution to exist for a 7 phase profile, The middle travel period is removed, so that the vehicle increases speed and then immediately decreases speed.
+In the case mentioned above, the speed limit is too high for a solution to exist for a 7 phase profile.
+The middle travel period is removed, so that the vehicle increases speed and then immediately decreases speed. 
 The peak speed in this case has a solution, but cannot be manipulated directly.
@@
 To decide which profile to employ, a cascade approach is used.
-Analytic values of time intervals are evaluated for each profile by plugging in values of $\bar{G}$, $\bar{a}_m$, and $\bar{j}_m$ until all time intervals are positive and real.
+Analytic values of time intervals are evaluated for each profile by plugging in values of $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ until all time intervals are positive and real. 
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 $\rightarrow$ 4R $\rightarrow$ 3.
@@
 The goal of the segmentation process is to divide a given path into multiple portions where each has a uniform upper limit on speed, acceleration, and braking.
+There are 3 steps in the segmentation process:
+\begin{enumerate}
+  \item \emph{Ceiling calculation}, where $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ are found for every point in the given path.
+  \item \emph{Clustering}, in which path points are grouped sequentially.
+  \item \emph{Consolidation}, in which a single $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ is selected for each segment.
+\end{enumerate}
 After this is done, the process described above in Section \ref{sec:jerkprofiles} may be applied to each of the segments.
 
-To begin, 5 values must be computed for each point in the planned path consisting of $N_p$ points: $v_{m,k}$, $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ for $k = 1, ...
-N_p$ .
-Each value is a minimax, i.e., a minimum of several maxima computed from various constraints.
-For instance, ...
+The ceiling calculation is a minimax problem.
+Each of the 9 scalars is a minimum of several maxima computed from an arbitrary set of constraints.
+For instance, ... \emph{put examples of things we didn't do...}
+% How we actually did this.
+For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ( $SL$, set constant for the whole path) and lateral acceleration limit ($a_{B,y}^{max}$.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
+For each path point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+The value $\kappa_k$ denotes the path curvature at each point.
+For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
 
-For the purposes of this work, two constraints are used for $v_m$: legal speed limit (set constant for the whole path) and lateral acceleration limit.
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
-For each path point, the value $v_{m,k}(a_{B,y})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
-The value $\kappa_k$ denotes the path curvature at each point.
-For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using recommendations from ~\cite{Maurya2012}
-An arbitrary set of constraints may
+This information is used to choose the segment boundaries in the clustering step.
+Any number of clustering algorithms may be used here, and prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
+For the present example, we will choose a simple univariate heuristic which essentially identifies curves in the roadway.
+Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
+Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
+Figure~\ref{fig:course_highlight_turns} shows the result of this algorithm.
+
+\begin{figure}[thpb]
+  \label{fig:course_highlight_turns}
+  \centering
+  \includegraphics[width=\columnwidth]{graphics/course_highlighted_turns.png}
+  \caption{Clustering output for a short path when using an algorithm which draws segment boundaries around curves.}
+\end{figure}
+
+In the consolidation step, ...
+
+\begin{figure}[thpb]
+  \label{fig:consolidation_speed_ceiling}
+  \centering
+  \includegraphics[width=\columnwidth]{graphics/speed_ceiling.png}
+\end{figure}
+
+The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...}
 

### 2016-01-13 paper1.tex
started first paragraph of tracking section

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:introduction}
+\section{Introduction} \label{sec:introduction}
 
@@
 
-\section{Trajectory Planning}
-\label{sec:trajectoryplanning}
+\section{Trajectory Planning} \label{sec:trajectoryplanning}
 
@@
 This means that the desired trajectory plan is one dimensional, so using curvilinear coordinates will significantly simplify the solution.
-
-% Need some lead in here?
-% \subsection{Simple Case}
 
@@
 
-\subsection{Jerk Profiles}
-\label{sec:jerkprofiles}
+\subsection{Jerk Profiles} \label{sec:jerkprofiles}
 
@@
 
-\subsubsection{6 Phase}
-\label{sec:6phase}
+\subsubsection{6 Phase} \label{sec:6phase}
 
@@
 
-\subsubsection{4 Phase}
-\label{sec:4phase}
+\subsubsection{4 Phase} \label{sec:4phase}
 
@@
 
-\subsubsection{Reversed 4 Phase}
-\label{sec:reversed4phase}
+\subsubsection{Reversed 4 Phase} \label{sec:reversed4phase}
 
@@
 
-\subsubsection{3 Phase}
-\label{sec:3phase}
+\subsubsection{3 Phase} \label{sec:3phase}
 
@@
 
-\subsection{Path Segmentation}
-\label{sec:pathsegmentation}
+\subsection{Path Segmentation} \label{sec:pathsegmentation}
 
@@
 There are 3 steps in the segmentation process:
-\begin{enumerate}
+\begin{enumerate} \label{asdf}
   \item \emph{Ceiling calculation}, where $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ are found for every point in the given path.
@@
 Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
-Figure~\ref{fig:course_highlight_turns} shows the result of this algorithm.
+Figure \ref{fig:course_highlight_turns} shows the result of this algorithm.
 
 \begin{figure}[thpb]
-  \label{fig:course_highlight_turns}
   \centering
@@
   \caption{Clustering output for a short path when using an algorithm which draws segment boundaries around curves.}
+  \label{fig:course_highlight_turns}
 \end{figure}
 
-In the consolidation step, ...
+In the consolidation step, a single $\bf{g}$ must be calculated for each segment using the values of $\bf{g}$ corresponding to each of the path points which comprise the segment in question.
 
 \begin{figure}[thpb]
-  \label{fig:consolidation_speed_ceiling}
   \centering
   \includegraphics[width=\columnwidth]{graphics/speed_ceiling.png}
+  \caption{Example of $v_m$ consolidation for simple univariate path clustering.}
+  \label{fig:consolidation_speed_ceiling}
 \end{figure}
@@
 
-\section{Trajectory Tracking}
-\label{sec:trajectorytracking}
+\section{Trajectory Tracking} \label{sec:trajectorytracking}
 
-asdf
+After planning has been performed, a reference signal must be regularly supplied to a control system in order for the vehicle to carry out the plan.
+This is known as tracking.
+The primary difficulty here is that the trajectory plan has been developed in a one dimensional position space using a reference time which will have unknown drift relative to real time.
+As such, we must find some way to look up the coordinates of the vehicle within the reference trajectory using information typically available for autonomous cars.
+A minimal set of such information would include position, velocity vector, heading, and acceleration in the body frame.
+After this, it is assumed that the vehicle controller will need samples of a small window into the future trajectory in order to pilot the vehicle.
 
@@
 
-Thanks.
+Thanks...
 

### 2016-01-13 paper1.tex
updated bib

@@
 
+\emph{Put kinematic equations here? Many are extrememly large, and I don't have the code.}
+
 \subsection{Path Segmentation} \label{sec:pathsegmentation}
@@
 The value $\kappa_k$ denotes the path curvature at each point.
-For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{graphics/course_highlighted_turns.png}
+  \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
   \caption{Clustering output for a short path when using an algorithm which draws segment boundaries around curves.}
@@
 
-In the consolidation step, a single $\bf{g}$ must be calculated for each segment using the values of $\bf{g}$ corresponding to each of the path points which comprise the segment in question.
+% need better notation here.
+In the consolidation step, a single set of $\bf{g}$, $\bf{a}_m$, and $\bf{j}_m$ must be calculated for each segment using the values corresponding to each of the path points which comprise the segment in question.
+The minimum of scalar across all constituent points is chosen.
+While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
+In the present example, this means that in the straight portions of the path $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entireity of the curve.
+As a result, many of the curves will have a constant speed.
+For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{graphics/speed_ceiling.png}
+  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
   \caption{Example of $v_m$ consolidation for simple univariate path clustering.}
@@
 
-The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...}
+The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...}\tabularnewline
+
+\subsection{Unification} \label{sec:unification}
+
+At this point, the path has been divided into segments and conditions have been computed such that an analytically derived plan can be fit to each segment individually.
+Now a reference trajectory must be generated for the entire path.
 
@@
 After this, it is assumed that the vehicle controller will need samples of a small window into the future trajectory in order to pilot the vehicle.
+
+The general idea will be to project the vehicle's current position onto the path then find the cumulative distance along the path between the start point and the projection point.
+This will be the 1D location corresponding to the vehicle's current reference time.
+Once this reference time is obtained, it may be used to look up position, speed, acceleration, and jerk along the curve of the path over the future time window.
+From there it is a matter of simple geometry to transform these values into the coordinate system of the path and, subsequently, any other coordinate system which may be needed.
+
+% \subsection{Path Projection} \label{sec:pathprojection}
+
+Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity, and acceleration vectors) onto the path.
+A simple approximation, which works well for the path shown in Figure~\ref{fig:course_highlight_turns} is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link''.
+Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
+The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection.
+Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex paths which overlap or self-intersect.
+
+Choosing an algorithm to obtain the current position of the vehicle along the length of the path (i.e., in the same one-dimensional coordinate system that planning was performed) is influenced by the form of the path plan.
+Assuming that the path is linear between waypoints reduces computational effort significantly, but the error in this approximation approaches zero as the spacing between points approaches zero.
+Other approximations which work well include cubic splines, arc splines, and Bezier curves, which have solutions for closest point projection and arc-length calculation~\cite{Wang2002,Wang2003,Schindler2011}.
+For this work, we will examine using the linear approximation between waypoints, meaning that the 1D position of the vehicle at any epoch is the sum of the link lengths behind it's perpendicular projection.
 

### 2016-01-14 paper1.tex
wrote about reverse t(s) lookup spline

@@
 
-Paragraphs here
-
-Paragraphs here
+Lit review...
 
@@
 This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
+
+How is this different from the others ...
 
@@
 Now a reference trajectory must be generated for the entire path.
+Continuity of position, speed, and acceleration is the primary goal of this process.
+
+A jerk vs. time profile is fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
+The value of $a_i$ for the first segment is set to be the value which is currently reported from sensor data.
+Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
+
+% talk about control points
 
@@
 Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
-The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection.
+The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection $\bf{r}_p$.
 Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex paths which overlap or self-intersect.
 
-Choosing an algorithm to obtain the current position of the vehicle along the length of the path (i.e., in the same one-dimensional coordinate system that planning was performed) is influenced by the form of the path plan.
-Assuming that the path is linear between waypoints reduces computational effort significantly, but the error in this approximation approaches zero as the spacing between points approaches zero.
-Other approximations which work well include cubic splines, arc splines, and Bezier curves, which have solutions for closest point projection and arc-length calculation~\cite{Wang2002,Wang2003,Schindler2011}.
-For this work, we will examine using the linear approximation between waypoints, meaning that the 1D position of the vehicle at any epoch is the sum of the link lengths behind it's perpendicular projection.
+Choosing an algorithm to obtain the current one dimensional position of the vehicle along the length of the path $s_{veh}$ (i.e., in the same one-dimensional coordinate system that planning was performed) is influenced by the form of the path plan.
+Assuming that the path is linear between waypoints reduces computational effort significantly.
+So $s_{veh}$ at any epoch is then sum of the link lengths behind it's perpendicular projection$\bf{r}_p$.
+However, the error in this approximation approaches zero as the spacing between path points approaches zero.
+Other, more accurate, approximations which work well include cubic splines, arc splines, and Bezier curves, which have solutions for closest point projection and arc-length calculation~\cite{Wang2002,Wang2003,Schindler2011}.
+
+Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to $v_{B,x}(t)$, $a_{B,x}(t)$, and $j_{B,x}(t)$.
+Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
+Since it is known that $s(t)$ is piecewise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+In this case, an $s(t)$ spline has already been created using the control points from the unification process, and it is now resampled at $\Delta t = 0.01 sec$.
+This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.
+Evaluating the $t(s_{veh})$ then yields the current reference time.
 

### 2016-01-14 paper1.tex
added algorithmic package

@@
 \usepackage[pdftex]{graphicx}
-% \usepackage
+\usepackage{algpseudocode}
 \usepackage[noadjust]{cite}
@@
 Figure \ref{fig:7phaseprofile} shows this trajectory over time.
-The problem is defined by the vector of known values $\bf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
+The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
 If the magnitude of the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
 Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
-The remaining parameters yet to be set now become $\bf{a}_m = [a^+_m , a^-_m]$ and $\bf{j}_m = [j^+_m , j^-_m]$.
+The remaining parameters yet to be set now become $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
 There are, however, multiple solutions for $\Delta t_2$, $\Delta t_4$, and $\Delta t_6$.
@@
 It is obvious that there are many cases in which this solution will fail.
-For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\bf{a}_m$ and $\bf{j}_m$.
-The authors then propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\bf{b}$.
+For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
+The authors then propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
 The other profiles are explained below in section \ref{sec:jerkprofiles}.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\bf{b}_k, k = 1 ... N_s$ and can be solved indivually in sequence.
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved indivually in sequence.
 Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
 
-There are a total of five potential jerk profiles which have been created to cover all feasible values in $\bf{b}$. All impossible scenarios are automatically ruled out.
+There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$. All impossible scenarios are automatically ruled out.
 
@@
 To decide which profile to employ, a cascade approach is used.
-Analytic values of time intervals are evaluated for each profile by plugging in values of $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ until all time intervals are positive and real. 
+Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real. 
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 $\rightarrow$ 4R $\rightarrow$ 3.
 
-\emph{Put kinematic equations here? Many are extrememly large, and I don't have the code.}
+\emph{Put kinematic equations here? Many are extrememly large, and I don't have the code. Perhaps better to summarize what was put into the symbolic solver.}
 
@@
 \begin{enumerate} \label{asdf}
-  \item \emph{Ceiling calculation}, where $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ are found for every point in the given path.
+  \item \emph{Ceiling calculation}, where $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ are found for every point in the given path.
   \item \emph{Clustering}, in which path points are grouped sequentially.
-  \item \emph{Consolidation}, in which a single $\bf{b}$, $\bf{a}_m$, and $\bf{j}_m$ is selected for each segment.
+  \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
 \end{enumerate}
@@
 % need better notation here.
-In the consolidation step, a single set of $\bf{g}$, $\bf{a}_m$, and $\bf{j}_m$ must be calculated for each segment using the values corresponding to each of the path points which comprise the segment in question.
+In the consolidation step, a single set of $\mathbf{g}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ must be calculated for each segment using the values corresponding to each of the path points which comprise the segment in question.
 The minimum of scalar across all constituent points is chosen.
@@
 
-\subsection{Unification} \label{sec:unification}
+\subsection{Unified Solution} \label{sec:unifiedsolution}
 
@@
 A jerk vs. time profile is fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
-The value of $a_i$ for the first segment is set to be the value which is currently reported from sensor data.
+Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
+The values of $a_i$ and $v_i$ for the first segment is set to be the value which is currently reported from sensor data.
 Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
+For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
+To set boundary speed conditions for the interior segments,
+
+\begin{figure}
+  \begin{algorithmic}[1]
+    \Procedure{SegmentBoundarySpeeds}{$v_m$}
+      \State $\mathbf{v}_p \gets \mathbf{v}_m$ \Comment{Initialize targets as ceilings}
+      \State 
+      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Final speed is 0}
+      \For{$k = 1, ..., N_s-2$}
+      \EndFor
+    \EndProcedure
+  \end{algorithmic}
+\caption{Algorithm to set speed boundary conditions for interior path segments}
+\label{fig:segmentspeedboundaryconditions}
+\end{figure}
 
@@
 Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
-The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection $\bf{r}_p$.
+The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection $\mathbf{r}_p$.
 Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex paths which overlap or self-intersect.
@@
 Assuming that the path is linear between waypoints reduces computational effort significantly.
-So $s_{veh}$ at any epoch is then sum of the link lengths behind it's perpendicular projection$\bf{r}_p$.
+So $s_{veh}$ at any epoch is then sum of the link lengths behind it's perpendicular projection$\mathbf{r}_p$.
 However, the error in this approximation approaches zero as the spacing between path points approaches zero.
@@
 Since it is known that $s(t)$ is piecewise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
-In this case, an $s(t)$ spline has already been created using the control points from the unification process, and it is now resampled at $\Delta t = 0.01 sec$.
+In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
 This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.

### 2016-01-14 paper1.tex
unified solution section close to finished

@@
 For instance, ... \emph{put examples of things we didn't do...}
+
 % How we actually did this.
@@
 For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
-To set boundary speed conditions for the interior segments,
+To set boundary speed conditions for the interior segments, Algorithm~\ref{alg:segmentspeedboundaryconditions}
 
@@
       \State $\mathbf{v}_p \gets \mathbf{v}_m$ \Comment{Initialize targets as ceilings}
-      \State 
-      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Final speed is 0}
-      \For{$k = 1, ..., N_s-2$}
+      \State $\mathbf{v}_{p,0} \gets v_i$ \Comment{Set starting speed to current}
+      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
+      \For{$k = 1, ..., N_s-1$}
+        \If{$v_{m,i+1} > v_{m,i}$}
+          \State $v_{p,i+1} \gets v_{m,i}$
+        \ElsIf{$v_{m,i+1} > v_{m,i}$}
+          \State $v_{p,i+1} \gets v_{m,i+1}$
+        \Else
+          \State $v_{p,i+1} \gets v_{m,i}$
+        \EndIf
       \EndFor
@@
 \caption{Algorithm to set speed boundary conditions for interior path segments}
-\label{fig:segmentspeedboundaryconditions}
+\label{alg:segmentspeedboundaryconditions}
 \end{figure}
@@
 % talk about control points
+Now each segment is solved individually.
+In the case where a segment requires a speed change which is infeasible given limits on acceleration and jerk, then the final speed for that segment is planned to be the closes possible value.
+When these discrepancies in the feasible $v_f$ for a segment arise, the corresponding value in $v_p$ is updated so that the new conditions may be accounted for in the subsequent segment.
+
+At the end of the process, there will be a set of $N_{cp}$ ``control points''.
+A control point is a vector $[t_{ref}, s, v_{B,x}, a_{B,x}, j_{B,x}]$ which describes the longitudinal kinematic state of the target vehicle at the moment of an infinite jounce jerk change.
+Fitting a jerk profile to each segment results in control points which must be stitched into the control points for the preceding segments.
+While no further information than initial conditions, time durations, and jerk values is needed (simple integration would yield lower derivatives of position), this information is already available since it must be solved for each segment.
+This extra information may be used as a redundancy layer to verify that no errors occured.
+For each segment with $M$ jerk phases, the following is produced:
+\begin{itemize}
+  \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
+  \item $M+1$ values of distance elapsed within the segment $\Delta s$.
+These must be shifted forward by the last $s$ value from the previous segment to translate them into the 1D path domain.
+Additionally, the first value must be erased.
+  \item $M+1$ values of $v_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
+  \item $M+1$ values of $a_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
+  \item $M$ jerk values, which may be appended directly to the preceding jerk values
+\end{itemize}
 

### 2016-01-14 paper1.tex
finished up the unification section

@@
 
+After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
+Since $j(t)$ is piecewise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
+To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
+Position along the path $s(t)$ takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
+Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
+Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
+The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
+
+However, the time at which the trajectory plan must be referenced is not always known.
+The solution to this problem is known as trajectory tracking, and will be discussed in the next section.
+
 \section{Trajectory Tracking} \label{sec:trajectorytracking}
@@
 
+\section{Implementation}
+
+\emph{How much should we say here?}
+
 \section*{Acknowledgments}

### 2016-01-14 paper1.tex
ready??

@@
 For the purposes of this work, the objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
+Elevation changes will not be addressed here.
 Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
@@
 This means that the desired trajectory plan is one dimensional, so using curvilinear coordinates will significantly simplify the solution.
+It will be shown how two dimensional concerns (specifically cornering speeds for this work) may be expressed in one dimension to guide planning.
+All planning will be longitidinal; that is, speed will be the primary focus, and it is assumed that a separate yaw controller will ensure that the origin of the vehicle.
 
@@
 The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
+It is assume that the final forward acceleration will be zero.
 If the magnitude of the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
@@
 Each of the 9 scalars is a minimum of several maxima computed from an arbitrary set of constraints.
-For instance, ... \emph{put examples of things we didn't do...}
+For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the path.
 
 % How we actually did this.
-For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ( $SL$, set constant for the whole path) and lateral acceleration limit ($a_{B,y}^{max}$.
+For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ($SL$, set constant for the whole path) and lateral acceleration limit ($a_{B,y}^{max}$.
 Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
@@
 Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity, and acceleration vectors) onto the path.
-A simple approximation, which works well for the path shown in Figure~\ref{fig:course_highlight_turns} is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link''.
+A simple approximation, which works well for the path shown in Figure~\ref{fig:course_highlight_turns} is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'').
 Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.

### 2016-01-19 paper1.tex
fixed tense in intro and added description to segmentation figure

@@
   \thanks{
-    * This work is supported by Honda Research Institute USA.
+    * This work was performed at Honda Research Institute USA.
   }
   \thanks{
-    $^{1}$ Robert G. Cofield is a graduate researcher at Auburn University's GPS \& Vehicle Dynamics Laboratory.
+    $^{1}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
   }
   \thanks{
-    $^{2}$ Rakesh Gupta ...
+    $^{2}$ Honda Research Institute USA
   }
   \thanks{
-    $^{3}$ Ambarish Goswami ... 
+    $^{3}$ Honda Research Institute USA
   }
   \thanks{
-    $^{4}$ David Bevly ...
+    $^{4}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
   }
@@
 % We will use the path-velocity decomposition
-% mention using curvilinear CS
-%
 
@@
 Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
-It is assumed that the path planner operates independently from the trajectory planner and tracker.
+We assume that the path planner operates independently from the trajectory planner and tracker.
 Prescribed initial and end conditions for position, speed, and acceleration must also be honored. 
 This means that the desired trajectory plan is one dimensional, so using curvilinear coordinates will significantly simplify the solution.
-It will be shown how two dimensional concerns (specifically cornering speeds for this work) may be expressed in one dimension to guide planning.
-All planning will be longitidinal; that is, speed will be the primary focus, and it is assumed that a separate yaw controller will ensure that the origin of the vehicle.
+We show how two dimensional concerns (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
+All planning is longitudinal; that is, speed is the primary focus, and we assume that a separate yaw controller will ensure that the vehicle follows the prescribed path.
 
@@
   \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
-  \caption{Clustering output for a short path when using an algorithm which draws segment boundaries around curves.}
+  \caption{Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
+Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).}
   \label{fig:course_highlight_turns}
@@
   \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
-  \caption{Example of $v_m$ consolidation for simple univariate path clustering.}
+  \caption{Example of $v_m$ consolidation for simple univariate path clustering which corresponds the the path shown in Figure \ref{fig:course_highlight_turns}.
+The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
+The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
+The resultant speed ceiling is shown in blue.}
   \label{fig:consolidation_speed_ceiling}

### 2016-01-19 paper1.tex
replace unlabeled full 7 phase diagram with the full spec version

@@
   \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/7phaseprofile.png}
+  \includegraphics[width=1.0\columnwidth]{graphics/Full7PhaseSpecVertical.png}
   \caption{7 phase constant jerk profile along curvilinear distance axis.}
-  \label{fig:7phaseprofile}
+  \label{fig:full7phasespec}
 \end{figure}
@@
 
-Consider, as a starting point, the case in which a vehicle is at rest and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
-Figure \ref{fig:7phaseprofile} shows this trajectory over time.
+Consider, as a starting point, the case in which a vehicle is moving at some small initial speed and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+Figure \ref{fig:full7phasespec} shows this trajectory over time.
 The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
@@
 The proper solution must be chosen by eliminating negative and complex solutions.
+As we will show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
+To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs.
+\ref{eq:am} and \ref{eq:jm}.
+This is computed for every point along the path.
+In the simplest case, one value for each of the quantities expressed in those relations may be chosen for the entire path.
+
+\begin{equation}
+  a^{+,max}_m >= a^+_m > 0 > a^-_m >= a^{-,max}_m
+  \label{eq:am}
+\end{equation}
+\begin{equation}
+  j^{+,max}_m >= j^+_m > 0 > j^-_m >= j^{-,max}_m
+  \label{eq:jm}
+\end{equation}
 
@@
 For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
-The authors then propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
-The other profiles are explained below in section \ref{sec:jerkprofiles}.
+We propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
+The other profiles are explained below in Section \ref{sec:jerkprofiles}.
 A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved indivually in sequence.
@@
 There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$. All impossible scenarios are automatically ruled out.
+To find closed-form solutions for each, a symbolic solver was used. The target values for each profile are the time steps between jerk changes.
+These are solved by plugging in initial conditions, default jerk values
+
+For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
+This results in negative time intervals for that phase.
+In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
+Algorithm \ref{alg:modifyacceljerkpeaks} describes this procedure.
+When it fails, a new jerk profile must be chosen.
+
+\begin{figure}
+  \begin{algorithmic}[1]
+    \Procedure{ModifyAccelJerkPeaks}{$\Delta t$, $\mathbf{a}_m$, $\mathbf{j}_m$, $\mathbf{a}^{max}_m$, $\mathbf{j}^{max}_m$}
+
+    \EndProcedure
+  \end{algorithmic}
+\caption{Procedure for modifying peak accel and jerk magnitude for rememdy negative time intervals for a given phase. This procedure is employed for each of the profiles.}
+\label{alg:modifyacceljerkpeaks}
+\end{figure}
+
+To decide which profile to employ, a cascade approach is used.
+Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real. 
+The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 $\rightarrow$ 4R $\rightarrow$ 3.
+
+\subsubsection{7 Phase} \label{sec:7phase}
+
+This profile, described above, is always attempted first.
+All time steps have a single solution, enabled by the fact that the initial and end conditions are known, and the two peak accelerations and two jerk magnitudes are provided as configuration parameters.
 
@@
 A change from one speed to another.
+All of the above profiles are composed of some combination of a 3 phase profile and a $j_k = 0$ phase.
 This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
@@
 \end{figure}
-
-To decide which profile to employ, a cascade approach is used.
-Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real. 
-The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 $\rightarrow$ 4R $\rightarrow$ 3.
-
-\emph{Put kinematic equations here? Many are extrememly large, and I don't have the code. Perhaps better to summarize what was put into the symbolic solver.}
 
@@
 
-The general idea will be to project the vehicle's current position onto the path then find the cumulative distance along the path between the start point and the projection point.
-This will be the 1D location corresponding to the vehicle's current reference time.
+The general idea is to project the vehicle's current position onto the path then find the cumulative distance along the path between the start point and the projection point.
+This is the 1D location corresponding to the vehicle's current reference time.
 Once this reference time is obtained, it may be used to look up position, speed, acceleration, and jerk along the curve of the path over the future time window.

### 2016-01-19 scripts/general_segment_solving_vertical.m

@@ -1,0 +1,144 @@
+%%
+% clear all; close all; clc
+% 
+% j = [1 0 -1 0 -1 0 1];
+% t = [0 1 2 3 4 5 6 7];
+% jpp = mkpp(t,j);
+% app = fnint(jpp, 0);
+% vpp = fnint(app, 0);
+% spp = fnint(vpp, 0);
+% 
+% tt = linspace(t(1), t(end), 500);
+% ss = fnval(spp, tt);
+% vv = fnval(vpp, tt);
+% aa = fnval(app, tt);
+% jj = fnval(jpp, tt);
+% 
+% f = figure
+% subplot(4,1,1)
+% plot(tt,ss, 'linewidth', 6)
+% grid on
+% whitebg('k')
+% ylabel('Pos', 'FontSize', 16, 'FontWeight', 'bold')
+% title('7 Phase Profile')
+% subplot(4,1,2)
+% plot(tt,vv, 'linewidth', 6)
+% grid on
+% whitebg('k')
+% ylabel('Speed', 'FontSize', 16, 'FontWeight', 'bold')
+% subplot(4,1,3)
+% plot(tt,aa, 'linewidth', 6)
+% grid on
+% ylabel('Accel', 'FontSize', 16, 'FontWeight', 'bold')
+% whitebg('k')
+% subplot(4,1,4)
+% plot(tt,jj, 'linewidth', 6)
+% grid on
+% whitebg('k')
+% ylabel('Jerk', 'FontSize', 16, 'FontWeight', 'bold')
+% xlabel('Time', 'FontSize', 16, 'FontWeight', 'bold')
+
+%%
+clear all; close all; clc
+
+j = [1 0 -1 0 -1 0 1];
+t = [0 .5 1.5 2.5 4 5 6.9+[0 1]];
+jpp = mkpp(t,j);
+app = fnint(jpp, .5);
+vpp = fnint(app, 1);
+spp = fnint(vpp, 0);
+
+tt = linspace(t(1), t(end), 500);
+ss = fnval(spp, tt);
+vv = fnval(vpp, tt);
+aa = fnval(app, tt);
+jj = fnval(jpp, tt);
+
+t_ = [t(1), t(end)];
+
+f = figure;
+p1 = subplot(4,1,1);
+hold on
+h = plot(tt,ss, 'linewidth', 6);
+plot(t_, [15 15], '--c', 'linewidth', 3)
+text(1, 14, 'L', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'top')
[…]

### 2016-01-19 paper1.tex
removed a/j tuning algorithm and simply stated it textually

@@
 The remaining parameters yet to be set now become $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
-There are, however, multiple solutions for $\Delta t_2$, $\Delta t_4$, and $\Delta t_6$.
-The proper solution must be chosen by eliminating negative and complex solutions.
 As we will show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
@@
 
-There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$. All impossible scenarios are automatically ruled out.
-To find closed-form solutions for each, a symbolic solver was used. The target values for each profile are the time steps between jerk changes.
-These are solved by plugging in initial conditions, default jerk values
+There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$.
+All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, travelling speed of zero, negative speeds in $\mathbf{b}$, etc).
+It should be noted that travelling backward (i.e., in reverse gear) is not supported in this work.
+To find closed-form solutions for each, a symbolic solver was used.
+The target values for each profile are the time steps between jerk changes.
+These are solved by plugging in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$.
 
 For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
-This results in negative time intervals for that phase.
+This results in a negative time interval for that phase.
 In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
-Algorithm \ref{alg:modifyacceljerkpeaks} describes this procedure.
-When it fails, a new jerk profile must be chosen.
-
-\begin{figure}
-  \begin{algorithmic}[1]
-    \Procedure{ModifyAccelJerkPeaks}{$\Delta t$, $\mathbf{a}_m$, $\mathbf{j}_m$, $\mathbf{a}^{max}_m$, $\mathbf{j}^{max}_m$}
-
-    \EndProcedure
-  \end{algorithmic}
-\caption{Procedure for modifying peak accel and jerk magnitude for rememdy negative time intervals for a given phase. This procedure is employed for each of the profiles.}
-\label{alg:modifyacceljerkpeaks}
-\end{figure}
+Specifically, jerk may be increased while peak acceleration is decreased.
+If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.
 

### 2016-01-19 paper1.tex
updated indiviual jerk profile sections

@@
 Specifically, jerk may be increased while peak acceleration is decreased.
+This process is referred to herein as ``accel-jerk tuning''
 If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.
@@
 All time steps have a single solution, enabled by the fact that the initial and end conditions are known, and the two peak accelerations and two jerk magnitudes are provided as configuration parameters.
+In the case where $\Delta t_2$ or $\Delta t_6$ are negative, accel-jerk tuning must be employed.
+In the instance where $\Delta t_4$ is negative, the total segment length $L$ is too short to reach top speed of $v_m$, so a 6 phase profile must be used.
 
@@
 In the case mentioned above, the speed limit is too high for a solution to exist for a 7 phase profile.
-The middle travel period is removed, so that the vehicle increases speed and then immediately decreases speed. 
+The middle travel period $\Delta t_4$ is removed, so that the vehicle increases speed and then immediately decreases speed. 
 The peak speed in this case has a solution, but cannot be manipulated directly.
+Without this known value, the problem becomes underdefined, so $\Delta t_2$ and $\Delta t_5$ each have two solutions: one negative and one positive.
+The positive solution is chosen.
+In the case where both solutions are negative, accel-jerk tuning is attempted. 
+Should that yield no positive time intervals, one of the two 4 phase profiles then becomes appropriate, since $L$ is not large enough to accomodate accelerating past the end speed.
+If $v_i > v_f$, then a reversed 4 phase profile is attempted.
+However, if $v_i > v_f$, then a regular 4 phase profile is attempted.
 
@@
   \includegraphics[width=0.7\columnwidth]{graphics/4phase_v(s).png}
-  \caption{4 phase profile, displayed as speed vs. position along path.}
+  \caption{4 phase profile, displayed as speed vs. position along path within the segment.}
   \label{fig:4phaseprofile}
@@
   \includegraphics[width=0.7\columnwidth]{graphics/4Rphase_all_derivatives.png}
-  \caption{Reversed 4 phase profile, displayed as speed vs. position along path.}
+  \caption{Reversed 4 phase profile, displayed as speed vs. position along path within the segment.}
   \label{fig:4rphaseprofile}
@@
 This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
-As such, this profile is chosen for a segment in which it is impossible to completely arrive at the end speed given the initial speed.
+As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4 phase profiles has been attempted and subsequently failed.
 

### 2016-01-19 scripts/general_segment_solving_vertical.m
redid notation in 7phase graphic, and made it bigger

@@ -89,13 +89,13 @@ p3 = subplot(4,1,3);
 hold on
 h = plot(tt,aa, 'linewidth', 6);
 plot([0 3], [1 1], '--c', 'linewidth', 3)
-text(3, 1, 'a_p^+', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'top', 'FontWeight', 'bold')
+text(3, 1, 'a_m^+', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'top', 'FontWeight', 'bold')
 plot([0 4], [1.5 1.5], '--r', 'linewidth', 3)
-text(4, 1.5, 'a_m^+', 'Color', 'r', 'FontSize', 16, 'VerticalAlignment', 'top', 'FontWeight', 'bold')
+text(4, 1.5, 'a_m^{+,max}', 'Color', 'r', 'FontSize', 16, 'VerticalAlignment', 'top', 'FontWeight', 'bold')
 plot([2.5 8], [-1.5, -1.5], '--r', 'linewidth', 3)
-text(2.5, -1.5, 'a_m^- ', 'Color', 'r', 'FontSize', 16, 'VerticalAlignment', 'bottom', 'FontWeight', 'bold', 'HorizontalAlignment', 'right')
+text(2.5, -1.5, 'a_m^{-,max} ', 'Color', 'r', 'FontSize', 16, 'VerticalAlignment', 'bottom', 'FontWeight', 'bold', 'HorizontalAlignment', 'right')
 plot([3.5 8], [-1 -1], '--c', 'linewidth', 3)
-text(3.5, -1, 'a_p^-', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'bottom', 'FontWeight', 'bold', 'HorizontalAlignment', 'right')
+text(3.5, -1, 'a_m^-', 'Color', 'c', 'FontSize', 16, 'VerticalAlignment', 'bottom', 'FontWeight', 'bold', 'HorizontalAlignment', 'right')
 grid on
 ylabel('Accel', 'FontSize', 16, 'FontWeight', 'bold')
 uistack(h, 'top')
@@ -107,28 +107,28 @@ p4 = subplot(4,1,4);
 hold on
 h = plot(tt,jj, 'linewidth', 6);
 plot([0 2.5], [1 1], '--c', 'linewidth', 3)
[…]

### 2016-01-19 paper1.tex
redid notation in 7phase graphic, and made it bigger
removed Ambarish as author

@@
   Rakesh Gupta $^{2}$,
-  Ambarish Goswami $^{3}$,
   and
@@
     $^{1}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
-  }
-  \thanks{
-    $^{2}$ Honda Research Institute USA
   }
@@
   \includegraphics[width=1.0\columnwidth]{graphics/Full7PhaseSpecVertical.png}
-  \caption{7 phase constant jerk profile along curvilinear distance axis.}
+  \caption{7 phase constant jerk profile along curvilinear distance axis. All values needed to completely parameterize a single path segment for trajectory are labelled here.}
   \label{fig:full7phasespec}

### 2016-01-20 paper2.tex
added paper2, not yet merged into single paper

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.

\usepackage[pdftex]{graphicx}
\usepackage{algpseudocode}
\usepackage[noadjust]{cite}

\title{\LARGE \bf
  Reactive Path Planning with Pedestrains for Autonomous Driving
}

\author{
  Robert G. Cofield $^{1}$,
  Rakesh Gupta $^{2}$,
  Ambarish Goswami $^{3}$,
  and
  David Bevly $^{4}$
  \thanks{
    * This work was performed at Honda Research Institute USA.
  }
  \thanks{
    $^{1}$ Robert G. Cofield is a graduate researcher at Auburn University's GPS \& Vehicle Dynamics Laboratory.
  }
  \thanks{
    $^{2}$ Rakesh Gupta ...
  }
  \thanks{
    $^{3}$ Ambarish Goswami ... 
  }
  \thanks{
    $^{4}$ David Bevly ...
  }
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Minimize of navigation time from origin to the destination implies two tasks that can be carried out sequentially or in parallel. First is Path Planning which involves computation of the shortest path. Second task is Velocity Planning (or Trajectory Generation) which involves the computation of the fastest velocity profile that is smooth and comfortable for the human occupant.

For human comfort, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort. A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}. Some preliminary work on optimizing these constraints on urban roads has been done ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene. On the other hand systems that find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} do not work as part of an integrated system that can navigate urban roads. In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort as well is reactive to presence of pedestrians on the road.

\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:introduction}

In this paper we present an autonomy framework for pedestrian aware ground vehicles. We describe the objective and architecture.
Then we describe other subsystems that our system needs to work with, and planned stops. We then describe
reactivity to pedestrians in the environment by extending the trajectory planner. We then summarize and discuss future extensions.
Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done
at very low speeds (5 miles/hour). There has also been some work in simulating vehicles and modeling
uncertainty, but pedestrians are less predictable than cars and can appear suddenly on the road 
(when someone is jaywalking for example).

Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway. We assume that there is no other traffic on the roadway. Desired behavior from the car includes following traffic laws at the intersections such as stop signs. At stop signs the car should stop and wait for an appropriate amount of time. It should also react to pedestrians on the road - at intersections as well as on the road. If pedestrians are not on the road they should be ignored by the autonomous cars. Further the pedestrians can be static or moving. The car should stop, slow down and circum-navigate around pedestrians when appropriate.

In urban scenarios we need to handle intersections, turns in the road and planned stops due 
to stop signs (that have already been marked in the map). In addition we exhibit reactionary 
behavior and stop of pedestrians that may appear on the road. We handle this via a high 
level path manager that calls the trajectory planner as necessary as shown in Figure \ref{fig:addreact}.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/MissingReactionPiece2.png}
  \caption{Adding reactive behavior to trajectory planning}
  \label{fig:addreact}
\end{figure}

\section{Input Subsystems} \label{sec:inputsubsystems}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/zenrin.png}
  \caption{High resolution map. Red color is at road edges. Green color shows lane center-lines}
  \label{fig:map}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/GoMentumSatelliteSmall.png}
  \caption{Gomentum map top view}
  \label{fig:gomentum}
\end{figure}

We assume that we have access to Path Planner. This can be for example a service similar to
Google Maps except that it provides lane level maps. We also assume that we have a High
Resolution digital map such as one shown in Figure \ref{fig:map} [corresponding to 
Figure \ref{fig:gomentum}] with Lane centers, Stop 
signs, crosswalk geometry, and Turning curves at intersections. The starting position the 
car is determined via high grade GPS. The destination is manually input into the system.
Using the path planner, we obtain a path along lane centers from start to the destination.
This plan is output as waypoints, with the waypoints corresponding to stop signs marked.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/ped-detection-screenshot.png}
  \caption{Showing detected pedestrian by the computer vision system}
  \label{fig:ped}
\end{figure}

%\begin{figure}[thpb]
%  \centering
%  \includegraphics[width=1.0\columnwidth]{graphics/3023096-slide-c4carafter.png}
%  \caption{Pedestrain at the intersection}
%  \label{fig:ped2}
%\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/screenshot-objectmap.png}
  \caption{Showing car with road and path covered so far. The red filled circle shows the pedestrian for which the car is stopping}
  \label{fig:car}
\end{figure}

Pedestrian Detection is performed via a Camera, and LiDAR system. It computes a bounding 
box for each pedestrian as shown in Figure \ref{fig:ped}. The pedestrian position is 
transformed to the vehicle body frame, and subsequently transformed to global frame and 
can be shown on the map as in Figure \ref{fig:car}.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/IntersectionCrosswalks.png}
  \caption{A car turning right at the intersection only needs to care about pedestrians in the 
pedestrian crossings shown yellow. Pedestrians in the green crossings can be ignored unless they
appear somewhere on the road}
  \label{fig:intersect}
\end{figure}

We also have an In-Out Algorithm implemented via polygon intersection that can be used to
ignore “safe” pedestrians that are on sidewalks, off-path crosswalks, and behind planned stop
as shown in Figure \ref{fig:intersect}.
Remaining pedestrians pose collision risk and we need to reactively slow down or stop for them.
This is accomplished by sending stop message with pedestrian locations to path manager.

For Navigation Filtering we use an ADMA-G based Commercial automotive GPS/INS system.
It receives RTK corrections via cell modem. It outputs Global position, global heading,
body frame velocity and body frame acceleration to 10s of cms.

\section{Path Manager} \label{sec:pathmanager}

Path Pre-processor conforms the path to planner assumptions. The Path Manager breaks up the
path at stop signs, decides when to stop, when to go and invokes the trajectory planner
as necessary.

There are two types of stops - Planned (PSTOP) and Reactive (RSTOP). At reactive stops
to pedestrians, the car may slow down and restart if the pedestrian leaves before the
car comes to a stop; otherwise the car would come to a stop and restarts once the 
pedestrian leaves.

The state of the car is represented by a State Machine as shown in figure \ref{fig:st}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
  \caption{State Transition Diagram}
  \label{fig:st}
\end{figure}

The assumption on our work is that there is a unique mapping from a time to s coordinate 
system. This mapping breaks down when we are stopped at a planned or reactive stop. 
The solution is to break the path into sub-paths, and plan and execute incrementally.
This is shown in Figures \ref{fig:segmentation} \ref{fig:stcur} and \ref{fig:scoord}.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
  \caption{Segmentation of the path by path manager. The Trajectory generation is serially called on each segment colored differently. Need to update figure with stop signs????????????????}
  \label{fig:segmentation}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/course_highlighted_turns.png}
  \caption{Further characterization of green path shown in segmentation figure into straight and curved parts}
  \label{fig:stcur}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
  \caption{Showing the path projected onto the one dimensional s coordinate system}
  \label{fig:scoord}
\end{figure}

The algorithm is:
\begin{itemize}
\item Send first sub-path to trajectory planner
\item Wait until completed
\item Pause for fixed time
\item Send next sub-path to planner
\item Repeat until end of path
\end{itemize}

For unplanned, reactive stops  we need to slow down and stop. There are three scenarios:
\begin{enumerate}
\item Normal pedestrian → appears
\item Stopping pedestrian → moves
\item Stopped pedestrian leaves →
\end{enumerate}

In the next section we describe when/where to begin braking, when/where to stop
and when to go again.

\section{Reactive Subsystem} \label{sec:reactiveubsystem}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL.png}
  \caption{Normal mode}
  \label{fig:react1}
\end{figure}
\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}
  \caption{Stopping for the pedestrian mode}
  \label{fig:react2}
\end{figure}
\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}
  \caption{Resuming after stop sign}
  \label{fig:react3}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/3phase_all_derivatives_SlowingDown.png}
  \caption{3 phase R stop}
  \label{fig:3phaserstop}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
  \caption{Green color path is the one over which the car finds a trajectory to stop}
  \label{fig:pathtostop}
\end{figure}

For reactive stopping, if there are multiple pedestrian ahead, we only consider the
closest pedestrian. That pedestrian is projected on the path of the car to determine the
required braking distance. The stopping trajectory is similar to rest of profile,
and consists of a single segment with 3 phases as shown in Figure \ref{fig:3phase}. 
The trajectory planner uses total distance as parameter, and is not suitable for
our purpose as the end speed constraints are not met. So we reformulate and 
replace $L$ with $v_f$ as hard constraint.

Typically while with the Trajectory Planner, we get initial conditions from the
navigation filter. We use normal (sedate) peak parameters for acceleration
and jerk. We plan a 3 phase stop profile with the current required stopping distance.

Once a pedestrian enters the stop zone, we need to plan, and execute RSTOP trajectory.
When available distance <= “required” distance, we increase acceleration, jerk peaks to
match available distance.

After a reactive stop we split the path before the stop from the one after the stop
sign and execute the later as a separate trajectory. The plan is stored and sampled 
the same as in normal operation.

To get back to normal driving after we are fully stopped, we resume
after pedestrian leaves roadway, or pedestrian moves far down roadway.
When he moved further down, we stop again later after we are a threshold
distance away. The effect is that we follow the person at the distance with some 
delay in chunks of few meters each time. If the pedestrian leaves before 
the car comes to a full stop, the car can speed up without stopping using the 
same algorithm.

\section{Summary} \label{sec:summary}

We have built a robust stopping model as part of Trajectory planner reliably integrated with path manager.
We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
they leave. Our system handles corner cases gracefully, and works with multiple pedestrians as we only 
need to consider the closest pedestrian ahead of us. In future, we want to handle other roadway traffic, 
coexisting with other vehicles. We would also like to perceive and stop for red-lights.

\section{Examples}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/FullStopSpec.png}
  \caption{Example 3 phase jerk profile}
  \label{fig:3phase}
\end{figure}

\section{2d to 1d conversion}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/PathProjection.png}
  \caption{Showing projection of current position of car shown by green triangle to s coordinate system}
  \label{fig:cartos}
\end{figure}
\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/PathProjectionSlice.png}
  \caption{car to s coord with additional points}
  \label{fig:cartos2}
\end{figure}
\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/s(t)_generic.png}
  \caption{Transformation from s coordinate system to time}
  \label{fig:stot}
\end{figure}

\section{Other Issues}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/LowResPathSample.png}
  \caption{Grid discretization issue}
  \label{fig:discretization}
\end{figure}
\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/linestring.png}
  \caption{Exaggerated view of path points which may have discontinuous curvature}
  \label{fig:disc}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/CircularCurvature.png}
  \caption{circular curvature}
  \label{fig:curv}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=0.5\columnwidth]{graphics/speed_ceiling.png}
  \caption{Speed Ceiling}
  \label{fig:speedceil}
\end{figure}

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}
  \caption{Reaction Response: Normal Mode}
  \label{fig:7phaseprofile}
\end{figure}

\section{Implementation}

\emph{How much should we say here?}

\section*{Acknowledgments}

Thanks...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{IEEEtran}
\bibliography{HRI_internship_trajectory_planner_articles}

\end{document}

### 2016-01-20 paper2.tex
remove gomentum images

@@
 \usepackage{algpseudocode}
+\usepackage{subfigure}
 \usepackage[noadjust]{cite}
@@
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/zenrin.png}
-  \caption{High resolution map. Red color is at road edges. Green color shows lane center-lines}
-  \label{fig:map}
-\end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/GoMentumSatelliteSmall.png}
-  \caption{Gomentum map top view}
-  \label{fig:gomentum}
-\end{figure}
-
 We assume that we have access to Path Planner. This can be for example a service similar to
 Google Maps except that it provides lane level maps. We also assume that we have a High
-Resolution digital map such as one shown in Figure \ref{fig:map} [corresponding to 
-Figure \ref{fig:gomentum}] with Lane centers, Stop 
-signs, crosswalk geometry, and Turning curves at intersections. The starting position the 
+Resolution digital map with Lane centers, Stop signs, crosswalk geometry, and Turning 
+curves at intersections. The starting position the 
 car is determined via high grade GPS. The destination is manually input into the system.
@@
   \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/ped-detection-screenshot.png}
+  \includegraphics[width=0.5\columnwidth]{graphics/ped2.png}
   \caption{Showing detected pedestrian by the computer vision system}
@@
 \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL.png}
-  \caption{Normal mode}
-  \label{fig:react1}
-\end{figure}
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}
-  \caption{Stopping for the pedestrian mode}
-  \label{fig:react2}
-\end{figure}
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}
-  \caption{Resuming after stop sign}
-  \label{fig:react3}
+  \subfigure[Normal mode]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL.png}}
+\\
+  \subfigure[Stopping for the pedestrian]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}}
+\\
+  \subfigure[Resuming after stop sign]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}}
+  \caption{Reactive behavior to pedestrain}
+  \label{fig:react}
 \end{figure}

### 2016-01-20 paper1.tex
added section buffer comments
began merging two documents, about to reevaluate.

@@
 \title{\LARGE \bf
-  Longitudinal Trajectory Planning and Tracking for Fixed-Path Autonomous Driving
+  % Longitudinal Trajectory Planning and Tracking for Fixed-Path Autonomous Driving
+  % Reactive Path Planning with Pedestrains for Autonomous Driving
+  Reactive Trajectory Planning and Tracking for Pedestrian-Aware Autonomous Driving on Fixed Paths
 }
@@
 
-When planning trajectories for an autonomous ground vehicle operating on roadways, it is common to employ the path-velocity decomposition (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
+When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
+It is common to employ the path-velocity decomposition to create trajectory plans, meaning there are two tasks which can be carried out sequentially or in parallel.
+% (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
+First is path planning which involves computation of the shortest path.
+The second task is velocity planning (or trajectory generation) which involves the computation of the fastest velocity profile that is appropriate for the cargo or passengers.
+
 Given a desired path, we present a novel method of planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
@@
 Each segement is then planned sequentially by choosing a piecewise constant jerk profile.
-The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous throughout the entire path.
-The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control.
-This method is shown to be efficient and reliable for use in online planning with a test vehicle.
+The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
+The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is trajectory tracking.
+We then present an online method of revising the trajectory plan to safely avoid pedestrians who enter the roadway by slowing or stopping when no alternative paths are available.
+This framework is shown to be efficient and reliable for use in online planning with a test vehicle and real pedestrians, while optimizing rider comfort.
 
@@
 These may include legal speed limits, engine dynamics, desired sideslip limits, traction availability, and rollover concerns.
+
+For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort.
+A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
+Some preliminary work on optimizing these constraints on urban roads has been done ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
+On the other hand, many systems in the literature which find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} do not work as part of an integrated system that can navigate urban roads.
+In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort as well as reactive to presence of pedestrians on the road.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Trajectory Planning} \label{sec:trajectoryplanning}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Path Segmentation} \label{sec:pathsegmentation}
@@
 The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...}\tabularnewline
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Trajectory Tracking} \label{sec:trajectorytracking}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Implementation}
@@
 \emph{How much should we say here?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-01-21 paper1.tex
rough integration of two papers

@@
   and
-  David Bevly $^{4}$
+  David Bevly $^{3}$
   \thanks{
@@
   \thanks{
-    $^{4}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
+    $^{3}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
   }
@@
 First is path planning which involves computation of the shortest path.
-The second task is velocity planning (or trajectory generation) which involves the computation of the fastest velocity profile that is appropriate for the cargo or passengers.
+The second task is velocity planning (or trajectory generation) which involves the computation of the fastest velocity profile that is appropriate for the scenario.
 
@@
 These may include legal speed limits, engine dynamics, desired sideslip limits, traction availability, and rollover concerns.
+Also of paramount importance is the comfort of the passengers.
+
+In this paper we present an autonomy framework for pedestrian aware ground vehicles.
+We describe the objective and architecture.
+Then we describe other subsystems that our system needs to work with, and planned stops.
+We then describe
+reactivity to pedestrians in the environment by extending the trajectory planner.
+We then summarize and discuss future extensions.
+Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done
+at very low speeds (5 miles/hour).
+There has also been some work in simulating vehicles and modeling
+uncertainty, but pedestrians are less predictable than cars and can appear suddenly on the road 
+(when someone is jaywalking for example).
+
+Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
+We assume that there is no other traffic on the roadway.
+Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
+At stop signs the car should stop and wait for an appropriate amount of time.
+It should also react to pedestrians on the road - at intersections as well as on the road.
+If pedestrians are not on the road they should be ignored by the autonomous cars.
+Further the pedestrians can be static or moving.
+The car should stop, slow down and circum-navigate around pedestrians when appropriate.
+
+In urban scenarios we need to handle intersections, turns in the road and planned stops due 
+to stop signs (that have already been marked in the map).
+In addition we exhibit reactionary 
+behavior and stop of pedestrians that may appear on the road.
+We handle this via a high 
+level path manager that calls the trajectory planner as necessary as shown in Figure \ref{fig:addreact}.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/MissingReactionPiece2.png}
+  \caption{
+    Adding reactive behavior to trajectory planning.
+    \newline
+    \textbf{!!!!!!!!!!!!!!!!!!!!!!REDO THIS!!!!!!!!!!!!!!!!!!}
+  }
+  \label{fig:addreact}
+\end{figure}
 
@@
 
-% intersection of path-planning constraints and speed-planning constraints is vehicle dynamics
-
-% review of literature for a few paragraphs]
-%  - searching planners (OMPL stuff)
-%  - robot joint planning --> constant jerk intervals
-
-% We will use the path-velocity decomposition
-
-Lit review...
-
 % tracking
@@
 
-How is this different from the others ...
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% mention using same cuvilinear cs scheme to project onto path
+\section{Input Subsystems} \label{sec:inputsubsystems}
+
+We assume that we have access to Path Planner. This can be for example a service similar to
+Google Maps except that it provides lane level maps. We also assume that we have a High
+Resolution digital map with Lane centers, Stop signs, crosswalk geometry, and Turning 
+curves at intersections. The starting position the 
+car is determined via high grade GPS. The destination is manually input into the system.
+Using the path planner, we obtain a path along lane centers from start to the destination.
+This plan is output as waypoints, with the waypoints corresponding to stop signs marked.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/ped2.png}
+  \caption{Showing detected pedestrian by the computer vision system
+  \newline
+  \textbf{!!!!!!!!Crop to save space!!!!!!!!!!!!!!!}
+  }
+  \label{fig:ped}
+\end{figure}
+
+%\begin{figure}[thpb]
+%  \centering
+%  \includegraphics[width=1.0\columnwidth]{graphics/3023096-slide-c4carafter.png}
+%  \caption{Pedestrain at the intersection}
+%  \label{fig:ped2}
+%\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/screenshot-objectmap.png}
+  \caption{Showing car with road and path covered so far. The red filled circle shows the pedestrian for which the car is stopping}
+  \label{fig:car}
+\end{figure}
+
+Pedestrian Detection is performed via a Camera, and LiDAR system. It computes a bounding 
+box for each pedestrian as shown in Figure \ref{fig:ped}. The pedestrian position is 
+transformed to the vehicle body frame, and subsequently transformed to global frame and 
+can be shown on the map as in Figure \ref{fig:car}.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/IntersectionCrosswalks.png}
+  \caption{A car turning right at the intersection only needs to care about pedestrians in the 
+pedestrian crossings shown yellow. Pedestrians in the green crossings can be ignored unless they
+appear somewhere on the road}
+  \label{fig:intersect}
+\end{figure}
+
+We also have an In-Out Algorithm implemented via polygon intersection that can be used to
+ignore “safe” pedestrians that are on sidewalks, off-path crosswalks, and behind planned stop
+as shown in Figure \ref{fig:intersect}.
+Remaining pedestrians pose collision risk and we need to reactively slow down or stop for them.
+This is accomplished by sending stop message with pedestrian locations to path manager.
+
+For Navigation Filtering we use an ADMA-G based Commercial automotive GPS/INS system.
+It receives RTK corrections via cell modem. It outputs Global position, global heading,
+body frame velocity and body frame acceleration to 10s of cms.
 
@@
   \includegraphics[width=1.0\columnwidth]{graphics/Full7PhaseSpecVertical.png}
-  \caption{7 phase constant jerk profile along curvilinear distance axis. All values needed to completely parameterize a single path segment for trajectory are labelled here.}
+  \caption{
+    7 phase constant jerk profile along curvilinear distance axis.
+    All values needed to completely parameterize a single path segment for trajectory are labelled here.}
   \label{fig:full7phasespec}
@@
   \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
-  \caption{Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
-Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).}
+  \caption{
+    Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
+    Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
+  }
   \label{fig:course_highlight_turns}
@@
   \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
-  \caption{Example of $v_m$ consolidation for simple univariate path clustering which corresponds the the path shown in Figure \ref{fig:course_highlight_turns}.
+  \caption{
+    Example of $v_m$ consolidation for simple univariate path clustering which corresponds the the path shown in Figure \ref{fig:course_highlight_turns}.
     The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
     The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
-The resultant speed ceiling is shown in blue.}
+    The resultant speed ceiling is shown in blue.
+  }
   \label{fig:consolidation_speed_ceiling}
@@
 
+\section{Path Manager} \label{sec:pathmanager}
+
+Path Pre-processor conforms the path to planner assumptions. The Path Manager breaks up the
+path at stop signs, decides when to stop, when to go and invokes the trajectory planner
+as necessary.
+
+There are two types of stops - Planned (PSTOP) and Reactive (RSTOP). At reactive stops
+to pedestrians, the car may slow down and restart if the pedestrian leaves before the
+car comes to a stop; otherwise the car would come to a stop and restarts once the 
+pedestrian leaves.
+
+The state of the car is represented by a State Machine as shown in figure \ref{fig:st}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
+  \caption{State Transition Diagram}
+  \label{fig:st}
+\end{figure}
+
+The assumption on our work is that there is a unique mapping from a time to s coordinate 
+system. This mapping breaks down when we are stopped at a planned or reactive stop. 
+The solution is to break the path into sub-paths, and plan and execute incrementally.
+This is shown in Figures \ref{fig:segmentation} \ref{fig:stcur} and \ref{fig:scoord}.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
+  \caption{Segmentation of the path by path manager. The Trajectory generation is serially called on each segment colored differently. Need to update figure with stop signs????????????????}
+  \label{fig:segmentation}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/course_highlighted_turns.png}
+  \caption{Further characterization of green path shown in segmentation figure into straight and curved parts}
+  \label{fig:stcur}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
+  \caption{Showing the path projected onto the one dimensional s coordinate system}
+  \label{fig:scoord}
+\end{figure}
+
+The algorithm is:
+\begin{itemize}
+\item Send first sub-path to trajectory planner
+\item Wait until completed
+\item Pause for fixed time
+\item Send next sub-path to planner
+\item Repeat until end of path
+\end{itemize}
+
+For unplanned, reactive stops  we need to slow down and stop. There are three scenarios:
+\begin{enumerate}
+\item Normal pedestrian → appears
+\item Stopping pedestrian → moves
+\item Stopped pedestrian leaves →
+\end{enumerate}
+
+In the next section we describe when/where to begin braking, when/where to stop
+and when to go again.
+
+\section{Reactive Subsystem} \label{sec:reactiveubsystem}
+
+\begin{figure}[thpb]
+\centering
+  \subfigure[Normal mode]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL.png}}
+\\
+  \subfigure[Stopping for the pedestrian]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}}
+\\
+  \subfigure[Resuming after stop sign]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}}
+  \caption{Reactive behavior to pedestrain}
+  \label{fig:react}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/3phase_all_derivatives_SlowingDown.png}
+  \caption{3 phase R stop}
+  \label{fig:3phaserstop}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
+  \caption{Green color path is the one over which the car finds a trajectory to stop}
+  \label{fig:pathtostop}
+\end{figure}
+
+For reactive stopping, if there are multiple pedestrian ahead, we only consider the
+closest pedestrian. That pedestrian is projected on the path of the car to determine the
+required braking distance. The stopping trajectory is similar to rest of profile,
+and consists of a single segment with 3 phases as shown in Figure \ref{fig:3phase}. 
+The trajectory planner uses total distance as parameter, and is not suitable for
+our purpose as the end speed constraints are not met. So we reformulate and 
+replace $L$ with $v_f$ as hard constraint.
+
+Typically while with the Trajectory Planner, we get initial conditions from the
+navigation filter. We use normal (sedate) peak parameters for acceleration
+and jerk. We plan a 3 phase stop profile with the current required stopping distance.
+
+Once a pedestrian enters the stop zone, we need to plan, and execute RSTOP trajectory.
+When available distance <= “required” distance, we increase acceleration, jerk peaks to
+match available distance.
+
+After a reactive stop we split the path before the stop from the one after the stop
+sign and execute the later as a separate trajectory. The plan is stored and sampled 
+the same as in normal operation.
+
+To get back to normal driving after we are fully stopped, we resume
+after pedestrian leaves roadway, or pedestrian moves far down roadway.
+When he moved further down, we stop again later after we are a threshold
+distance away. The effect is that we follow the person at the distance with some 
+delay in chunks of few meters each time. If the pedestrian leaves before 
+the car comes to a full stop, the car can speed up without stopping using the 
+same algorithm.
+
+\section{Summary} \label{sec:summary}
+
+We have built a robust stopping model as part of Trajectory planner reliably integrated with path manager.
+We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
+they leave. Our system handles corner cases gracefully, and works with multiple pedestrians as we only 
+need to consider the closest pedestrian ahead of us. In future, we want to handle other roadway traffic, 
+coexisting with other vehicles. We would also like to perceive and stop for red-lights.
+
+\section{Examples}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/FullStopSpec.png}
+  \caption{Example 3 phase jerk profile}
+  \label{fig:3phase}
+\end{figure}
+
+\section{2d to 1d conversion}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/PathProjection.png}
+  \caption{Showing projection of current position of car shown by green triangle to s coordinate system}
+  \label{fig:cartos}
+\end{figure}
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/PathProjectionSlice.png}
+  \caption{car to s coord with additional points}
+  \label{fig:cartos2}
+\end{figure}
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/s(t)_generic.png}
+  \caption{Transformation from s coordinate system to time}
+  \label{fig:stot}
+\end{figure}
+
+\section{Other Issues}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/LowResPathSample.png}
+  \caption{Grid discretization issue}
+  \label{fig:discretization}
+\end{figure}
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/linestring.png}
+  \caption{Exaggerated view of path points which may have discontinuous curvature}
+  \label{fig:disc}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/CircularCurvature.png}
+  \caption{circular curvature}
+  \label{fig:curv}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/speed_ceiling.png}
+  \caption{Speed Ceiling}
+  \label{fig:speedceil}
+\end{figure}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}
+  \caption{Reaction Response: Normal Mode}
+  \label{fig:7phaseprofile}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Implementation}

### 2016-01-21 paper1.tex
a lot of cutting out figures

@@
 For Navigation Filtering we use an ADMA-G based Commercial automotive GPS/INS system.
-It receives RTK corrections via cell modem. It outputs Global position, global heading,
+It receives RTK corrections via cell modem.
+It outputs Global position, global heading,
 body frame velocity and body frame acceleration to 10s of cms.
@@
 Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
-Figure \ref{fig:course_highlight_turns} shows the result of this algorithm.
+Figure \ref{fig:1Dto2DandSegmentation} shows the result of this algorithm.
+
+% \begin{figure}[thpb]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
+%   \caption{
+%     Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
+%     Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
+%   }
+%   \label{fig:course_highlight_turns}
+% \end{figure}
+
+% \begin{figure}[thpb]
+%   \centering
+%   \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
+%   \caption{Showing the path projected onto the one dimensional s coordinate system}
+%   \label{fig:scoord}
+% \end{figure}
+
+% \begin{figure}[thpb]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
+%   \caption{
+%     Example of $v_m$ consolidation for simple univariate path clustering which corresponds the the path shown in Figure \ref{fig:course_highlight_turns}.
+%     The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
+%     The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
+%     The resultant speed ceiling is shown in blue.
+%   }
+%   \label{fig:consolidation_speed_ceiling}
+% \end{figure}
 
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
-  \caption{
+  \subfigure[
     Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
     Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
+  ]{
+    \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
   }
-  \label{fig:course_highlight_turns}
+  \\
+  \subfigure[
+    Showing the path projected onto the one dimensional s coordinate system
+  ]{
+    \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
+  }
+  \\
+  \subfigure[
+    Example of $v_m$ consolidation for simple univariate path clustering which corresponds to the path shown above.
+    The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
+    The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
+    The resultant speed ceiling is shown in blue.
+  ]{
+  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
+  }
+  \caption{\textbf{Cut down the subfigure captions and put it here!!!}}
+  \label{fig:1Dto2DandSegmentation}
 \end{figure}
@@
 For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
-  \caption{
-    Example of $v_m$ consolidation for simple univariate path clustering which corresponds the the path shown in Figure \ref{fig:course_highlight_turns}.
-    The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
-    The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
-    The resultant speed ceiling is shown in blue.
-  }
-  \label{fig:consolidation_speed_ceiling}
-\end{figure}
 
@@
 
+\subsection{2d to 1d conversion}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/PathProjection.png}
+  \caption{Showing projection of current position of car shown by green triangle to s coordinate system}
+  \label{fig:cartos}
+\end{figure}
+
 The general idea is to project the vehicle's current position onto the path then find the cumulative distance along the path between the start point and the projection point.
@@
 
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/s(t)_generic.png}
+  \caption{
+    Transformation from s coordinate system to reference time.
+    The figure depicts s(t) for a 3 phase jerk profile with positive forward acceleration.
+  }
+  \label{fig:stot}
+\end{figure}
+
 % \subsection{Path Projection} \label{sec:pathprojection}
@@
 Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity, and acceleration vectors) onto the path.
-A simple approximation, which works well for the path shown in Figure~\ref{fig:course_highlight_turns} is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'').
+A simple approximation, which works well for the path shown in Figure~\ref{fig:1Dto2DandSegmentation} is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'').
 Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
@@
 
-Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to $v_{B,x}(t)$, $a_{B,x}(t)$, and $j_{B,x}(t)$.
+\subsection{Control Reference Signal} \label{sec:controlreferencesignal}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/PathProjectionSlice.png}
+  \caption{car to s coord with additional points}
+  \label{fig:cartos2}
+\end{figure}
+
+Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to look up $v_{B,x}(t_{veh})$, $a_{B,x}(t_{veh})$, and $j_{B,x}(t_{veh})$.
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
@@
 
-Path Pre-processor conforms the path to planner assumptions. The Path Manager breaks up the
+Path Pre-processor conforms the path to planner assumptions.
+The path manager breaks up the
 path at stop signs, decides when to stop, when to go and invokes the trajectory planner
@@
 
-There are two types of stops - Planned (PSTOP) and Reactive (RSTOP). At reactive stops
+There are two types of stops: planned (PSTOP) and reactive (RSTOP).
+At reactive stops
 to pedestrians, the car may slow down and restart if the pedestrian leaves before the
@@
 pedestrian leaves.
-
 The state of the car is represented by a State Machine as shown in figure \ref{fig:st}
@@
 The assumption on our work is that there is a unique mapping from a time to s coordinate 
-system. This mapping breaks down when we are stopped at a planned or reactive stop. 
-The solution is to break the path into sub-paths, and plan and execute incrementally.
-This is shown in Figures \ref{fig:segmentation} \ref{fig:stcur} and \ref{fig:scoord}.
+system.
+This mapping breaks down when we are stopped at a planned or reactive stop, as there is now way to do a $t(s)$ when multiple reference time solutions exist for certain path positions.
+The solution is to break the path into sub-paths, and plan and execute them incrementally as shown in Fig. \ref{fig:segmentation}.
 
@@
   \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
-  \caption{Segmentation of the path by path manager. The Trajectory generation is serially called on each segment colored differently. Need to update figure with stop signs????????????????}
+  \caption{Segmentation of the path by path manager. The trajectory generation is serially called on each segment colored differently. \textbf{Need to update figure with stop signs????????????????}
+  }
   \label{fig:segmentation}
@@
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/course_highlighted_turns.png}
-  \caption{Further characterization of green path shown in segmentation figure into straight and curved parts}
-  \label{fig:stcur}
-\end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
-  \caption{Showing the path projected onto the one dimensional s coordinate system}
-  \label{fig:scoord}
-\end{figure}
-
-The algorithm is:
+The procedure is:
 \begin{itemize}
@@
 
-For unplanned, reactive stops  we need to slow down and stop. There are three scenarios:
+For unplanned, reactive stops  we need to slow down and stop. 
+There are three scenarios:
 \begin{enumerate}
-\item Normal pedestrian → appears
-\item Stopping pedestrian → moves
-\item Stopped pedestrian leaves →
+\item \emph{Normal}: pedestrian appears
+\item \emph{Stopping}: pedestrian moves
+\item \emph{Stopped}: pedestrian leaves
 \end{enumerate}
@@
 \centering
-  \subfigure[Normal mode]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL.png}}
+  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
   \\
-  \subfigure[Stopping for the pedestrian]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}}
+  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}}
   \\
@@
   \label{fig:react}
-\end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/3phase_all_derivatives_SlowingDown.png}
-  \caption{3 phase R stop}
-  \label{fig:3phaserstop}
 \end{figure}
@@
 
-\section{Examples}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/FullStopSpec.png}
-  \caption{Example 3 phase jerk profile}
-  \label{fig:3phase}
-\end{figure}
-
-\section{2d to 1d conversion}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/PathProjection.png}
-  \caption{Showing projection of current position of car shown by green triangle to s coordinate system}
-  \label{fig:cartos}
-\end{figure}
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/PathProjectionSlice.png}
-  \caption{car to s coord with additional points}
-  \label{fig:cartos2}
-\end{figure}
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/s(t)_generic.png}
-  \caption{Transformation from s coordinate system to time}
-  \label{fig:stot}
-\end{figure}
-
-\section{Other Issues}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/LowResPathSample.png}
-  \caption{Grid discretization issue}
-  \label{fig:discretization}
-\end{figure}
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/linestring.png}
-  \caption{Exaggerated view of path points which may have discontinuous curvature}
-  \label{fig:disc}
-\end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/CircularCurvature.png}
-  \caption{circular curvature}
-  \label{fig:curv}
-\end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/speed_ceiling.png}
-  \caption{Speed Ceiling}
-  \label{fig:speedceil}
-\end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}
-  \caption{Reaction Response: Normal Mode}
-  \label{fig:7phaseprofile}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Implementation}
-
-\emph{How much should we say here?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Acknowledgments}
-
-Thanks...
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-01-21 root.tex
removed paper2, renamed paper1 to root as per IEEE convention

--- 2016-01-21 paper1.tex
@@
 
-\subsection{2d to 1d conversion}
-
 \begin{figure}[thpb]
@@
 Other, more accurate, approximations which work well include cubic splines, arc splines, and Bezier curves, which have solutions for closest point projection and arc-length calculation~\cite{Wang2002,Wang2003,Schindler2011}.
-
-\subsection{Control Reference Signal} \label{sec:controlreferencesignal}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/PathProjectionSlice.png}
-  \caption{car to s coord with additional points}
-  \label{fig:cartos2}
-\end{figure}
 
@@
 
-\section{Reactive Subsystem} \label{sec:reactiveubsystem}
-
 \begin{figure}[thpb]
@@
   \\
-  \subfigure[Resuming after stop sign]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}}
+  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}}
   \caption{Reactive behavior to pedestrain}
@@
 For reactive stopping, if there are multiple pedestrian ahead, we only consider the
-closest pedestrian. That pedestrian is projected on the path of the car to determine the
-required braking distance. The stopping trajectory is similar to rest of profile,
-and consists of a single segment with 3 phases as shown in Figure \ref{fig:3phase}. 
-The trajectory planner uses total distance as parameter, and is not suitable for
-our purpose as the end speed constraints are not met. So we reformulate and 
-replace $L$ with $v_f$ as hard constraint.
+closest pedestrian.
+That pedestrian is projected on the path of the car to determine the required braking distance.
+The stopping trajectory is similar to rest of profile, and consists of a single segment with 3 phases as shown in Figure \textbf{reference! include 3 phase full spec profile?}. 
+The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met. So we reformulate and replace $L$ with $v_f$ as hard constraint.
 
-Typically while with the Trajectory Planner, we get initial conditions from the
-navigation filter. We use normal (sedate) peak parameters for acceleration
-and jerk. We plan a 3 phase stop profile with the current required stopping distance.
+Typically while with the Trajectory Planner, we get initial conditions from the navigation filter.
+We use normal (sedate) peak parameters for acceleration and jerk.
+We plan a 3 phase stop profile with the current required stopping distance.
 
 Once a pedestrian enters the stop zone, we need to plan, and execute RSTOP trajectory.
-When available distance <= “required” distance, we increase acceleration, jerk peaks to
-match available distance.
+When available distance is less than “required” distance, we increase acceleration, jerk peaks to match available distance.
 

### 2016-01-21 root.tex
renamed bib file
finished rough draft of integrated paper

@@
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/screenshot-objectmap.png}
-  \caption{Showing car with road and path covered so far. The red filled circle shows the pedestrian for which the car is stopping}
-  \label{fig:car}
-\end{figure}
-
 Pedestrian Detection is performed via a Camera, and LiDAR system. It computes a bounding 
 box for each pedestrian as shown in Figure \ref{fig:ped}. The pedestrian position is 
-transformed to the vehicle body frame, and subsequently transformed to global frame and 
-can be shown on the map as in Figure \ref{fig:car}.
+transformed to the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
 
@@
 Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real.
-The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 $\rightarrow$ 4R $\rightarrow$ 3.
+The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
 
@@
   \\
-  \subfigure[
-    Showing the path projected onto the one dimensional s coordinate system
-  ]{
-    \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
-  }
-  \\
+  % \subfigure[
+  %   Showing the path projected onto the one dimensional s coordinate system
+  % ]{
+  %   \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
+  % }
+  % \\
   \subfigure[
@@
   }
-  \caption{\textbf{Cut down the subfigure captions and put it here!!!}}
+  \caption{Path segmentation and speed ceiling consolidation procedures.}
   \label{fig:1Dto2DandSegmentation}
@@
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/s(t)_generic.png}
-  \caption{
-    Transformation from s coordinate system to reference time.
-    The figure depicts s(t) for a 3 phase jerk profile with positive forward acceleration.
-  }
-  \label{fig:stot}
-\end{figure}
-
 % \subsection{Path Projection} \label{sec:pathprojection}
@@
 
-Path Pre-processor conforms the path to planner assumptions.
-The path manager breaks up the
-path at stop signs, decides when to stop, when to go and invokes the trajectory planner
-as necessary.
+The assumption on our work is that there is a unique mapping from a reference time $t_{ref}$ to a scalar value indicating the distance along the current path, $s$.
+system.
+This mapping breaks down when we are stopped at a planned or reactive stop, as there is no way to do a $t_{ref}(s)$ lookup when multiple reference time solutions exist for certain path positions.
+The solution is to break the path into sub-paths at each stop location (i.e., stop sign) which is known a priori, and plan then execute them incrementally as shown in Fig. \ref{fig:segmentation}.
+This is the first function of the path manager, and can be broken down as follows:
+\begin{itemize}
+\item Send first sub-path to trajectory planner
+\item Wait until completed.
+\item Pause for fixed time, typically 2-3 second.
+\item Continue waiting if any pedestrians are in the roadway ahead (along the intended path) within some fixed distance.
+\item Send next sub-path to planner.
+\item Repeat until end of path is reached.
+\end{itemize}
 
-There are two types of stops: planned (PSTOP) and reactive (RSTOP).
-At reactive stops
-to pedestrians, the car may slow down and restart if the pedestrian leaves before the
-car comes to a stop; otherwise the car would come to a stop and restarts once the 
-pedestrian leaves.
-The state of the car is represented by a State Machine as shown in figure \ref{fig:st}
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
+  \caption{
+    Segmentation of the path by path manager.
+    The trajectory generation is serially called on each segment colored differently.
+    \textbf{Need to update figure with stop signs???? ??????? ?????}
+  }
+  \label{fig:segmentation}
+\end{figure}
+
+Both behaviors dictate that the path manager govern the operation of the trajectory planner.
+To enable both of these features, the path manager is implemented as a discrete state machine, as shown in Fig. \ref{fig:st}.
 
@@
 
-The assumption on our work is that there is a unique mapping from a time to s coordinate 
-system.
-This mapping breaks down when we are stopped at a planned or reactive stop, as there is now way to do a $t(s)$ when multiple reference time solutions exist for certain path positions.
-The solution is to break the path into sub-paths, and plan and execute them incrementally as shown in Fig. \ref{fig:segmentation}.
+The distinction between momentarily braking for a pedestrian which quickly leaves the roadway and coming to a full stop needs no formal logic.
+They follow the same state transition, and since both trajectory plans are extremely smooth, riders are rarely aware that any change took place.
+It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
-  \caption{Segmentation of the path by path manager. The trajectory generation is serially called on each segment colored differently. \textbf{Need to update figure with stop signs????????????????}
-  }
-  \label{fig:segmentation}
-\end{figure}
-
-The procedure is:
-\begin{itemize}
-\item Send first sub-path to trajectory planner
-\item Wait until completed
-\item Pause for fixed time
-\item Send next sub-path to planner
-\item Repeat until end of path
-\end{itemize}
-
-For unplanned, reactive stops  we need to slow down and stop. 
-There are three scenarios:
-\begin{enumerate}
-\item \emph{Normal}: pedestrian appears
-\item \emph{Stopping}: pedestrian moves
-\item \emph{Stopped}: pedestrian leaves
-\end{enumerate}
-
-In the next section we describe when/where to begin braking, when/where to stop
-and when to go again.
+Next, the triggers for state transition will be described.
+They are implemented as a set of static and dynamic thresholds, which are depicted graphically in Fig. \ref{fig:react}.
 
@@
 
+All detected pedestrians locations are compared against the high-resolution digital map, so that pedestrians which are not on the roadway or near an upcoming crosswalk are not considered.
+Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
+Pedestrians which are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
+All the pedestrians which remain pose a potential collision risk, however only the closest is considered for RSTOP.
+
+Next the nominal stopping distance $Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
+The procedure for doing this will be the same as the procedure for planning the RSTOP later with a few steps omitted, since the planning algorithm is extremely efficient.
+A 3 phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+
+While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
+As was mentioned in previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
+Normal planning chooses enforces $L$ to avoid skewing segment boundaries within each subpath.
+However, reactive stopping requires that $v_f$ be enforced, and $L$ be a parameter which may be solved dependently.
+Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
+%%% This is the brief way of wording the preceding three sentences.
+% The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met.
+So we reformulate and replace $L$ with $v_f$ as the hard constraint.
+
+\textbf{3 phase full spec profile figure?}.
+
+Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.
+For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used.
+Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
+RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= s_{ped} - s_{veh}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
+The trajectory is then recomputed with increasingly higher-magnitude values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
+Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
+
+After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
+The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
+Within the trajectory tracker, all portions of the path leading up to the location of RSTOP tiggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
+This is depicted in Fig. \ref{fig:pathtostop}
+The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
+
 \begin{figure}[thpb]
@@
 
-For reactive stopping, if there are multiple pedestrian ahead, we only consider the
-closest pedestrian.
-That pedestrian is projected on the path of the car to determine the required braking distance.
-The stopping trajectory is similar to rest of profile, and consists of a single segment with 3 phases as shown in Figure \textbf{reference! include 3 phase full spec profile?}. 
-The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met. So we reformulate and replace $L$ with $v_f$ as hard constraint.
-
-Typically while with the Trajectory Planner, we get initial conditions from the navigation filter.
-We use normal (sedate) peak parameters for acceleration and jerk.
-We plan a 3 phase stop profile with the current required stopping distance.
-
-Once a pedestrian enters the stop zone, we need to plan, and execute RSTOP trajectory.
-When available distance is less than “required” distance, we increase acceleration, jerk peaks to match available distance.
-
-After a reactive stop we split the path before the stop from the one after the stop
-sign and execute the later as a separate trajectory. The plan is stored and sampled 
-the same as in normal operation.
-
-To get back to normal driving after we are fully stopped, we resume
-after pedestrian leaves roadway, or pedestrian moves far down roadway.
-When he moved further down, we stop again later after we are a threshold
-distance away. The effect is that we follow the person at the distance with some 
-delay in chunks of few meters each time. If the pedestrian leaves before 
-the car comes to a full stop, the car can speed up without stopping using the 
-same algorithm.
+This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
+This value is typically set to 1 meter.
+Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$.
+The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
+The preceding buffers are depicted graphically in Fig. \ref{fig:react}.
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{HRI_internship_trajectory_planner_articles}
+\bibliography{root}
 

### 2016-01-21 root.tex
update abstract, intro and summary

@@
   % Reactive Path Planning with Pedestrains for Autonomous Driving
-  Reactive Trajectory Planning and Tracking for Pedestrian-Aware Autonomous Driving on Fixed Paths
+  Reactive Trajectory Planning and Tracking Control for Pedestrian-Aware Autonomous Driving in Urban Environments
 }
@@
 \author{
-  Robert G. Cofield $^{1}$,
-  Rakesh Gupta $^{2}$,
-  and
-  David Bevly $^{3}$
+  Robert G. Cofield $^{1}$ and
+  Rakesh Gupta $^{2}$
   \thanks{
@@
   \thanks{
-    $^{3}$ Honda Research Institute USA
-  }
-  \thanks{
-    $^{3}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
+    $^{2}$ Honda Research Institute USA Inc, 375 Ravendale Drive, Mountain View, CA 94043. rgupta@hra.com
   }
@@
 
-When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
-It is common to employ the path-velocity decomposition to create trajectory plans, meaning there are two tasks which can be carried out sequentially or in parallel.
-% (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-First is path planning which involves computation of the shortest path.
-The second task is velocity planning (or trajectory generation) which involves the computation of the fastest velocity profile that is appropriate for the scenario.
+In this paper, we address the problem of trajectory planning for urban roads that satisfies 
+constraints for human comfort while being reactive to the presence of pedestrians.
+Past work on reducing jerk and limiting velocities and acceleration does not consider 
+reactive behaviors such as responding to pedestrians. On the other hand, systems for collision 
+avoidance, find and plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
+In this paper, we present an integrated trajectory generation and tracking control system for urban 
+environments that simultaneously considers constraints to maximize human comfort and real-time  
+trajectory updates to react to pedestrians on the road.
 
-Given a desired path, we present a novel method of planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
-This method first segments the path using an arbitrary set of kinematic or dynamic constaints which may be functions of the desired path.
-Each segement is then planned sequentially by choosing a piecewise constant jerk profile.
-The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
-The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is trajectory tracking.
-We then present an online method of revising the trajectory plan to safely avoid pedestrians who enter the roadway by slowing or stopping when no alternative paths are available.
-This framework is shown to be efficient and reliable for use in online planning with a test vehicle and real pedestrians, while optimizing rider comfort.
+Given a desired path, we present a novel online method for planning longitudinal trajectories 
+to follow the urban path with turns and intersections while honoring traffic regulations such 
+as stop signs. Based on sensor data, our method updates the trajectory plan to safely avoid 
+pedestrians on the road by slowing or stopping when no alternative paths are available. 
+Our method has closed form solutions, and runs at 20 Hz ???????, and is efficient and reliable 
+for use in online planning while optimizing rider comfort. We have confirmed this with
+a test vehicle and pedestrians with 10s of hours of testing under autonomous driving conditions.
 
@@
 
-... autonomous vehicles ... \emph{The 1st sentence is the hardest}.
-One such task is computing a tentative reference trajectory plan for travelling from one point to another.
-While the set of possible paths between any two points is infinite, it is typically heavily restricted by consideration of constraints such as traversability and legally available travel lanes, among others.
-The intended speed, acceleration, and higher derivates of position with respect to time are also subject to constraints.
-These may include legal speed limits, engine dynamics, desired sideslip limits, traction availability, and rollover concerns.
-Also of paramount importance is the comfort of the passengers.
+When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
+It is common to employ the path-velocity decomposition to create trajectory plans, meaning there are two tasks which can be carried out sequentially or in parallel.
+% (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
+First is Path Planning, which involves computation of the shortest path.
+The second task is Trajectory Planning, which involves the computation of the fastest velocity profile that is appropriate for the scenario.
 
-In this paper we present an autonomy framework for pedestrian aware ground vehicles.
-We describe the objective and architecture.
-Then we describe other subsystems that our system needs to work with, and planned stops.
-We then describe
-reactivity to pedestrians in the environment by extending the trajectory planner.
-We then summarize and discuss future extensions.
-Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done
-at very low speeds (5 miles/hour).
-There has also been some work in simulating vehicles and modeling
-uncertainty, but pedestrians are less predictable than cars and can appear suddenly on the road 
-(when someone is jaywalking for example).
+Computation of a tentative reference trajectory plan for traveling from one point to another is often guided by
+passenger comfort. While the set of possible paths between any two points is infinite, it is 
+typically heavily restricted by consideration of constraints such as traversability and legally available travel lanes, among others.
+The intended speed, acceleration, and higher derivatives of position with respect to time are also subject to constraints.
+These include legal speed limits, engine dynamics, desired side-slip limits, traction availability, and rollover concerns.
 
+For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort.
+A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
+Some preliminary work on optimizing these constraints on urban roads has been done ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
+On the other hand, many systems in the literature which find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} do not work as part of an integrated system that can navigate urban roads.
+In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort as well as reactive to presence of pedestrians on the road.
+%Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
+%There has also been some work in simulating vehicles and modeling uncertainty, but pedestrians are less predictable than cars 
+%and can appear suddenly on the road (when someone is jaywalking for example).
+
+In this paper, we present an autonomy framework for pedestrian aware ground vehicles.
 Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
-We assume that there is no other traffic on the roadway.
 Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
@@
 It should also react to pedestrians on the road - at intersections as well as on the road.
-If pedestrians are not on the road they should be ignored by the autonomous cars.
+Pedestrians on the sidewalks should be ignored by the autonomous cars.
 Further the pedestrians can be static or moving.
@@
 
-In urban scenarios we need to handle intersections, turns in the road and planned stops due 
+Given a desired path, we present a novel method of planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
+Our method first segments the path using an arbitrary set of kinematic or dynamic constraints which may be functions of the desired path.
+Each segment is then planned sequentially by choosing a piece-wise constant jerk profile.
+The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
+The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking control.
+We then present an online method of revising the trajectory plan to safely avoid pedestrians who enter the roadway by slowing or stopping when no alternative paths are available.
+
+We first describe the architecture of our system in Section II. We also describe other subsystems that our system needs to work with, and planned stops.
+In Section III, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies jerk and maximum velocity and acceleration. 
+We explain how we handle intersections, turns in the road and planned stops due 
 to stop signs (that have already been marked in the map).
-In addition we exhibit reactionary 
-behavior and stop of pedestrians that may appear on the road.
-We handle this via a high 
-level path manager that calls the trajectory planner as necessary as shown in Figure \ref{fig:addreact}.
+We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
+Once a trajectory plan is formulated, it must be utilized online in order for a reference signal to be sent to the controllers which govern steering, braking, and engine speed.
+This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
+Section V describes the high level Path Manager that calls the trajectory planner as necessary to 
+exhibit reactionary behavior and stop for pedestrians that may appear on the road.
+We then conclude and discuss future extensions.
+% tracking
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{System Architecture} \label{sec:systemarchirectureg}
+
+Our system architecture is shown in Figure \ref{fig:addreact}.
+????? describe ???
 
@@
 
-For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort.
-A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
-Some preliminary work on optimizing these constraints on urban roads has been done ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
-On the other hand, many systems in the literature which find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} do not work as part of an integrated system that can navigate urban roads.
-In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort as well as reactive to presence of pedestrians on the road.
-
-% tracking
-Once a trajectory plan is formulated, it must be utilized online in order for a reference signal to be sent to the controllers which govern steering, braking, and engine speed.
-This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Input Subsystems} \label{sec:inputsubsystems}
-
 We assume that we have access to Path Planner. This can be for example a service similar to
@@
   \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/ped2.png}
+  \includegraphics[width=0.5\columnwidth]{graphics/ped3.png}
   \caption{Showing detected pedestrian by the computer vision system
   \newline
-  \textbf{!!!!!!!!Crop to save space!!!!!!!!!!!!!!!}
   }
@@
 Elevation changes will not be addressed here.
-Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
+Integrating piece-wise constant jerk over time yields acceleration which is piece-wise linear over time, speed which is piece-wise quadratic over time, and position which is piece-wise cubic over time.
 We assume that the path planner operates independently from the trajectory planner and tracker.
@@
     7 phase constant jerk profile along curvilinear distance axis.
-    All values needed to completely parameterize a single path segment for trajectory are labelled here.}
+    All values needed to completely parameterize a single path segment for trajectory are labeled here.}
   \label{fig:full7phasespec}
@@
 
-Consider, as a starting point, the case in which a vehicle is moving at some small initial speed and needs to accelerate to some travelling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+Consider, as a starting point, the case in which a vehicle is moving at some small initial speed and needs to accelerate to some traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
 Figure \ref{fig:full7phasespec} shows this trajectory over time.
@@
 The other profiles are explained below in Section \ref{sec:jerkprofiles}.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved indivually in sequence.
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved individually in sequence.
 Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
@@
 There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$.
-All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, travelling speed of zero, negative speeds in $\mathbf{b}$, etc).
-It should be noted that travelling backward (i.e., in reverse gear) is not supported in this work.
+All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
+It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
 To find closed-form solutions for each, a symbolic solver was used.
@@
 The peak speed in this case has a solution, but cannot be manipulated directly.
-Without this known value, the problem becomes underdefined, so $\Delta t_2$ and $\Delta t_5$ each have two solutions: one negative and one positive.
+Without this known value, the problem becomes under-defined, so $\Delta t_2$ and $\Delta t_5$ each have two solutions: one negative and one positive.
 The positive solution is chosen.
 In the case where both solutions are negative, accel-jerk tuning is attempted. 
-Should that yield no positive time intervals, one of the two 4 phase profiles then becomes appropriate, since $L$ is not large enough to accomodate accelerating past the end speed.
+Should that yield no positive time intervals, one of the two 4 phase profiles then becomes appropriate, since $L$ is not large enough to accommodate accelerating past the end speed.
 If $v_i > v_f$, then a reversed 4 phase profile is attempted.
@@
 While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
-In the present example, this means that in the straight portions of the path $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entireity of the curve.
+In the present example, this means that in the straight portions of the path $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
 As a result, many of the curves will have a constant speed.
@@
 While no further information than initial conditions, time durations, and jerk values is needed (simple integration would yield lower derivatives of position), this information is already available since it must be solved for each segment.
-This extra information may be used as a redundancy layer to verify that no errors occured.
+This extra information may be used as a redundancy layer to verify that no errors occurred.
 For each segment with $M$ jerk phases, the following is produced:
@@
 After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
-Since $j(t)$ is piecewise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
+Since $j(t)$ is piece-wise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
 To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
@@
 Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
-Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
+Acceleration $a_{B,x}(t)$ is piece-wise linear, so lookup is a trivial matter.
 The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
@@
 
-\section{Trajectory Tracking} \label{sec:trajectorytracking}
+\section{Trajectory Tracking Control} \label{sec:trajectorytracking}
 
 After planning has been performed, a reference signal must be regularly supplied to a control system in order for the vehicle to carry out the plan.
-This is known as tracking.
+This is known as tracking control.
 The primary difficulty here is that the trajectory plan has been developed in a one dimensional position space using a reference time which will have unknown drift relative to real time.
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+Since it is known that $s(t)$ is piece-wise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
 In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
@@
   \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}}
-  \caption{Reactive behavior to pedestrain}
+  \caption{Reactive behavior to pedestrian}
   \label{fig:react}
@@
 The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
-Within the trajectory tracker, all portions of the path leading up to the location of RSTOP tiggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
+Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
 This is depicted in Fig. \ref{fig:pathtostop}
@@
 
-We have built a robust stopping model as part of Trajectory planner reliably integrated with path manager.
+In this paper, we have presented a novel online method for planning longitudinal trajectories 
+to follow the urban path with real time updates to avoid pedestrians on the roadway by slowing or 
+stopping. We built a robust stopping model as part of Trajectory planner reliably integrated with 
+the Path Manager.
 We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
-they leave. Our system handles corner cases gracefully, and works with multiple pedestrians as we only 
-need to consider the closest pedestrian ahead of us. In future, we want to handle other roadway traffic, 
+they leave. Our system handles corner cases gracefully, following pedestrians if they continue walking 
+on the road away from the car. Our system works with multiple pedestrians as we only 
+need to consider the closest pedestrian ahead of us.
+
+We make three contributions in this paper. First, we present an integrated trajectory planner that
+simultaneously limits jerk, velocity and acceleration while being responsive to presence of pedestrians.
+Second, our trajectory planner has closed form solutions, and is suitable for online implementation
+on a car, unlike systems requiring solutions to non linear optimizations. Finally, we have confirmed the
+efficiency and reliability of our trajectory planner on a test vehicle with 10s of hours of testing
+under autonomous driving conditions on urban roads with pedestrians in random locations.
+
+In future, we would like to handle other roadway traffic, 
 coexisting with other vehicles. We would also like to perceive and stop for red-lights.

### 2016-01-21 root.tex
update first 4 pages

@@
 
-In this paper, we address the problem of trajectory planning for urban roads that satisfies 
-constraints for human comfort while being reactive to the presence of pedestrians.
-Past work on reducing jerk and limiting velocities and acceleration does not consider 
+In this paper, we address the problem of trajectory planning for urban roads while being 
+reactive to the presence of pedestrians.
+Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider 
 reactive behaviors such as responding to pedestrians. On the other hand, systems for collision 
-avoidance, find and plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
+avoidance, plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
 In this paper, we present an integrated trajectory generation and tracking control system for urban 
@@
 as stop signs. Based on sensor data, our method updates the trajectory plan to safely avoid 
-pedestrians on the road by slowing or stopping when no alternative paths are available. 
+pedestrians on the road by slowing or stopping.
 Our method has closed form solutions, and runs at 20 Hz ???????, and is efficient and reliable 
-for use in online planning while optimizing rider comfort. We have confirmed this with
+for use in online planning. We have confirmed this with
 a test vehicle and pedestrians with 10s of hours of testing under autonomous driving conditions.
@@
 When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
-It is common to employ the path-velocity decomposition to create trajectory plans, meaning there are two tasks which can be carried out sequentially or in parallel.
+It is common to employ the path-velocity decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
@@
 
-Computation of a tentative reference trajectory plan for traveling from one point to another is often guided by
-passenger comfort. While the set of possible paths between any two points is infinite, it is 
-typically heavily restricted by consideration of constraints such as traversability and legally available travel lanes, among others.
+Computation of a trajectory plan for traveling from one point to another is often guided by
+passenger comfort. The set of possible paths to the destination is typically restricted 
+by constraints such as traversability and legally available travel lanes, among others.
 The intended speed, acceleration, and higher derivatives of position with respect to time are also subject to constraints.
@@
 Some preliminary work on optimizing these constraints on urban roads has been done ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
-On the other hand, many systems in the literature which find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} do not work as part of an integrated system that can navigate urban roads.
-In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort as well as reactive to presence of pedestrians on the road.
+On the other hand, many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
+In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort and is reactive to presence of pedestrians on the road.
 %Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
@@
 Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
-Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
+%Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
 At stop signs the car should stop and wait for an appropriate amount of time.
@@
 
-Given a desired path, we present a novel method of planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
+Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
 Our method first segments the path using an arbitrary set of kinematic or dynamic constraints which may be functions of the desired path.
-Each segment is then planned sequentially by choosing a piece-wise constant jerk profile.
+Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
 The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
 The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking control.
-We then present an online method of revising the trajectory plan to safely avoid pedestrians who enter the roadway by slowing or stopping when no alternative paths are available.
+We present an online method of revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing or stopping when no alternative paths are available.
 
 We first describe the architecture of our system in Section II. We also describe other subsystems that our system needs to work with, and planned stops.
-In Section III, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies jerk and maximum velocity and acceleration. 
-We explain how we handle intersections, turns in the road and planned stops due 
-to stop signs (that have already been marked in the map).
+In Section III, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies jerk and maximum velocity and acceleration constraints. 
+We explain how we handle intersections, turns in the road and planned stops due to stop signs (that have already been marked in the map).
 We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
-Once a trajectory plan is formulated, it must be utilized online in order for a reference signal to be sent to the controllers which govern steering, braking, and engine speed.
+Once a trajectory plan is formulated, it must be utilized online to send the reference signal to the controllers that govern steering, braking, and engine speed.
 This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
@@
 We assume that we have access to Path Planner. This can be for example a service similar to
-Google Maps except that it provides lane level maps. We also assume that we have a High
+Google Maps except that it provides lane level maps. 
+We assume that the path planner operates independently from the trajectory planner and tracker.
+We also assume that we have a High
 Resolution digital map with Lane centers, Stop signs, crosswalk geometry, and Turning 
@@
   \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/IntersectionCrosswalks.png}
+  \includegraphics[width=0.3\columnwidth]{graphics/IntersectionCrosswalks.png}
   \caption{A car turning right at the intersection only needs to care about pedestrians in the 
@@
 It outputs Global position, global heading,
-body frame velocity and body frame acceleration to 10s of cms.
+body frame velocity and body frame acceleration to cm. level accuracy.
 
@@
 
-For the purposes of this work, the objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
+The objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
 Elevation changes will not be addressed here.
-Integrating piece-wise constant jerk over time yields acceleration which is piece-wise linear over time, speed which is piece-wise quadratic over time, and position which is piece-wise cubic over time.
-We assume that the path planner operates independently from the trajectory planner and tracker.
+Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
+
 Prescribed initial and end conditions for position, speed, and acceleration must also be honored.
@@
 The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
-It is assume that the final forward acceleration will be zero.
-If the magnitude of the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then the profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+It is assumed that the final forward acceleration will be zero.
+If the magnitude of the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then a profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
 Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
 The remaining parameters yet to be set now become $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
-As we will show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
+As we show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
 To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs.
@@
 This is computed for every point along the path.
-In the simplest case, one value for each of the quantities expressed in those relations may be chosen for the entire path.
+In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.
 
@@
 
-It is obvious that there are many cases in which this solution will fail.
+There are many cases in which this solution will fail.
 For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
-We propose that 4 other basic profiles be defined, and an algorithm constructed which judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
+We propose to define 4 other basic profiles, and our algorithm judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
 The other profiles are explained below in Section \ref{sec:jerkprofiles}.
@@
 
-This profile, described above, is always attempted first.
+This profile, with $\Delta t_i$ as shown in Figure {fig:full7phasespec}, is always attempted first.
 All time steps have a single solution, enabled by the fact that the initial and end conditions are known, and the two peak accelerations and two jerk magnitudes are provided as configuration parameters.
@@
   \centering
-  \includegraphics[width=0.7\columnwidth]{graphics/6phase_v(s).png}
-  \caption{6 phase profile, displayed as speed vs. position along path}
+  \begin{tabular}{cc}
+  \subfigure[6 phase profile] {
+  \includegraphics[width=0.4\columnwidth]{graphics/6phase_v(s).png}
   \label{fig:6phaseprofile}
+  } &
+  \subfigure[4 phase profile] {
+  \includegraphics[width=0.4\columnwidth]{graphics/4phase_v(s).png}
+  \label{fig:4phaseprofile}
+  } \\
+  \subfigure[4 Reversed phase profile] {
+  \includegraphics[width=0.4\columnwidth]{graphics/4Rphase_all_derivatives.png}
+  \label{fig:4rphaseprofile}
+  } &
+  \subfigure[3 phase profile] {
+  \includegraphics[width=0.4\columnwidth]{graphics/3phase_v(s).png}
+  \label{fig:3phaseprofile}
+  }
+  \end{tabular}
+  \caption{6, 4, 4Reversed and 3 phase profiles, displayed as speed vs. position along path}
+
 \end{figure}
@@
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.7\columnwidth]{graphics/4phase_v(s).png}
-  \caption{4 phase profile, displayed as speed vs. position along path within the segment.}
-  \label{fig:4phaseprofile}
-\end{figure}
-
 \subsubsection{Reversed 4 Phase} \label{sec:reversed4phase}
@@
 The initial speed is maintained for a period before applying the brakes.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.7\columnwidth]{graphics/4Rphase_all_derivatives.png}
-  \caption{Reversed 4 phase profile, displayed as speed vs. position along path within the segment.}
-  \label{fig:4rphaseprofile}
-\end{figure}
 
@@
 As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4 phase profiles has been attempted and subsequently failed.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.7\columnwidth]{graphics/3phase_v(s).png}
-  \caption{3 phase profile, displayed as speed vs. position along path.}
-  \label{fig:3phaseprofile}
-\end{figure}
 
@@
 % How we actually did this.
-For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ($SL$, set constant for the whole path) and lateral acceleration limit ($a_{B,y}^{max}$.
+For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ($SL$, set constant for the whole path) and lateral acceleration limit $a_{B,y}^{max}$.
 Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
+The value $\kappa_k$ denotes the path curvature at each point.
 For each path point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
-The value $\kappa_k$ denotes the path curvature at each point.
 
@@
 
-The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...}\tabularnewline
+The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...???????????}\tabularnewline
 
@@
 For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
-To set boundary speed conditions for the interior segments, Algorithm~\ref{alg:segmentspeedboundaryconditions}
+Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments.
 
@@
 After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
-Since $j(t)$ is piece-wise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
+Since $j(t)$ is piecewise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
 To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
@@
 Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
-Acceleration $a_{B,x}(t)$ is piece-wise linear, so lookup is a trivial matter.
+Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
 The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piece-wise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+Since it is known that $s(t)$ is piecewise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
 In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.

### 2016-01-22 root.tex
update pages 5-8

@@
 Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
+\emph{RG: $N_p$ not defined. this is the first place we introduce two subscripts. please explain what the two subscripts stand for in general for example in $a_{B,y}$ and $v_{m,k}$ ?????? }
 The value $\kappa_k$ denotes the path curvature at each point.
@@
   \end{algorithmic}
-\caption{Algorithm to set speed boundary conditions for interior path segments}
+\caption{Algorithm to set speed boundary conditions for interior path segments. We iterate over ??? in line 5 to set ?????? }
 \label{alg:segmentspeedboundaryconditions}
 \end{figure}
+
+\emph{RG: lines 6 and 8 in algorithm fire ate same. please check.??????}
 
@@
 
+\emph{RG: please explain reference time, $s$ and $\Delta s$ ????????}
+
 After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
@@
 However, the time at which the trajectory plan must be referenced is not always known.
-The solution to this problem is known as trajectory tracking, and will be discussed in the next section.
+The solution to this problem is known as trajectory tracking control, and will be discussed in the next section.
 
@@
 The primary difficulty here is that the trajectory plan has been developed in a one dimensional position space using a reference time which will have unknown drift relative to real time.
-As such, we must find some way to look up the coordinates of the vehicle within the reference trajectory using information typically available for autonomous cars.
+As such, we must find a way to look up the coordinates of the vehicle within the reference trajectory using information typically available for autonomous cars.
 A minimal set of such information would include position, velocity vector, heading, and acceleration in the body frame.
-After this, it is assumed that the vehicle controller will need samples of a small window into the future trajectory in order to pilot the vehicle.
+After this, it is assumed that the vehicle controller will need samples of a small window into the future trajectory to pilot the vehicle.
 
@@
 Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity, and acceleration vectors) onto the path.
-A simple approximation, which works well for the path shown in Figure~\ref{fig:1Dto2DandSegmentation} is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'').
+A simple approximation, which works well for the path shown in Figure~\ref{fig:1Dto2DandSegmentation} ????????
+is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'').
+\emph{ RG: please show link in figure ???????}
 Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
 The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection $\mathbf{r}_p$.
-Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex paths which overlap or self-intersect.
+Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex paths which overlap or self-intersections.
 
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, $t(s)$ may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+Since it is known that $s(t)$ is piecewise cubic, $t(s)$ \emph{RG: please define $t(s)$ ?????????} may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
 In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
@@
 \end{figure}
+\emph{RG: a figure text etc is smaller than b and c. maybe crop from right to make same size}
 
@@
 
-Next the nominal stopping distance $Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
+Next the nominal stopping distance $\Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
 The procedure for doing this will be the same as the procedure for planning the RSTOP later with a few steps omitted, since the planning algorithm is extremely efficient.
@@
 While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
-As was mentioned in previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
-Normal planning chooses enforces $L$ to avoid skewing segment boundaries within each subpath.
-However, reactive stopping requires that $v_f$ be enforced, and $L$ be a parameter which may be solved dependently.
+As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
+Normal planning enforces $L$ to avoid skewing segment boundaries within each subpath.
+However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
 Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
@@
   \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
-  \caption{Green color path is the one over which the car finds a trajectory to stop}
+  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
   \label{fig:pathtostop}

### 2016-01-22 root.tex
fixing up figures. added ped detection citation. also cleaning up commented lines

@@
 pedestrians on the road by slowing or stopping.
-Our method has closed form solutions, and runs at 20 Hz ???????, and is efficient and reliable 
+Our method has closed form solutions, and runs at 20 Hz, and is efficient and reliable 
 for use in online planning. We have confirmed this with
@@
 Our system architecture is shown in Figure \ref{fig:addreact}.
-????? describe ???
+There are 4 primary input components.
+A path planner determines the optimal course to take by searching a map database.
+This same map database is used to search for stop signs and provide their locations.
+A GPS/INS navigation filter provides estimates of global position, velocity vector, attitude, and acceleration.
+A pedestrian detection component \cite{Gepperth2013,Gepperth2014} leverages computer vision in conjunction with the navigation filter to estimate the global location of pedestrians which are walking in front of the vehicle.
+Techniques for designing these components are abundant in the literature and are addressed here.
 
@@
   \caption{
-    Adding reactive behavior to trajectory planning.
-    \newline
-    \textbf{!!!!!!!!!!!!!!!!!!!!!!REDO THIS!!!!!!!!!!!!!!!!!!}
+   Primary system components. The ``Path Manager'' and ``Trajectory Planner'' scomponents (in blue) are presented herein.
   }
@@
 
-We assume that we have access to Path Planner. This can be for example a service similar to
+We assume that we have access to Path Planner. For example, this can be a service similar to
 Google Maps except that it provides lane level maps. 
@@
 For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ($SL$, set constant for the whole path) and lateral acceleration limit $a_{B,y}^{max}$.
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$.
-\emph{RG: $N_p$ not defined. this is the first place we introduce two subscripts. please explain what the two subscripts stand for in general for example in $a_{B,y}$ and $v_{m,k}$ ?????? }
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$, where $N_p$ is the number of waypoints in the full path.
+\emph{RG: this is the first place we introduce two subscripts. please explain what the two subscripts stand for in general for example in $a_{B,y}$ and $v_{m,k}$ ?????? }
 The value $\kappa_k$ denotes the path curvature at each point.
@@
 
-The power of this algorithm is that any set and number of constraints may be used ... \emph{etc., etc., ...???????????}\tabularnewline
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
       \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
-      \For{$k = 1, ..., N_s-1$}
+      \For{$i = 1, ..., N_s-1$}
         \If{$v_{m,i+1} > v_{m,i}$}
-          \State $v_{p,i+1} \gets v_{m,i}$
-        \ElsIf{$v_{m,i+1} > v_{m,i}$}
           \State $v_{p,i+1} \gets v_{m,i+1}$
@@
   \end{algorithmic}
-\caption{Algorithm to set speed boundary conditions for interior path segments. We iterate over ??? in line 5 to set ?????? }
+  \caption{
+    Algorithm to set speed boundary conditions for interior path segments.
+    Having consolidated the waypoints in to a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
+    This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
+  }
 \label{alg:segmentspeedboundaryconditions}
 \end{figure}
-
-\emph{RG: lines 6 and 8 in algorithm fire ate same. please check.??????}
 
@@
 To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
-Position along the path $s(t)$ takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
+Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
 Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
@@
 Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity, and acceleration vectors) onto the path.
-A simple approximation, which works well for the path shown in Figure~\ref{fig:1Dto2DandSegmentation} ????????
-is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'').
-\emph{ RG: please show link in figure ???????}
+A simple approximation is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'', and is depicted in Fig.
+\ref{fig:cartos}).
 Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, $t(s)$ \emph{RG: please define $t(s)$ ?????????} may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$,F may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
 In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
@@
 
-\textbf{3 phase full spec profile figure?}.
-
 Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.

### 2016-01-22 root.tex
fixing RSTOP diagrams
finished updating RSTOP/PSTOP pics

@@
     The trajectory generation is serially called on each segment colored differently.
-    \textbf{Need to update figure with stop signs???? ??????? ?????}
   }
@@
   \\
-  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RSTOP.png}}
+  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Replan.png}}
   \\
-  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_RESUME.png}}
+  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
   \caption{Reactive behavior to pedestrian}

### 2016-01-22 root.tex
cleaned up FIXME sentence in paper
changed assosciation details
rearranging begun

@@
   % Reactive Path Planning with Pedestrains for Autonomous Driving
-  Reactive Trajectory Planning and Tracking Control for Pedestrian-Aware Autonomous Driving in Urban Environments
+  Reactive Trajectory Planning and Trajectory Tracking for Pedestrian-Aware Autonomous Driving in Urban Environments
 }
@@
   \thanks{
-    $^{1}$ Auburn University, Dept. of Mechanical Engineering, GPS \& Vehicle Dynamics Laboratory.
+    $^{1}$ Auburn University GAVLab. 1418 Wiggins Hall, Auburn, AL 36832. robertgcofield@gmail.com
   }
@@
 
-In this paper, we address the problem of trajectory planning for urban roads while being 
-reactive to the presence of pedestrians.
-Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider 
-reactive behaviors such as responding to pedestrians. On the other hand, systems for collision 
-avoidance, plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
-In this paper, we present an integrated trajectory generation and tracking control system for urban 
-environments that simultaneously considers constraints to maximize human comfort and real-time  
-trajectory updates to react to pedestrians on the road.
+In this paper, we address the problem of trajectory planning for urban roads while reacting to the presence of pedestrians.
+Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider reactive behaviors such as responding to pedestrians.
+On the other hand, systems for collision avoidance plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
+In this paper, we present an integrated trajectory generation and tracking system for urban environments that simultaneously considers constraints to maximize human comfort and real-time trajectory updates to react to pedestrians on the road.
 
-Given a desired path, we present a novel online method for planning longitudinal trajectories 
-to follow the urban path with turns and intersections while honoring traffic regulations such 
-as stop signs. Based on sensor data, our method updates the trajectory plan to safely avoid 
-pedestrians on the road by slowing or stopping.
-Our method has closed form solutions, and runs at 20 Hz, and is efficient and reliable 
-for use in online planning. We have confirmed this with
-a test vehicle and pedestrians with 10s of hours of testing under autonomous driving conditions.
+Given a desired path, we present a novel online method for planning longitudinal trajectories to follow the urban path with turns and intersections while honoring traffic regulations such as stop signs.
+Based on sensor data, our method updates the trajectory plan to safely avoid pedestrians on the road by slowing or stopping.
+This is critical in urban environments, when modifying the vehicle's intended path is often not possible.
+Our method has closed form solutions, runs at 20 Hz, and is efficient and reliable for use in online planning.
+We have confirmed this with a test vehicle and pedestrians with over 100 hours of testing under driverless operation.
 
@@
 
-We first describe the architecture of our system in Section II. We also describe other subsystems that our system needs to work with, and planned stops.
-In Section III, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies jerk and maximum velocity and acceleration constraints. 
+We first describe the architecture of our system in Sec. \ref{sec:systemarchirecture}.
+We also describe other subsystems that our system needs to work with, and planned stops.
+In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies jerk and maximum velocity and acceleration constraints. 
 We explain how we handle intersections, turns in the road and planned stops due to stop signs (that have already been marked in the map).
@@
 
-\section{System Architecture} \label{sec:systemarchirectureg}
+\section{System Architecture} \label{sec:systemarchirecture}
 
 Our system architecture is shown in Figure \ref{fig:addreact}.
-There are 4 primary input components.
-A path planner determines the optimal course to take by searching a map database.
-This same map database is used to search for stop signs and provide their locations.
-A GPS/INS navigation filter provides estimates of global position, velocity vector, attitude, and acceleration.
-A pedestrian detection component \cite{Gepperth2013,Gepperth2014} leverages computer vision in conjunction with the navigation filter to estimate the global location of pedestrians which are walking in front of the vehicle.
-Techniques for designing these components are abundant in the literature and are addressed here.
+There are 4 primary input components: a path planner, navigation filter, pedestrian detector, and a map which provides the location of stop signs.
+% nav filter
+For navigation filtering we use an ADMA-G  commercial automotive GPS/INS system.
+It receives RTK corrections via cell modem.
+It outputs global position, global heading, body frame velocity and body frame acceleration to centimeter level accuracy.
 
@@
 
-We assume that we have access to Path Planner. For example, this can be a service similar to
-Google Maps except that it provides lane level maps. 
-We assume that the path planner operates independently from the trajectory planner and tracker.
-We also assume that we have a High
-Resolution digital map with Lane centers, Stop signs, crosswalk geometry, and Turning 
-curves at intersections. The starting position the 
-car is determined via high grade GPS. The destination is manually input into the system.
-Using the path planner, we obtain a path along lane centers from start to the destination.
-This plan is output as waypoints, with the waypoints corresponding to stop signs marked.
+Pedestrian detection is performed via a camera and LiDAR system based on \cite{Gepperth2013,Gepperth2014}.
+It computes a bounding box for each pedestrian as shown in Figure \ref{fig:ped}.
+The pedestrian position is transformed to the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
 
@@
 
+We assume that we have access to a path planner.
+For example, this can be a service similar to Google Maps except that it provides lane level maps. 
+We assume that the path planner operates independently from the trajectory planner and tracker.
+We also assume that we have a high resolution digital map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
+The destination is manually input into the system by the passengers.
+Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the destination.
+This plan is output as set of global waypoints, with the waypoints corresponding to stop signs flagged.
+This same map database is used to search for stop signs and provide their locations.
+
 %\begin{figure}[thpb]
@@
 
-Pedestrian Detection is performed via a Camera, and LiDAR system. It computes a bounding 
-box for each pedestrian as shown in Figure \ref{fig:ped}. The pedestrian position is 
-transformed to the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.3\columnwidth]{graphics/IntersectionCrosswalks.png}
+  \caption{
+    A car turning right at the intersection only needs to consider pedestrians in the pedestrian crossings shown yellow.
+    Pedestrians in the green crossings can be ignored unless they appear somewhere on the road.
+  }
+  \label{fig:intersect}
+\end{figure}
+
+We also have an In-Out Algorithm implemented via polygon intersection that can be used to ignore “safe” pedestrians
+Pedestrians that are on sidewalks, off-path crosswalks, and behind planned stops as shown in Figure \ref{fig:intersect} can be considered ``safe'' and are ignored.
+This is done by searching the same map database used for path planning.
+Remaining pedestrians pose collision risk and the vehicle needs to reactively slow down or stop for them.
+This is accomplished by sending stop message with pedestrian locations to the path manager.
+
+Beyond the input subsystems, there are three other major components: a high-level planner, a low-level planner, and a software controller which governs engine speed, braking, and steering.
+The controller takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
+This short-term trajectory is oriented in the vehicle body frame.
+\textbf{Citation for controller !!!!!}.
+The function of the low-level planner (``trajectory planner'') is two-fold:
+\begin{itemize}
+  \item Compute medium-range trajectories, as detailed in Sec. \ref{sec:trajectoryplanning}.
+  \item Sample medium-range trajectories for input to the controller, as detailed in Sec. \ref{sec:trajectoryplanning}
+\end{itemize}
+
+This work focuses on the high- and low-level planners.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Path Manager} \label{sec:pathmanager}
+
+The assumption on our work is that there is a unique mapping from a reference time $t_{ref}$ to a scalar value indicating the distance along the current path from the starting, $s$.
+system.
+This mapping breaks down when the vehicle is stopped (planned or unplanned), as there is no way to look up $t_{ref}(s)$ when multiple reference time solutions exist for certain path positions.
+The solution is to break the path into sub-paths at each stop location (i.e., stop sign) which is known a priori, and plan then execute them incrementally as shown in Fig. \ref{fig:segmentation}.
+This is the first function of the path manager.
+% , and can be broken down as follows:
+% \begin{itemize}
+% \item Send first sub-path to trajectory planner
+% \item Wait until completed.
+% \item Pause for fixed time, typically 2-3 second.
+% \item Continue waiting if any pedestrians are in the roadway ahead (along the intended path) within some fixed distance.
+% \item Send next sub-path to planner.
+% \item Repeat until end of path is reached.
+% \end{itemize}
 
@@
   \centering
-  \includegraphics[width=0.3\columnwidth]{graphics/IntersectionCrosswalks.png}
-  \caption{A car turning right at the intersection only needs to care about pedestrians in the 
-pedestrian crossings shown yellow. Pedestrians in the green crossings can be ignored unless they
-appear somewhere on the road}
-  \label{fig:intersect}
+  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
+  \caption{
+    Segmentation of the path by path manager.
+    The trajectory generation is serially called on each segment colored differently.
+  }
+  \label{fig:segmentation}
 \end{figure}
 
-We also have an In-Out Algorithm implemented via polygon intersection that can be used to
-ignore “safe” pedestrians that are on sidewalks, off-path crosswalks, and behind planned stop
-as shown in Figure \ref{fig:intersect}.
-Remaining pedestrians pose collision risk and we need to reactively slow down or stop for them.
-This is accomplished by sending stop message with pedestrian locations to path manager.
+Both behaviors dictate that the path manager govern the operation of the trajectory planner.
+To enable both of these features, the path manager is implemented as a discrete state machine, as shown in Fig. \ref{fig:st}.
 
-For Navigation Filtering we use an ADMA-G based Commercial automotive GPS/INS system.
-It receives RTK corrections via cell modem.
-It outputs Global position, global heading,
-body frame velocity and body frame acceleration to cm. level accuracy.
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
+  \caption{State Transition Diagram}
+  \label{fig:st}
+\end{figure}
+
+The distinction between momentarily braking for a pedestrian which quickly leaves the roadway and coming to a full stop needs no formal logic.
+They follow the same state transition, and since both trajectory plans are extremely smooth, riders are rarely aware that any change took place.
+It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.
+
+Next, the triggers for state transition will be described.
+They are implemented as a set of static and dynamic thresholds, which are depicted graphically in Fig. \ref{fig:react}.
+
+\begin{figure}[thpb]
+\centering
+  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
+  \\
+  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Replan.png}}
+  \\
+  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
+  \caption{Reactive behavior to pedestrian}
+  \label{fig:react}
+\end{figure}
+
+All detected pedestrians locations are compared against the high-resolution digital map, so that pedestrians which are not on the roadway or near an upcoming crosswalk are not considered.
+Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
+Pedestrians which are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
+All the pedestrians which remain pose a potential collision risk, however only the closest is considered for RSTOP.
+
+Next the nominal stopping distance $\Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
+The procedure for doing this will be the same as the procedure for planning the RSTOP later with a few steps omitted, since the planning algorithm is extremely efficient.
+A 3 phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+
+While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
+As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
+Normal planning enforces $L$ to avoid skewing segment boundaries within each subpath.
+However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
+Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
+%%% This is the brief way of wording the preceding three sentences.
+% The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met.
+So we reformulate and replace $L$ with $v_f$ as the hard constraint.
+
+Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.
+For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used.
+Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
+RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= s_{ped} - s_{veh}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
+The trajectory is then recomputed with increasingly higher-magnitude values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
+Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
+
+After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
+The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
+Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
+This is depicted in Fig. \ref{fig:pathtostop}
+The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
+  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
+  \label{fig:pathtostop}
+\end{figure}
+
+This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
+This value is typically set to 1 meter.
+Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$.
+The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
+The preceding buffers are depicted graphically in Fig. \ref{fig:react}.
 
@@
 \end{enumerate}
-After this is done, the process described above in Section \ref{sec:jerkprofiles} may be applied to each of the segments.
+After this is done, the process described in Section \ref{sec:jerkprofiles} may be applied to each of the segments.
 
@@
 
-\section{Trajectory Tracking Control} \label{sec:trajectorytracking}
+\section{Trajectory Tracking} \label{sec:trajectorytracking}
 
@@
 
-\section{Path Manager} \label{sec:pathmanager}
-
-The assumption on our work is that there is a unique mapping from a reference time $t_{ref}$ to a scalar value indicating the distance along the current path, $s$.
-system.
-This mapping breaks down when we are stopped at a planned or reactive stop, as there is no way to do a $t_{ref}(s)$ lookup when multiple reference time solutions exist for certain path positions.
-The solution is to break the path into sub-paths at each stop location (i.e., stop sign) which is known a priori, and plan then execute them incrementally as shown in Fig. \ref{fig:segmentation}.
-This is the first function of the path manager, and can be broken down as follows:
-\begin{itemize}
-\item Send first sub-path to trajectory planner
-\item Wait until completed.
-\item Pause for fixed time, typically 2-3 second.
-\item Continue waiting if any pedestrians are in the roadway ahead (along the intended path) within some fixed distance.
-\item Send next sub-path to planner.
-\item Repeat until end of path is reached.
-\end{itemize}
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
-  \caption{
-    Segmentation of the path by path manager.
-    The trajectory generation is serially called on each segment colored differently.
-  }
-  \label{fig:segmentation}
-\end{figure}
-
-Both behaviors dictate that the path manager govern the operation of the trajectory planner.
-To enable both of these features, the path manager is implemented as a discrete state machine, as shown in Fig. \ref{fig:st}.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
-  \caption{State Transition Diagram}
-  \label{fig:st}
-\end{figure}
-
-The distinction between momentarily braking for a pedestrian which quickly leaves the roadway and coming to a full stop needs no formal logic.
-They follow the same state transition, and since both trajectory plans are extremely smooth, riders are rarely aware that any change took place.
-It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.
-
-Next, the triggers for state transition will be described.
-They are implemented as a set of static and dynamic thresholds, which are depicted graphically in Fig. \ref{fig:react}.
-
-\begin{figure}[thpb]
-\centering
-  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
-  \\
-  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Replan.png}}
-  \\
-  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
-  \caption{Reactive behavior to pedestrian}
-  \label{fig:react}
-\end{figure}
-\emph{RG: a figure text etc is smaller than b and c. maybe crop from right to make same size}
-
-All detected pedestrians locations are compared against the high-resolution digital map, so that pedestrians which are not on the roadway or near an upcoming crosswalk are not considered.
-Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
-Pedestrians which are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
-All the pedestrians which remain pose a potential collision risk, however only the closest is considered for RSTOP.
-
-Next the nominal stopping distance $\Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
-The procedure for doing this will be the same as the procedure for planning the RSTOP later with a few steps omitted, since the planning algorithm is extremely efficient.
-A 3 phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
-
-While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
-As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
-Normal planning enforces $L$ to avoid skewing segment boundaries within each subpath.
-However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
-Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
-%%% This is the brief way of wording the preceding three sentences.
-% The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met.
-So we reformulate and replace $L$ with $v_f$ as the hard constraint.
-
-Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.
-For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used.
-Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
-RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= s_{ped} - s_{veh}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
-The trajectory is then recomputed with increasingly higher-magnitude values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
-Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
-
-After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
-The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
-Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
-This is depicted in Fig. \ref{fig:pathtostop}
-The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
-  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
-  \label{fig:pathtostop}
-\end{figure}
-
-This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
-This value is typically set to 1 meter.
-Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$.
-The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
-The preceding buffers are depicted graphically in Fig. \ref{fig:react}.
-
 \section{Summary} \label{sec:summary}

### 2016-01-22 root.tex
working on path manager section

@@
   \thanks{
-    $^{1}$ Auburn University GAVLab. 1418 Wiggins Hall, Auburn, AL 36832. robertgcofield@gmail.com
+    $^{1}$ Auburn University GAVLab, 1418 Wiggins Hall, Auburn, AL 36832. \textit{robertgcofield@gmail.com}
   }
   \thanks{
-    $^{2}$ Honda Research Institute USA Inc, 375 Ravendale Drive, Mountain View, CA 94043. rgupta@hra.com
+    $^{2}$ Honda Research Institute USA Inc, 375 Ravendale Drive, Mountain View, CA 94043. \textit{rgupta@hra.com}
   }
@@
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-First is Path Planning, which involves computation of the shortest path.
+First is Path Planning, which involves computation of the optimal path.
 The second task is Trajectory Planning, which involves the computation of the fastest velocity profile that is appropriate for the scenario.
@@
 This short-term trajectory is oriented in the vehicle body frame.
-\textbf{Citation for controller !!!!!}.
-The function of the low-level planner (``trajectory planner'') is two-fold:
+\textbf{Citation for controller ????}.
+
+The high-level planner is referred to as the ``path manager'', and performs decision-making as well as governing operation of the low-level planner.
+Its primary functions are:
+\begin{enumerate}
+  \item Stop the vehicle at stop signs and wait until the assosciated intersection is clear before continuing.
+  \item Reactively reduce speed or stop altogether for pedestrians which pose a collision risk, then continue when they are clear.
+  \item Govern the trajectory planner.
+\end{enumerate}
+The low-level planner is referred to as the ``trajectory planner'', and is responsible for quantifying the actions dictated by the path manager.
+Its primary functions are:
 \begin{itemize}
-  \item Compute medium-range trajectories, as detailed in Sec. \ref{sec:trajectoryplanning}.
-  \item Sample medium-range trajectories for input to the controller, as detailed in Sec. \ref{sec:trajectoryplanning}
+  \item Compute medium-range trajectories for a given path.
+  \item Revise trajectories to react to pedestrians.
+  \item Sample the current trajectory for input to the controller.
 \end{itemize}
-
-This work focuses on the high- and low-level planners.
+This work focuses on the path manager (Sec. \ref{sec:pathmanager}) and the trajectory planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
 
@@
 
-The assumption on our work is that there is a unique mapping from a reference time $t_{ref}$ to a scalar value indicating the distance along the current path from the starting, $s$.
-system.
-This mapping breaks down when the vehicle is stopped (planned or unplanned), as there is no way to look up $t_{ref}(s)$ when multiple reference time solutions exist for certain path positions.
-The solution is to break the path into sub-paths at each stop location (i.e., stop sign) which is known a priori, and plan then execute them incrementally as shown in Fig. \ref{fig:segmentation}.
-This is the first function of the path manager.
-% , and can be broken down as follows:
-% \begin{itemize}
-% \item Send first sub-path to trajectory planner
-% \item Wait until completed.
-% \item Pause for fixed time, typically 2-3 second.
-% \item Continue waiting if any pedestrians are in the roadway ahead (along the intended path) within some fixed distance.
-% \item Send next sub-path to planner.
-% \item Repeat until end of path is reached.
-% \end{itemize}
+The path manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
+Each of the states is described below.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
+  \caption{State Transition Diagram}
+  \label{fig:st}
+\end{figure}
+
+Many systems perform trajectory planning over a sliding window within the full path which progresses with the vehicle.
+This requires ... \textbf{something about computing actions through stop signs before it's necessary and how that's inefficient}.
+We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:segmentation}, and  execute them incrementally.
+
+In the absence of pedestrians, the path manager cycles between PSTOP and NORMAL states.
+In NORMAL state, the path manager send the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
+In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
+When pedestrians are present, the transition from PSTOP to NORMAL is delayed until there are no pedestrians detected in the range denoted by $\Delta s_{resume}$.
 
@@
 
-Both behaviors dictate that the path manager govern the operation of the trajectory planner.
-To enable both of these features, the path manager is implemented as a discrete state machine, as shown in Fig. \ref{fig:st}.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
-  \caption{State Transition Diagram}
-  \label{fig:st}
-\end{figure}
+All detected pedestrians locations are compared against the high-resolution digital map, so that pedestrians which are not on the roadway or near an upcoming crosswalk are not considered.
+Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
+Pedestrians which are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
+All the pedestrians which remain pose a potential collision risk, however only the closest is considered for RSTOP.
 
 The distinction between momentarily braking for a pedestrian which quickly leaves the roadway and coming to a full stop needs no formal logic.
-They follow the same state transition, and since both trajectory plans are extremely smooth, riders are rarely aware that any change took place.
+They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
 It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.
-
-Next, the triggers for state transition will be described.
-They are implemented as a set of static and dynamic thresholds, which are depicted graphically in Fig. \ref{fig:react}.
 
@@
   \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
-  \caption{Reactive behavior to pedestrian}
+  \caption{Graphical representation of static and dynamic thresholds used to trigger state transitions.}
   \label{fig:react}
 \end{figure}
-
-All detected pedestrians locations are compared against the high-resolution digital map, so that pedestrians which are not on the roadway or near an upcoming crosswalk are not considered.
-Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
-Pedestrians which are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
-All the pedestrians which remain pose a potential collision risk, however only the closest is considered for RSTOP.
 

### 2016-01-22 root.tex
moved RSTOP trajectory planning stuff to another section

@@
 
-Next the nominal stopping distance $\Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
-The procedure for doing this will be the same as the procedure for planning the RSTOP later with a few steps omitted, since the planning algorithm is extremely efficient.
-A 3 phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+Next the nominal distance needed to stop $\Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
+The procedure for doing this will be the same as the procedure for planning the RSTOP trajectory later with a few steps omitted, since the planning algorithm is extremely efficient.
 
-While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
-As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
-Normal planning enforces $L$ to avoid skewing segment boundaries within each subpath.
-However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
-Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
-%%% This is the brief way of wording the preceding three sentences.
-% The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met.
-So we reformulate and replace $L$ with $v_f$ as the hard constraint.
-
-Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.
-For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used.
 Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
 RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= s_{ped} - s_{veh}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
-The trajectory is then recomputed with increasingly higher-magnitude values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
-Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
-
-After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
-The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
-Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
-This is depicted in Fig. \ref{fig:pathtostop}
-The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
-  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
-  \label{fig:pathtostop}
-\end{figure}
+Planning and executing the trajectory for coming to a halt is detailed in Sec. \ref{sec:reactivestoptrajectory}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Reactive Stop Trajectories} \label{sec:reactivestoptrajectory}
+
+This is the procedure for determining nominal stopping distance required.
+A 3 phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+
+While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
+As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
+Normal planning enforces $L$ to avoid skewing segment boundaries within each subpath.
+However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
+Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
+%%% This is the brief way of wording the preceding three sentences.
+% The trajectory planner uses total distance as parameter, and is not suitable for our purpose as the end speed constraints are not met.
+So we reformulate and replace $L$ with $v_f$ as the hard constraint.
+
+Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.
+For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used.
+
+Once the path manager triggers RSTOP, the trajectory is then recomputed with increasingly higher-magnitude values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
+Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
+
+After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
+The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
+Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
+This is depicted in Fig. \ref{fig:pathtostop}
+The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
+  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
+  \label{fig:pathtostop}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-01-22 root.tex
updates

@@
 The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
-The preceding buffers are depicted graphically in Fig. \ref{fig:react}.
+The preceding quantities are depicted graphically in Fig. \ref{fig:react}.
 
@@
 
-The objective of trajectory planning is to take a 2 dimensional (horizontal plane) path, and compute a series of constant jerk intervals for motion along the length of the path which is time optimal yet still falls within constraints on speed, acceleration, and jerk.
+The task trajectory planning in this context is to take a 2 dimensional (horizontal plane) set of waypoints from the path manager, and compute a series of constant jerk intervals for motion along the length of the path.
+This trajectory must be time optimal yet still fall within constraints on speed, acceleration, and jerk.
 Elevation changes will not be addressed here.
@@
 Prescribed initial and end conditions for position, speed, and acceleration must also be honored.
-This means that the desired trajectory plan is one dimensional, so using curvilinear coordinates will significantly simplify the solution.
+This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
 We show how two dimensional concerns (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
-All planning is longitudinal; that is, speed is the primary focus, and we assume that a separate yaw controller will ensure that the vehicle follows the prescribed path.
+All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
 

### 2016-01-22 root.tex
clarified Fig 5, 9

@@
 This requires ... \textbf{something about computing actions through stop signs before it's necessary and how that's inefficient}.
-We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:segmentation}, and  execute them incrementally.
+We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and  execute them incrementally.
 
@@
   \caption{
-    Segmentation of the path by path manager.
-    The trajectory generation is serially called on each segment colored differently.
+    Dividing the path into subpaths at stop signs.
+    The trajectory generation is serially called on each segment which is colored differently here.
   }
-  \label{fig:segmentation}
+  \label{fig:subpathdivision}
 \end{figure}
@@
   \subfigure[
-    Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
-    Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
+    Result of dividing a single subpath into separate segments.
+    The clustering algorithm used here sets segments boundaries before and after curves.
+    The subpath segments in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
   ]{
@@
   \subfigure[
-    Example of $v_m$ consolidation for simple univariate path clustering which corresponds to the path shown above.
+    Example of how the subpath's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
+    The x-axis corresponds to distance along the subpath above.
     The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
@@
   }
-  \caption{Path segmentation and speed ceiling consolidation procedures.}
+  \caption{Dividing a subpath into segments and determining speed ceilings for each segment.}
   \label{fig:1Dto2DandSegmentation}

### 2016-01-22 root.tex
updated segmentation section to introduce math quantities earlier

@@
 Prescribed initial and end conditions for position, speed, and acceleration must also be honored.
+All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
 This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
 We show how two dimensional concerns (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
-All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Path Segmentation} \label{sec:pathsegmentation}
+
+Our approach to fitting appropriate jerk profiles to complex subpaths is to further break each subpath into segments which have very simple shape.
+Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
+Each segment will be defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
+Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.
+
+The segmentation process proceeds as follows:
+\begin{enumerate} \label{asdf}
+  \item \emph{Ceiling calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ are found for every point in the given subpath.
+  \item \emph{Clustering}, in which subpath points are grouped sequentially and the segment boundaries are determined..
+  \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
+\end{enumerate}
+After this is done, the process described in Section \ref{sec:jerkprofiles} may be applied to each of the segments so that a plan of jerk vs. time may be generated.
+
+The ceiling calculation is a minimax problem.
+Each of the 9 scalars is a minimum of several maxima computed from an arbitrary set of constraints.
+For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the subpath.
+
+% How we actually did this.
+For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$, where $N_p$ is the number of waypoints in the subpath.
+The value $\kappa_k$ denotes the path curvature at each point.
+For each pathpath point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+
+This information is used to choose the segment boundaries in the clustering step.
+Any number of clustering algorithms may be used here, and prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
+For the present example, we will choose a simple univariate heuristic which essentially identifies curves in the roadway using only the speed ceiling.
+Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
+Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
+Figure \ref{fig:1Dto2DandSegmentation} shows the result of this algorithm.
+
+\begin{figure}[thpb]
+  \centering
+  \subfigure[
+    Result of dividing a single subpath into separate segments.
+    The clustering algorithm used here sets segments boundaries before and after curves.
+    The subpath segments in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
+  ]{
+    \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
+  }
+  \\
+  \subfigure[
+    Example of how the subpath's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
+    The x-axis corresponds to distance along the subpath above.
+    The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
+    The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
+    The resultant speed ceiling is shown in blue.
+  ]{
+  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
+  }
+  \caption{Dividing a subpath into segments and determining speed ceilings for each segment.}
+  \label{fig:1Dto2DandSegmentation}
+\end{figure}
+
+% need better notation here.
+In the consolidation step, a single set of $\mathbf{g}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ must be calculated for each segment using the values corresponding to each of the waypoints which comprise the segment in question.
+The minimum of scalar across all constituent points is chosen.
+While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
+In the present example, this means that in the straight portions of the subpath $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
+As a result, many of the curves will have a constant speed.
+For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Fitting Jerk Profiles to Each Segment} \label{sec:jerkprofiles}
+
+To introduce the process of computing a jerk profile for each segment, let us consider the case in which a vehicle needs to travel some arbitrary distance ahead.
+The vehicle is moving at some small initial speed and needs to accelerate to some traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+Figure \ref{fig:full7phasespec} shows this trajectory over time.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Jerk Profiles} \label{sec:jerkprofiles}
-
-Consider, as a starting point, the case in which a vehicle is moving at some small initial speed and needs to accelerate to some traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
-Figure \ref{fig:full7phasespec} shows this trajectory over time.
 The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
-It is assumed that the final forward acceleration will be zero.
-If the magnitude of the desired peak acceleration and jerk ($a_m$ and $j_m$, respectively) are known, then a profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
-Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior \emph{Citation}.
-The remaining parameters yet to be set now become $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
+Final forward acceleration is set to zero.
+If $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
+are known, then a profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior.
 As we show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
@@
 \ref{eq:am} and \ref{eq:jm}.
-This is computed for every point along the path.
+This is computed for every point along the subpath.
 In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.
@@
 As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4 phase profiles has been attempted and subsequently failed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Path Segmentation} \label{sec:pathsegmentation}
-
-% make sure to incorporate extensibility to other constraints
-The goal of the segmentation process is to divide a given path into multiple portions where each has a uniform upper limit on speed, acceleration, and braking.
-There are 3 steps in the segmentation process:
-\begin{enumerate} \label{asdf}
-  \item \emph{Ceiling calculation}, where $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ are found for every point in the given path.
-  \item \emph{Clustering}, in which path points are grouped sequentially.
-  \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
-\end{enumerate}
-After this is done, the process described in Section \ref{sec:jerkprofiles} may be applied to each of the segments.
-
-The ceiling calculation is a minimax problem.
-Each of the 9 scalars is a minimum of several maxima computed from an arbitrary set of constraints.
-For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the path.
-
-% How we actually did this.
-For the purposes of developing a simple example application, two constraints will be used for $v_{m,k}$: legal speed limit ($SL$, set constant for the whole path) and lateral acceleration limit $a_{B,y}^{max}$.
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$, where $N_p$ is the number of waypoints in the full path.
-\emph{RG: this is the first place we introduce two subscripts. please explain what the two subscripts stand for in general for example in $a_{B,y}$ and $v_{m,k}$ ?????? }
-The value $\kappa_k$ denotes the path curvature at each point.
-For each path point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
-
-This information is used to choose the segment boundaries in the clustering step.
-Any number of clustering algorithms may be used here, and prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
-For the present example, we will choose a simple univariate heuristic which essentially identifies curves in the roadway.
-Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
-Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
-Figure \ref{fig:1Dto2DandSegmentation} shows the result of this algorithm.
-
-% \begin{figure}[thpb]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
-%   \caption{
-%     Clustering output for a short path when using an algorithm which draws segment boundaries around curves.
-%     Portions in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
-%   }
-%   \label{fig:course_highlight_turns}
-% \end{figure}
-
-% \begin{figure}[thpb]
-%   \centering
-%   \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
-%   \caption{Showing the path projected onto the one dimensional s coordinate system}
-%   \label{fig:scoord}
-% \end{figure}
-
-% \begin{figure}[thpb]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
-%   \caption{
-%     Example of $v_m$ consolidation for simple univariate path clustering which corresponds the the path shown in Figure \ref{fig:course_highlight_turns}.
-%     The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
-%     The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
-%     The resultant speed ceiling is shown in blue.
-%   }
-%   \label{fig:consolidation_speed_ceiling}
-% \end{figure}
-
-\begin{figure}[thpb]
-  \centering
-  \subfigure[
-    Result of dividing a single subpath into separate segments.
-    The clustering algorithm used here sets segments boundaries before and after curves.
-    The subpath segments in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
-  ]{
-    \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
-  }
-  \\
-  % \subfigure[
-  %   Showing the path projected onto the one dimensional s coordinate system
-  % ]{
-  %   \includegraphics[width=1.0\columnwidth]{graphics/1Dsegmentation.png}
-  % }
-  % \\
-  \subfigure[
-    Example of how the subpath's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
-    The x-axis corresponds to distance along the subpath above.
-    The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
-    The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
-    The resultant speed ceiling is shown in blue.
-  ]{
-  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
-  }
-  \caption{Dividing a subpath into segments and determining speed ceilings for each segment.}
-  \label{fig:1Dto2DandSegmentation}
-\end{figure}
-
-% need better notation here.
-In the consolidation step, a single set of $\mathbf{g}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ must be calculated for each segment using the values corresponding to each of the path points which comprise the segment in question.
-The minimum of scalar across all constituent points is chosen.
-While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
-In the present example, this means that in the straight portions of the path $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
-As a result, many of the curves will have a constant speed.
-For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
 

### 2016-01-23 root.tex
update to section IV-A... oops

@@
 
-The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
 Final forward acceleration is set to zero.
-If $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
-are known, then a profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
-Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior.
-As we show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
+A profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+As we show later, $\mathbf{a}_m$, and $\mathbf{j}_m$ must sometimes be modified to obtain physically possible time intervals.
 To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs.
 \ref{eq:am} and \ref{eq:jm}.
-This is computed for every point along the subpath.
 In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.

### 2016-01-23 root.tex
clarifying and reducing jerk profile sectionsn

@@
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k$, $\mathbf{a}_{m,k}$, and $\mathbf{j}_{m,k}$ where $k = 1 ...
+N_s$ and can be solved individually in sequence.
 
@@
 Final forward acceleration is set to zero.
-A profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+A profile with $M$ constant jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
 As we show later, $\mathbf{a}_m$, and $\mathbf{j}_m$ must sometimes be modified to obtain physically possible time intervals.
@@
 For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
-We propose to define 4 other basic profiles, and our algorithm judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
-The other profiles are explained below in Section \ref{sec:jerkprofiles}.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved individually in sequence.
-Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
+We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$., and our algorithm judiciously chooses the appropriate profile.
+The other profiles are explained below.
 
-There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$.
 All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
 It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
-To find closed-form solutions for each, a symbolic solver was used.
-The target values for each profile are the time steps between jerk changes.
-These are solved by plugging in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$.
+To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used.
 
 For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
-This results in a negative time interval for that phase.
+This results in a negative time interval for one or more phases.
 In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
@@
 To decide which profile to employ, a cascade approach is used.
-Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real.
+Analytic values of time intervals are evaluated for each profile by plugging in new values of $\mathbf{a}_m$ and $\mathbf{j}_m$ until all time intervals are positive and real.
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.

### 2016-01-23 root.tex
overall changes

--- 2016-01-22 root.tex
@@
 \author{
-  Robert G. Cofield $^{1}$ and
-  Rakesh Gupta $^{2}$
+  Robert G. Cofield$^{1}$ and
+  Rakesh Gupta$^{2}$
   \thanks{
@@
 \begin{abstract}
-
 In this paper, we address the problem of trajectory planning for urban roads while reacting to the presence of pedestrians.
@@
 Given a desired path, we present a novel online method for planning longitudinal trajectories to follow the urban path with turns and intersections while honoring traffic regulations such as stop signs.
-Based on sensor data, our method updates the trajectory plan to safely avoid pedestrians on the road by slowing or stopping.
+Based on sensor data, our method updates the trajectory plan to safely avoid pedestrians on the road by slowing down or stopping.
 This is critical in urban environments, when modifying the vehicle's intended path is often not possible.
@@
 When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
-It is common to employ the path-velocity decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
+It is common to employ the Path-Trajectory decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-First is Path Planning, which involves computation of the optimal path.
-The second task is Trajectory Planning, which involves the computation of the fastest velocity profile that is appropriate for the scenario.
+First is {\it Path Planning}, which involves computation of the optimal path.
+The second task is {\it Trajectory Planning}, which involves the computation of the fastest velocity profile that is appropriate for the scenario.
 
@@
 On the other hand, many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
-In this paper, we describe an trajectory generation system that satisfies these constraints for human comfort and is reactive to presence of pedestrians on the road.
+In this paper, we describe a trajectory generation system that satisfies these constraints for human comfort and is reactive to presence of pedestrians on the road.
 %Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
@@
 
-In this paper, we present an autonomy framework for pedestrian aware ground vehicles.
+We present an autonomy framework for pedestrian-aware ground vehicles.
 Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
@@
 At stop signs the car should stop and wait for an appropriate amount of time.
-It should also react to pedestrians on the road - at intersections as well as on the road.
-Pedestrians on the sidewalks should be ignored by the autonomous cars.
-Further the pedestrians can be static or moving.
-The car should stop, slow down and circum-navigate around pedestrians when appropriate.
+It should also react to pedestrians - at intersections as well as on the road.
+Pedestrians on the sidewalks should be ignored by the autonomous car.
+Further, the pedestrians can be static or moving.
+The car should stop, slow down and go around pedestrians when appropriate.
 
@@
 The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking control.
-We present an online method of revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing or stopping when no alternative paths are available.
+We present an online method of revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.
 
@@
 We also describe other subsystems that our system needs to work with, and planned stops.
-In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies jerk and maximum velocity and acceleration constraints. 
+Section \ref{sec:pathmanager} describes the high level Path Manager that calls the trajectory planner as necessary to 
+exhibit reactive behavior and stop for pedestrians that may appear on the road.
+In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
 We explain how we handle intersections, turns in the road and planned stops due to stop signs (that have already been marked in the map).
@@
 This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
-Section V describes the high level Path Manager that calls the trajectory planner as necessary to 
-exhibit reactionary behavior and stop for pedestrians that may appear on the road.
 We then conclude and discuss future extensions.
@@
 Our system architecture is shown in Figure \ref{fig:addreact}.
-There are 4 primary input components: a path planner, navigation filter, pedestrian detector, and a map which provides the location of stop signs.
+There are 4 primary input components: Navigation Filter, Pedestrian Detector, a Path Planner, and a Map which provides the location of stop signs.
 % nav filter
@@
   \caption{
-   Primary system components. The ``Path Manager'' and ``Trajectory Planner'' scomponents (in blue) are presented herein.
+   Primary system components. The {\it Path Manager} and {\it Trajectory Planner} components (in blue) are presented herein.
+   \textbf{RG: blue color is too dark to read text ????????????}
   }
@@
 It computes a bounding box for each pedestrian as shown in Figure \ref{fig:ped}.
-The pedestrian position is transformed to the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
+The pedestrian position is computed in the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
 
@@
 We assume that we have access to a path planner.
-For example, this can be a service similar to Google Maps except that it provides lane level maps. 
+For example, this can be a service similar to Google Maps except that instead of road level route, it provides lane level route. 
 We assume that the path planner operates independently from the trajectory planner and tracker.
@@
 Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the destination.
-This plan is output as set of global waypoints, with the waypoints corresponding to stop signs flagged.
+This plan is output as a set of global waypoints, with the waypoints corresponding to stop signs flagged.
 This same map database is used to search for stop signs and provide their locations.
@@
   \caption{
-    A car turning right at the intersection only needs to consider pedestrians in the pedestrian crossings shown yellow.
-    Pedestrians in the green crossings can be ignored unless they appear somewhere on the road.
+    A car turning right at the intersection only needs to consider pedestrians in the pedestrian crossings shown in yellow.
+    Pedestrians at the green crossings can be ignored unless they appear somewhere on the road.
   }
@@
 
-Beyond the input subsystems, there are three other major components: a high-level planner, a low-level planner, and a software controller which governs engine speed, braking, and steering.
-The controller takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
-This short-term trajectory is oriented in the vehicle body frame.
-\textbf{Citation for controller ????}.
+Beyond these input subsystems, our system has three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.
 
-The high-level planner is referred to as the ``path manager'', and performs decision-making as well as governing operation of the low-level planner.
+The high-level planner is referred to as the {\it Path Manager}, and performs decision-making as well as governing operation of the low-level planner.
 Its primary functions are:
 \begin{enumerate}
-  \item Stop the vehicle at stop signs and wait until the assosciated intersection is clear before continuing.
+  \item Stop the vehicle at stop signs and wait until the associated intersection is clear before continuing.
   \item Reactively reduce speed or stop altogether for pedestrians which pose a collision risk, then continue when they are clear.
-  \item Govern the trajectory planner.
+  \item Govern the Trajectory Planner.
 \end{enumerate}
-The low-level planner is referred to as the ``trajectory planner'', and is responsible for quantifying the actions dictated by the path manager.
+The low-level planner is referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
 Its primary functions are:
@@
 \end{itemize}
-This work focuses on the path manager (Sec. \ref{sec:pathmanager}) and the trajectory planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
+
+The Software Controller governs engine speed, braking, and steering.
+The controller takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
+This short-term trajectory is oriented in the vehicle body frame.
+\textbf{Citation for controller ????}.
+This work focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
 
@@
   \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
-  \caption{State Transition Diagram}
+  \caption{State Transition Diagram for ?????????? more detail????? }
   \label{fig:st}
@@
 This requires ... \textbf{something about computing actions through stop signs before it's necessary and how that's inefficient}.
-We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and  execute them incrementally.
+We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call Trajectory Planner for each sub-path incrementally.
 
-In the absence of pedestrians, the path manager cycles between PSTOP and NORMAL states.
+In the absence of pedestrians, the path manager cycles between NORMAL and PSTOP states.
 In NORMAL state, the path manager send the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
@@
   \caption{
-    Dividing the path into subpaths at stop signs.
-    The trajectory generation is serially called on each segment which is colored differently here.
+    Dividing the path into sub-paths at stop signs.
+    The trajectory generation is serially called on each sub-path which are colored differently here.
   }
@@
 
-All detected pedestrians locations are compared against the high-resolution digital map, so that pedestrians which are not on the roadway or near an upcoming crosswalk are not considered.
+All detected pedestrian locations are compared against the high-resolution digital map, so that pedestrians that are not on the roadway or near an upcoming crosswalk are not considered.
 Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
-Pedestrians which are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
-All the pedestrians which remain pose a potential collision risk, however only the closest is considered for RSTOP.
+Pedestrians that are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
+All the remaining pedestrians pose potential collision risks, however only the closest pedestrian is considered for RSTOP.
 
@@
 
-Next the nominal distance needed to stop $\Delta s_{RSTOP}$ must be computed given the vehicle's current location, speed, and forward acceleration.
-The procedure for doing this will be the same as the procedure for planning the RSTOP trajectory later with a few steps omitted, since the planning algorithm is extremely efficient.
+When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
+The procedure for doing this is the same as the procedure for planning the RSTOP trajectory later with a few steps omitted, since the planning algorithm is extremely efficient.
 
 Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
-RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= s_{ped} - s_{veh}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
+RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= \Delta s_{avail}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
 Planning and executing the trajectory for coming to a halt is detailed in Sec. \ref{sec:reactivestoptrajectory}.
@@
 This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
-This value is typically set to 1 meter.
+$\Delta s_{replan}$ value is typically set to 1 meter.
 Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$.
@@
 
-The task trajectory planning in this context is to take a 2 dimensional (horizontal plane) set of waypoints from the path manager, and compute a series of constant jerk intervals for motion along the length of the path.
+The task of Trajectory Planning is to take a 2 dimensional (horizontal plane) set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
 This trajectory must be time optimal yet still fall within constraints on speed, acceleration, and jerk.
@@
 
-Prescribed initial and end conditions for position, speed, and acceleration must also be honored.
+Prescribed initial and end conditions for position, speed, and acceleration must be honored.
 All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
 This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
-We show how two dimensional concerns (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
+We show how two dimensional \textbf{concerns????} (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
 
@@
 
-Our approach to fitting appropriate jerk profiles to complex subpaths is to further break each subpath into segments which have very simple shape.
+Our approach to fitting appropriate jerk profiles to complex sub-paths is to further break each subpath into segments which have very simple shape.
 Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
@@
 \begin{enumerate} \label{asdf}
-  \item \emph{Ceiling calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ are found for every point in the given subpath.
+  \item \emph{Ceiling calculation}, where 
+%$v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
+initial/final speeds, speed limit, peak acceleration and jerk 
+are found for every point in the given subpath.
   \item \emph{Clustering}, in which subpath points are grouped sequentially and the segment boundaries are determined..
@@
 The ceiling calculation is a minimax problem.
-Each of the 9 scalars is a minimum of several maxima computed from an arbitrary set of constraints.
+Each of the 9 scalars \textbf{RG: please explain 9 scalars ???} is a minimum of several maxima computed from an arbitrary set of constraints.
 For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the subpath.
@@
 The value $\kappa_k$ denotes the path curvature at each point.
-For each pathpath point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+For each \textbf{??? sub-path ??} point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
 
@@
   \subfigure[
-    Example of how the subpath's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
+    Example of how the sub-path's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
     The x-axis corresponds to distance along the subpath above.
@@
 % need better notation here.
-In the consolidation step, a single set of $\mathbf{g}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ must be calculated for each segment using the values corresponding to each of the waypoints which comprise the segment in question.
+In the consolidation step, a single set of 
+%$\mathbf{g}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
+initial/final speeds, speed limit, peak acceleration and jerk 
+must be calculated for each segment using the values corresponding to each of the waypoints which comprise the segment in question.
 The minimum of scalar across all constituent points is chosen.
@@
   \caption{
-    7 phase constant jerk profile along curvilinear distance axis.
+    7-phase constant jerk profile along curvilinear distance axis.
     All values needed to completely parameterize a single path segment for trajectory are labeled here.}
@@
 We propose to define 4 other basic profiles, and our algorithm judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
-The other profiles are explained below in Section \ref{sec:jerkprofiles}.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved individually in sequence.
+
+\textbf{RG: does not belong here: The other profiles are explained below in Section \ref{sec:jerkprofiles}.
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ...
+N_s$ and can be solved individually in sequence.
 Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
+}
 
@@
 It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
-To find closed-form solutions for each, a symbolic solver was used.
+To find closed-form solutions for each, a symbolic solver is used.
 The target values for each profile are the time steps between jerk changes.
@@
 
-\subsubsection{7 Phase} \label{sec:7phase}
+\subsubsection{7-phase} \label{sec:7phase}
 
-This profile, with $\Delta t_i$ as shown in Figure {fig:full7phasespec}, is always attempted first.
+This profile, with $\Delta t_i$ as shown in Figure \ref{fig:full7phasespec}, is always attempted first.
 All time steps have a single solution, enabled by the fact that the initial and end conditions are known, and the two peak accelerations and two jerk magnitudes are provided as configuration parameters.
 In the case where $\Delta t_2$ or $\Delta t_6$ are negative, accel-jerk tuning must be employed.
-In the instance where $\Delta t_4$ is negative, the total segment length $L$ is too short to reach top speed of $v_m$, so a 6 phase profile must be used.
+In the instance where $\Delta t_4$ is negative, the total segment length $L$ is too short to reach top speed of $v_m$, so a 6-phase profile must be used.
 
-\subsubsection{6 Phase} \label{sec:6phase}
+\subsubsection{6-phase} \label{sec:6phase}
 
-In the case mentioned above, the speed limit is too high for a solution to exist for a 7 phase profile.
+In the case mentioned above, the speed limit is too high for a solution to exist for a 7-phase profile.
 The middle travel period $\Delta t_4$ is removed, so that the vehicle increases speed and then immediately decreases speed. 
@@
 In the case where both solutions are negative, accel-jerk tuning is attempted. 
-Should that yield no positive time intervals, one of the two 4 phase profiles then becomes appropriate, since $L$ is not large enough to accommodate accelerating past the end speed.
-If $v_i > v_f$, then a reversed 4 phase profile is attempted.
-However, if $v_i > v_f$, then a regular 4 phase profile is attempted.
+Should that yield no positive time intervals, one of the two 4-phase profiles then becomes appropriate, since $L$ is not large enough to accommodate accelerating past the end speed.
+If $v_i > v_f$, then a reversed 4-phase profile is attempted.
+However, if $v_i > v_f$, then a regular 4-phase profile is attempted.
 
@@
   \begin{tabular}{cc}
-  \subfigure[6 phase profile] {
+  \subfigure[6-phase profile] {
   \includegraphics[width=0.4\columnwidth]{graphics/6phase_v(s).png}
@@
   } &
-  \subfigure[4 phase profile] {
+  \subfigure[4-phase profile] {
   \includegraphics[width=0.4\columnwidth]{graphics/4phase_v(s).png}
@@
   } &
-  \subfigure[3 phase profile] {
+  \subfigure[3-phase profile] {
   \includegraphics[width=0.4\columnwidth]{graphics/3phase_v(s).png}
@@
   \end{tabular}
-  \caption{6, 4, 4Reversed and 3 phase profiles, displayed as speed vs. position along path}
+  \caption{6, 4, 4Reversed and 3-phase profiles, displayed as speed vs. position along path}
 
@@
 
-\subsubsection{4 Phase} \label{sec:4phase}
+\subsubsection{4-phase} \label{sec:4phase}
 
@@
 
-\subsubsection{Reversed 4 Phase} \label{sec:reversed4phase}
+\subsubsection{Reversed 4-phase} \label{sec:reversed4phase}
 
@@
 
-\subsubsection{3 Phase} \label{sec:3phase}
+\subsubsection{3-phase} \label{sec:3phase}
 
 A change from one speed to another.
-All of the above profiles are composed of some combination of a 3 phase profile and a $j_k = 0$ phase.
+All of the above profiles are composed of some combination of a 3-phase profile and a $j_k = 0$ phase.
 This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
-As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4 phase profiles has been attempted and subsequently failed.
+As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4-phase profiles has been attempted and subsequently failed.
 
@@
 
-\subsection{Unified Solution} \label{sec:unifiedsolution}
+\subsection{Stitching segment profiles together} \label{sec:unifiedsolution}
 
-At this point, the path has been divided into segments and conditions have been computed such that an analytically derived plan can be fit to each segment individually.
-Now a reference trajectory must be generated for the entire path.
+At this point, each sub-path has been divided into segments and conditions have been computed such that an analytically derived plan can be fit to each segment individually.
+Now a reference trajectory must be generated for the entire sub-path.
 Continuity of position, speed, and acceleration is the primary goal of this process.
@@
 
-\emph{RG: please explain reference time, $s$ and $\Delta s$ ????????}
+\textbf{RG: please explain reference time, $s$ and $\Delta s$ ????????}
 
@@
 
-However, the time at which the trajectory plan must be referenced is not always known.
-The solution to this problem is known as trajectory tracking control, and will be discussed in the next section.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\textbf{RG: abrupt transition. please add a few lines on current sub-path being modified into 2 sub-paths. And there being a presence of a pedestrian}
+
 This is the procedure for determining nominal stopping distance required.
-A 3 phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+A 3-phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
 
 While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
-As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3 phase profile, but not both, since no closed-form solution exists for that case.
+As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3-phase profile, but not both, since no closed-form solution exists for that case.
 Normal planning enforces $L$ to avoid skewing segment boundaries within each subpath.
 However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
-Reformulating a new 3 phase jerk profile accordingly is a simple matter for a symbolic solver.
+Reformulating a new 3-phase jerk profile accordingly is a simple matter for a symbolic solver.
 %%% This is the brief way of wording the preceding three sentences.
@@
 
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
-  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
-  \label{fig:pathtostop}
-\end{figure}
+%\begin{figure}[thpb]
+%  \centering
+%  \includegraphics[width=1.0\columnwidth]{graphics/RSTOP_path_truncation.png}
+%  \caption{Green color path is the one over which the car finds a trajectory to stop. please show green as dashes for b\&w printing ??????????? }
+%  \label{fig:pathtostop}
+%\end{figure}
+
+However, the time at which the trajectory plan must be referenced is not always known.
+The solution to this problem is known as trajectory tracking control, and will be discussed in the next section.
 
@@
   \includegraphics[width=0.5\columnwidth]{graphics/PathProjection.png}
-  \caption{Showing projection of current position of car shown by green triangle to s coordinate system}
+  \caption{Showing projection of current position of car shown by green triangle to s coordinate system
+   \textbf{ RG: please label link in figure if possible} }
+
   \label{fig:cartos}
@@
 
-The general idea is to project the vehicle's current position onto the path then find the cumulative distance along the path between the start point and the projection point.
+\textbf{RG: this paragraph along with above figure should be merged with 2D to 1D s transformation section.
+in the figure please show a line for s coordinate as well} The general idea is to project the vehicle's current position onto the path, and then find the cumulative distance along the path between the start point and the projection point.
 This is the 1D location corresponding to the vehicle's current reference time.
@@
 \ref{fig:cartos}).
-Projections which lie outside of the link's two constituent path points are snapped to that which is nearest.
+Projections which lie outside of the link's two constituent path points are snapped to that which is the nearest.
 The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection $\mathbf{r}_p$.
-Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex paths which overlap or self-intersections.
+Any number of modifications can be made using heading, velocity vectors, and previous information to allow this method to be extended to complex overlapping paths or self-intersections.
 
-Choosing an algorithm to obtain the current one dimensional position of the vehicle along the length of the path $s_{veh}$ (i.e., in the same one-dimensional coordinate system that planning was performed) is influenced by the form of the path plan.
+Choosing an algorithm to obtain the current one dimensional position of the vehicle along the length of the path $s_{veh}$ (i.e., in the same one-dimensional coordinate system in which the planning was performed) is influenced by the form of the path plan.
 Assuming that the path is linear between waypoints reduces computational effort significantly.
-So $s_{veh}$ at any epoch is then sum of the link lengths behind it's perpendicular projection$\mathbf{r}_p$.
+So $s_{veh}$ at any epoch is the sum of the link lengths behind its perpendicular projection $\mathbf{r}_p$.
 However, the error in this approximation approaches zero as the spacing between path points approaches zero.
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$,F may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$, F may be approximated \textbf {??????
+as cubic if a spline is ????
+} being constructed from knots which have a very high resolution.
 In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
@@
 
+\textbf{RG: abrupt transition. need to summarize here}
+
+\textbf{RG: Ambarish suggests to replace all speed except speed limit by velocity and speeds -> velocities}
+
 \section{Summary} \label{sec:summary}
 
-In this paper, we have presented a novel online method for planning longitudinal trajectories 
-to follow the urban path with real time updates to avoid pedestrians on the roadway by slowing or 
-stopping. We built a robust stopping model as part of Trajectory planner reliably integrated with 
+In this paper, we have presented a novel real-time method for planning longitudinal trajectories 
+in an autonomous car to follow an urban path with online updates to avoid pedestrians on the roadway by slowing down or 
+stopping. We built a robust stopping model as part of Trajectory Planner reliably integrated with 
 the Path Manager.
 We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
-they leave. Our system handles corner cases gracefully, following pedestrians if they continue walking 
-on the road away from the car. Our system works with multiple pedestrians as we only 
-need to consider the closest pedestrian ahead of us.
+they leave. Our system handles complex cases that were not explicitly programmed gracefully, such as following 
+pedestrians if they continue walking on the road away from the car. Our system works with multiple 
+pedestrians as we only need to consider the closest pedestrian ahead of us.
 
 We make three contributions in this paper. First, we present an integrated trajectory planner that
-simultaneously limits jerk, velocity and acceleration while being responsive to presence of pedestrians.
+simultaneously limits jerk, velocity and acceleration to pre-set desired values while being responsive 
+to the presence of pedestrians.
 Second, our trajectory planner has closed form solutions, and is suitable for online implementation
 on a car, unlike systems requiring solutions to non linear optimizations. Finally, we have confirmed the
-efficiency and reliability of our trajectory planner on a test vehicle with 10s of hours of testing
-under autonomous driving conditions on urban roads with pedestrians in random locations.
+efficiency and reliability of our trajectory planner on a test vehicle with over 100 hours of testing
+under fully autonomous driving conditions on urban roads with pedestrians in random locations.
 
 In future, we would like to handle other roadway traffic, 
-coexisting with other vehicles. We would also like to perceive and stop for red-lights.
+coexisting with other vehicles. We would also like to perceive and stop for stop signs and red-lights,
+rather than using map information.
 

### 2016-01-23 root.tex
killed unification section by merging with preceding sections

--- 2016-01-23 root.tex
@@
 
+A jerk vs. time profile will be fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
+Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
+The values of $a_i$ and $v_i$ for the first segment are set to be the value which is currently reported from sensor data.
+Since the final acceleration is to be set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
+For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
+Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments. 
+Continuity of position, speed, and acceleration is the primary goal of this process.
+
+\begin{figure}
+  \begin{algorithmic}[1]
+    \Procedure{SegmentBoundarySpeeds}{$v_m$}
+      \State $\mathbf{v}_p \gets \mathbf{v}_m$ \Comment{Initialize targets as ceilings}
+      \State $\mathbf{v}_{p,0} \gets v_i$ \Comment{Set starting speed to current}
+      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
+      \For{$i = 1, ..., N_s-1$}
+        \If{$v_{m,i+1} > v_{m,i}$}
+          \State $v_{p,i+1} \gets v_{m,i+1}$
+        \Else
+          \State $v_{p,i+1} \gets v_{m,i}$
+        \EndIf
+      \EndFor
+    \EndProcedure
+  \end{algorithmic}
+  \caption{
+    Algorithm to set speed boundary conditions for interior segments.
+    Having consolidated the waypoints in to a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
+    This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
+  }
+\label{alg:segmentspeedboundaryconditions}
+\end{figure}
+
+Now that all segments within the subpath have been parameterized, they may be solved sequentially.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
+Each is described below.
 
@@
 A change from one speed to another.
-All of the above profiles are composed of some combination of a 3 phase profile and a $j_k = 0$ phase.
+All of the above profiles are composed of some combination of a 3 phase profile and a $j = 0$ phase.
 This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
-As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4 phase profiles has been attempted and subsequently failed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Unified Solution} \label{sec:unifiedsolution}
-
-At this point, the path has been divided into segments and conditions have been computed such that an analytically derived plan can be fit to each segment individually.
-Now a reference trajectory must be generated for the entire path.
-Continuity of position, speed, and acceleration is the primary goal of this process.
-
-A jerk vs. time profile is fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
-Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
-The values of $a_i$ and $v_i$ for the first segment is set to be the value which is currently reported from sensor data.
-Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
-For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
-Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments.
-
-\begin{figure}
-  \begin{algorithmic}[1]
-    \Procedure{SegmentBoundarySpeeds}{$v_m$}
-      \State $\mathbf{v}_p \gets \mathbf{v}_m$ \Comment{Initialize targets as ceilings}
-      \State $\mathbf{v}_{p,0} \gets v_i$ \Comment{Set starting speed to current}
-      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
-      \For{$i = 1, ..., N_s-1$}
-        \If{$v_{m,i+1} > v_{m,i}$}
-          \State $v_{p,i+1} \gets v_{m,i+1}$
-        \Else
-          \State $v_{p,i+1} \gets v_{m,i}$
-        \EndIf
-      \EndFor
-    \EndProcedure
-  \end{algorithmic}
-  \caption{
-    Algorithm to set speed boundary conditions for interior path segments.
-    Having consolidated the waypoints in to a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
-    This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
-  }
-\label{alg:segmentspeedboundaryconditions}
-\end{figure}
-
-% talk about control points
-Now each segment is solved individually.
-In the case where a segment requires a speed change which is infeasible given limits on acceleration and jerk, then the final speed for that segment is planned to be the closes possible value.
+As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial conditions.
 When these discrepancies in the feasible $v_f$ for a segment arise, the corresponding value in $v_p$ is updated so that the new conditions may be accounted for in the subsequent segment.
 
-At the end of the process, there will be a set of $N_{cp}$ ``control points''.
-A control point is a vector $[t_{ref}, s, v_{B,x}, a_{B,x}, j_{B,x}]$ which describes the longitudinal kinematic state of the target vehicle at the moment of an infinite jounce jerk change.
-Fitting a jerk profile to each segment results in control points which must be stitched into the control points for the preceding segments.
-While no further information than initial conditions, time durations, and jerk values is needed (simple integration would yield lower derivatives of position), this information is already available since it must be solved for each segment.
-This extra information may be used as a redundancy layer to verify that no errors occurred.
-For each segment with $M$ jerk phases, the following is produced:
+% \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}
+
+At this point, the subpath has been divided into segments and an analytically derived plan has been fit to each segment individually.
+Now a unified reference trajectory must be constructed by stitching these together.
+
+For each segment with $M$ jerk phases, the following values are computed, which correspond to the instants at which jerk changes:
 \begin{itemize}
   \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
-  \item $M+1$ values of distance elapsed within the segment $\Delta s$.
-These must be shifted forward by the last $s$ value from the previous segment to translate them into the 1D path domain.
-Additionally, the first value must be erased.
-  \item $M+1$ values of $v_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
-  \item $M+1$ values of $a_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
-  \item $M$ jerk values, which may be appended directly to the preceding jerk values
+  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values
 \end{itemize}
+It is now possible to integrate $j(t)$ to determine an instantaneous trajectory at any moment in time.
 
-\emph{RG: please explain reference time, $s$ and $\Delta s$ ????????}
+% After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
+% To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
+% Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
+% Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
+% Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
+% The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
 
-After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
-Since $j(t)$ is piecewise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
-To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
-Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
-Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
-Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
-The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
-
-However, the time at which the trajectory plan must be referenced is not always known.
-The solution to this problem is known as trajectory tracking control, and will be discussed in the next section.
+% However, the time at which the trajectory plan must be referenced is not always known.
+% The solution to this problem is known as trajectory tracking control, and will be discussed in Sec. \ref{sec:trajectorytracking}.
 
@@
 
-After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
+% After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
 The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
@@
 Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$,F may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
-In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
+In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:jerkprofiles}, and it is now resampled at $\Delta t = 0.01 sec$.
 This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.

### 2016-01-23 root.tex
removed dangling reference to deleted figure

--- 2016-01-23 root.tex
@@
 The value $\kappa_k$ denotes the path curvature at each point.
-For each \textbf{??? sub-path ??} point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+For each sub-path point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
 
@@
 
-The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
 Final forward acceleration is set to zero.
-If $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$.
-are known, then a profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
-Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior.
-As we show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
+A profile with $M$ constant jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
+As we show later, $\mathbf{a}_m$, and $\mathbf{j}_m$ must sometimes be modified to obtain physically possible time intervals.
 To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs.
 \ref{eq:am} and \ref{eq:jm}.
-This is computed for every point along the subpath.
 In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.
@@
 For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
-We propose to define 4 other basic profiles, and our algorithm judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.
+We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$., and our algorithm judiciously chooses the appropriate profile.
+The other profiles are explained below.
 
-\textbf{RG: does not belong here: The other profiles are explained below in Section \ref{sec:jerkprofiles}.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ...
-N_s$ and can be solved individually in sequence.
-Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}.
-}
-
-There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$.
 All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
 It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
-To find closed-form solutions for each, a symbolic solver is used.
-The target values for each profile are the time steps between jerk changes.
-These are solved by plugging in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$.
+To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used.
 
 For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
-This results in a negative time interval for that phase.
+This results in a negative time interval for one or more phases.
 In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
@@
 To decide which profile to employ, a cascade approach is used.
-Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real.
+Analytic values of time intervals are evaluated for each profile by plugging in new values of $\mathbf{a}_m$ and $\mathbf{j}_m$ until all time intervals are positive and real.
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
+Each is described below.
 
@@
 A change from one speed to another.
-All of the above profiles are composed of some combination of a 3-phase profile and a $j_k = 0$ phase.
+All of the above profiles are composed of some combination of a 3 phase profile and a $j = 0$ phase.
 This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
-As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4-phase profiles has been attempted and subsequently failed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Stitching segment profiles together} \label{sec:unifiedsolution}
-
-At this point, each sub-path has been divided into segments and conditions have been computed such that an analytically derived plan can be fit to each segment individually.
-Now a reference trajectory must be generated for the entire sub-path.
-Continuity of position, speed, and acceleration is the primary goal of this process.
-
-A jerk vs. time profile is fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
-Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
-The values of $a_i$ and $v_i$ for the first segment is set to be the value which is currently reported from sensor data.
-Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
-For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
-Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments.
-
-\begin{figure}
-  \begin{algorithmic}[1]
-    \Procedure{SegmentBoundarySpeeds}{$v_m$}
-      \State $\mathbf{v}_p \gets \mathbf{v}_m$ \Comment{Initialize targets as ceilings}
-      \State $\mathbf{v}_{p,0} \gets v_i$ \Comment{Set starting speed to current}
-      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
-      \For{$i = 1, ..., N_s-1$}
-        \If{$v_{m,i+1} > v_{m,i}$}
-          \State $v_{p,i+1} \gets v_{m,i+1}$
-        \Else
-          \State $v_{p,i+1} \gets v_{m,i}$
-        \EndIf
-      \EndFor
-    \EndProcedure
-  \end{algorithmic}
-  \caption{
-    Algorithm to set speed boundary conditions for interior path segments.
-    Having consolidated the waypoints in to a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
-    This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
-  }
-\label{alg:segmentspeedboundaryconditions}
-\end{figure}
-
-% talk about control points
-Now each segment is solved individually.
-In the case where a segment requires a speed change which is infeasible given limits on acceleration and jerk, then the final speed for that segment is planned to be the closes possible value.
+As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial conditions.
 When these discrepancies in the feasible $v_f$ for a segment arise, the corresponding value in $v_p$ is updated so that the new conditions may be accounted for in the subsequent segment.
 
-At the end of the process, there will be a set of $N_{cp}$ ``control points''.
-A control point is a vector $[t_{ref}, s, v_{B,x}, a_{B,x}, j_{B,x}]$ which describes the longitudinal kinematic state of the target vehicle at the moment of an infinite jounce jerk change.
-Fitting a jerk profile to each segment results in control points which must be stitched into the control points for the preceding segments.
-While no further information than initial conditions, time durations, and jerk values is needed (simple integration would yield lower derivatives of position), this information is already available since it must be solved for each segment.
-This extra information may be used as a redundancy layer to verify that no errors occurred.
-For each segment with $M$ jerk phases, the following is produced:
+% \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}
+
+At this point, the subpath has been divided into segments and an analytically derived plan has been fit to each segment individually.
+Now a unified reference trajectory must be constructed by stitching these together.
+
+For each segment with $M$ jerk phases, the following values are computed, which correspond to the instants at which jerk changes:
 \begin{itemize}
   \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
-  \item $M+1$ values of distance elapsed within the segment $\Delta s$.
-These must be shifted forward by the last $s$ value from the previous segment to translate them into the 1D path domain.
-Additionally, the first value must be erased.
-  \item $M+1$ values of $v_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
-  \item $M+1$ values of $a_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
-  \item $M$ jerk values, which may be appended directly to the preceding jerk values
+  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values
 \end{itemize}
+It is now possible to integrate $j(t)$ to determine an instantaneous trajectory at any moment in time.
 
-\textbf{RG: please explain reference time, $s$ and $\Delta s$ ????????}
+% After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
+% To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
+% Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
+% Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
+% Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
+% The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
 
-After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
-Since $j(t)$ is piecewise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
-To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
-Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
-Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
-Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
-The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
+% However, the time at which the trajectory plan must be referenced is not always known.
+% The solution to this problem is known as trajectory tracking control, and will be discussed in Sec. \ref{sec:trajectorytracking}.
 
@@
 
-After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
+% After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
 The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
 Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
-This is depicted in Fig. \ref{fig:pathtostop}
+% This is depicted in Fig. \ref{fig:pathtostop}
 The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$, F may be approximated \textbf {??????
-as cubic if a spline is ????
-} being constructed from knots which have a very high resolution.
-In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
+Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$,F may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:jerkprofiles}, and it is now resampled at $\Delta t = 0.01 sec$.
 This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.

### 2016-01-23 root.tex
fix transition to summary and other minor changes

@@
 We also describe other subsystems that our system needs to work with, and planned stops.
-Section \ref{sec:pathmanager} describes the high level Path Manager that calls the trajectory planner as necessary to 
-exhibit reactive behavior and stop for pedestrians that may appear on the road.
+Section \ref{sec:pathmanager} describes the high level Path Manager that calls the Trajectory Planner as necessary to 
+exhibit reactive behavior and stop for pedestrians that appear on the road.
 In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
@@
 Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the destination.
-This plan is output as a set of global waypoints, with the waypoints corresponding to stop signs flagged.
-This same map database is used to search for stop signs and provide their locations.
+This plan is then output as a set of global waypoints.
 
@@
 
+Our High Resolution Map database is used to search for stop signs and provide their locations.
+The plan waypoints corresponding to stop signs are then flagged for later processing.
 We also have an In-Out Algorithm implemented via polygon intersection that can be used to ignore “safe” pedestrians
@@
 
-The task of Trajectory Planning is to take a 2 dimensional (horizontal plane) set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
+The task of Trajectory Planning is to take a 2 dimensional (horizontal plane) set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the sub-path.
 This trajectory must be time optimal yet still fall within constraints on speed, acceleration, and jerk.
@@
 \begin{enumerate} \label{asdf}
-  \item \emph{Ceiling calculation}, where 
-%$v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
-initial/final speeds, speed limit, peak acceleration and jerk 
+  \item \emph{Ceiling calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
+%$initial/final speeds, speed limit, peak acceleration and jerk 
 are found for every point in the given subpath.
@@
 % need better notation here.
-In the consolidation step, a single set of 
-%$\mathbf{g}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
-initial/final speeds, speed limit, peak acceleration and jerk 
+In the consolidation step, a single set of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
+%initial/final speeds, speed limit, peak acceleration and jerk 
 must be calculated for each segment using the values corresponding to each of the waypoints which comprise the segment in question.
@@
   } \\
-  \subfigure[4 Reversed phase profile] {
+  \subfigure[Reversed 4-phase profile] {
   \includegraphics[width=0.4\columnwidth]{graphics/4Rphase_all_derivatives.png}
@@
   \end{tabular}
-  \caption{6, 4, 4Reversed and 3-phase profiles, displayed as speed vs. position along path}
+  \caption{6, 4, Reversed 4 and 3-phase profiles, displayed as speed vs. position along path}
 
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$,F may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
+Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$, F may be approximated \textbf {as cubic if a spline ???????
+} is being constructed from knots which have a very high resolution.
 In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:jerkprofiles}, and it is now resampled at $\Delta t = 0.01 sec$.
@@
 
-\textbf{RG: abrupt transition. need to summarize here}
+To summarize, in the last three sections we have described how a urban path is broken into sub-paths,
+and computation of a velocity and jerk profile for segments within sub-paths that honor constraints
+for speed, velocity and acceleration as well as reactively slow down or stop for pedestrians on
+the urban path.
 
@@
 We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
-they leave. Our system handles complex cases that were not explicitly programmed gracefully, such as following 
+they leave. Our system handles complex cases that were not explicitly programmed, gracefully, such as following 
 pedestrians if they continue walking on the road away from the car. Our system works with multiple 

### 2016-01-23 root.tex
move lit review, changed 1d -> 2D paragraph location, changed subscripts

--- 2016-01-23 root.tex
@@
 When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
-It is common to employ the Path-Trajectory decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
+% "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
+It is common to employ the path-velocity decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
@@
 A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
-Some preliminary work on optimizing these constraints on urban roads has been done ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
+Similar work has been performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013}, which sets the travel time a priori and optimally minimizes jerk but does not set hard limitations or control jerk directly.
+Some preliminary work on optimizing these constraints on urban roads has also been done by ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
 On the other hand, many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
+This approach also makes the assumption that alternative paths around pedestrians are available, which may be invalid in urban scenarios.
+Planners which optimize human comfort by modifying the whole trajectory to seek obstacle-free space \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
 In this paper, we describe a trajectory generation system that satisfies these constraints for human comfort and is reactive to presence of pedestrians on the road.
+We enforce hard upper limits on jerk without requiring that travel time be known before planning occurs.
 %Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
@@
 
-Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path.
+Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path without requiring prior knowledge of travel time.
 Our method first segments the path using an arbitrary set of kinematic or dynamic constraints which may be functions of the desired path.
@@
 The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
-The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking control.
+The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
+% We are not doing control or tracking control. We're doing tracking.
 We present an online method of revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.
@@
 The Software Controller governs engine speed, braking, and steering.
-The controller takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
+It takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
 This short-term trajectory is oriented in the vehicle body frame.
@@
 We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call Trajectory Planner for each sub-path incrementally.
+
+Prescribed initial and end conditions for position, speed, and acceleration must be honored.
+All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
+This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
+We show how two dimensional \textbf{concerns????} (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
 
@@
 
-The task of Trajectory Planning is to take a 2 dimensional (horizontal plane) set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
+The task of Trajectory Planning is to take a one dimensional set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
 This trajectory must be time optimal yet still fall within constraints on speed, acceleration, and jerk.
@@
 Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
-
-Prescribed initial and end conditions for position, speed, and acceleration must be honored.
-All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
-This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
-We show how two dimensional \textbf{concerns????} (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
 
@@
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k$, $\mathbf{a}_{m,k}$, and $\mathbf{j}_{m,k}$ where $k = 1 ...
+A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1 ...
 N_s$ and can be solved individually in sequence.
@@
 For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$, where $N_p$ is the number of waypoints in the subpath.
-The value $\kappa_k$ denotes the path curvature at each point.
-For each sub-path point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the subpath.
+The value $\kappa_{k_P}$ denotes the path curvature at each point.
+For each sub-path point, the value $v_{m,k_p}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
 
@@
 However, the time at which the trajectory plan must be referenced is not always known.
-The solution to this problem is known as trajectory tracking control, and will be discussed in the next section.
+The solution to this problem is known as trajectory tracking, and will be discussed in the next section.
 
@@
 After planning has been performed, a reference signal must be regularly supplied to a control system in order for the vehicle to carry out the plan.
-This is known as tracking control.
+This is known as tracking .
 The primary difficulty here is that the trajectory plan has been developed in a one dimensional position space using a reference time which will have unknown drift relative to real time.

### 2016-01-23 root.tex
explain figure 7 and other minor changes

@@
 In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
-We explain how we handle intersections, turns in the road and planned stops due to stop signs (that have already been marked in the map).
+We explain how we handle intersections, turns in the road and planned stops due to stop signs.
 We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
@@
 This requires ... \textbf{something about computing actions through stop signs before it's necessary and how that's inefficient}.
-We propose to break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call Trajectory Planner for each sub-path incrementally.
+We break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call Trajectory Planner for each sub-path in the path.
 
@@
     Dividing the path into sub-paths at stop signs.
-    The trajectory generation is serially called on each sub-path which are colored differently here.
+    The Trajectory Planner is serially called for each sub-path shown in different color.
   }
@@
 Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
-Pedestrians that are on a different subpath than the vehicle are disregarded, as well as those which are behind the vehicle (i.e., $s_{ped} < s_{veh}$).
+Pedestrians that are on a different subpath than the vehicle, and behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
 All the remaining pedestrians pose potential collision risks, however only the closest pedestrian is considered for RSTOP.
 
-The distinction between momentarily braking for a pedestrian which quickly leaves the roadway and coming to a full stop needs no formal logic.
+The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no formal logic.
 They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
@@
 
-Our approach to fitting appropriate jerk profiles to complex sub-paths is to further break each subpath into segments which have very simple shape.
+Our approach to fitting appropriate jerk profiles to complex sub-paths is to further break each sub-path into segments which have very simple shape.
 Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
@@
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.
-A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1 ...
+A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1 ...
 N_s$ and can be solved individually in sequence.
@@
 \begin{enumerate} \label{asdf}
-  \item \emph{Ceiling calculation}, where 
-%$v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
-initial/final speeds, speed limit, peak acceleration and jerk 
-are found for every point in the given subpath.
-  \item \emph{Clustering}, in which subpath points are grouped sequentially and the segment boundaries are determined..
+  \item \emph{Ceiling calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
+%$initial/final speeds, speed limit, peak acceleration and jerk 
+are found for every point in the given sub-path.
+  \item \emph{Clustering}, in which sub-path points are grouped sequentially and the segment boundaries are determined..
   \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
 \end{enumerate}
-After this is done, the process described in Section \ref{sec:jerkprofiles} may be applied to each of the segments so that a plan of jerk vs. time may be generated.
+After this is done, the process described in next section may be applied to each of the segments so that a plan of jerk vs. time may be generated.
 
@@
 Each of the 9 scalars \textbf{RG: please explain 9 scalars ???} is a minimum of several maxima computed from an arbitrary set of constraints.
-For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the subpath.
+For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the sub-path.
 
@@
 For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the subpath.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
 The value $\kappa_{k_P}$ denotes the path curvature at each point.
@@
 Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
+
 Figure \ref{fig:1Dto2DandSegmentation} shows the result of this algorithm. 
+We start with a sub-path which is similar in structure to the sub-path between the stop signs in Fig.
+\ref{fig:subpathdivision}.
+This sub-path is broken up into five segments as shown in Fig.
+\ref{fig:1Dto2DandSegmentationa}, three shown in blue corresponding to straight line part of the sub-path and two shown in red corresponding to turns in the sub-path.
+As shown in Fig \ref{fig:1Dto2DandSegmentationb} we use the speed limit of 11.1 m/sec for the straight line segments and 3.0 m/sec for the turn segments to determine the segment boundaries.
 
@@
   \subfigure[
-    Result of dividing a single subpath into separate segments.
+    Result of dividing a single sub-path into separate segments.
     The clustering algorithm used here sets segments boundaries before and after curves.
-    The subpath segments in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.176 m/s (25 mi/hr).
+    The sub-path segments in red indicate areas where the longitudinal speed required to adhere to lateral acceleration limits is below the legal speed limit of 11.1 m/s (25 mi/hr).
   ]{
+  \label{fig:1Dto2DandSegmentationa}
     \includegraphics[width=0.8\columnwidth]{graphics/course_highlighted_turns.png}
@@
     The x-axis corresponds to distance along the subpath above.
-    The legal speed limit of 11.176 m/s (25 mi/hr) is shown in cyan.
+    The legal speed limit of 11.1 m/s (25 mi/hr) is shown in cyan.
     The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
@@
   ]{
+  \label{fig:1Dto2DandSegmentationb}
   \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
@@
 
-A jerk vs. time profile will be fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
+A jerk vs. time profile is fit to each segment using the algorithm in the next section.
 Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
-The values of $a_i$ and $v_i$ for the first segment are set to be the value which is currently reported from sensor data.
+The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
 Since the final acceleration is to be set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
@@
 Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments. 
-Continuity of position, speed, and acceleration is the primary goal of this process.
+Primary goal of this process is the continuity of position, speed, and acceleration.
 
@@
 The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
-Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
+Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$.
 % This is depicted in Fig. \ref{fig:pathtostop}
@@
 
-After planning has been performed, a reference signal must be regularly supplied to a control system in order for the vehicle to carry out the plan.
-This is known as tracking .
+After planning has been performed, a reference signal must be regularly supplied to the control system in order for the vehicle to carry out the plan.
+This is known as tracking.
 The primary difficulty here is that the trajectory plan has been developed in a one dimensional position space using a reference time which will have unknown drift relative to real time.
@@
 
-\textbf{RG: abrupt transition. need to summarize here}
-
-\textbf{RG: Ambarish suggests to replace all speed except speed limit by velocity and speeds -> velocities}
+To summarize, in the last three sections we have described how a urban path is broken into sub-paths,
+and computation of a velocity and jerk profile for segments within sub-paths that honor constraints
+for speed, velocity and acceleration as well as reactively slow down or stop for pedestrians on
+the urban path.
 

### 2016-01-23 root.tex
fixed bold comments

@@
     Primary system components. The {\it Path Manager} and {\it Trajectory Planner} components (in blue) are presented herein.
-   \textbf{RG: blue color is too dark to read text ????????????}
   }
@@
 This short-term trajectory is oriented in the vehicle body frame.
-\textbf{Citation for controller ????}.
 This work focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
@@
 Many systems perform trajectory planning over a sliding window within the full path which progresses with the vehicle.
-This requires ... \textbf{something about computing actions through stop signs before it's necessary and how that's inefficient}.
 We break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call Trajectory Planner for each sub-path in the path.
@@
 This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
-We show how two dimensional \textbf{concerns????} (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
+We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
 
@@
 The ceiling calculation is a minimax problem.
-Each of the 9 scalars \textbf{RG: please explain 9 scalars ???} is a minimum of several maxima computed from an arbitrary set of constraints.
+Each of the 9 scalars in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is a minimum of several maxima computed from an arbitrary set of constraints.
 For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the sub-path.
@@
 
-\textbf{RG: abrupt transition. please add a few lines on current sub-path being modified into 2 sub-paths. And there being a presence of a pedestrian}
-
-This is the procedure for determining nominal stopping distance required.
+The procedure for planning NORMAL mode trajectories is separate from that used to compute RSTOP trajectories.
 A 3-phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
-
 While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
@@
 Now appropriate values of $a_m^-$ and $j_m^-$ must be chosen to completely parameterize the trajectory.
-For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used.
+For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used and the distance traversed in the resultant trajectory is reported as the current value of $\Delta s_{RSTOP}$.
 
@@
   \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/PathProjection.png}
-  \caption{Showing projection of current position of car shown by green triangle to s coordinate system
-   \textbf{ RG: please label link in figure if possible} }
+  \includegraphics[width=0.8\columnwidth]{graphics/PathProjectionSlice.png}
+  \caption{Showing projection of current position of car shown by green triangle to s coordinate system.
+Each circle represents a node, and the lines between them are links.
+The orange dots indicate future trajectory samples which are sent to the controller.
+  }
 
@@
 
-\textbf{RG: this paragraph along with above figure should be merged with 2D to 1D s transformation section.
-in the figure please show a line for s coordinate as well} The general idea is to project the vehicle's current position onto the path, and then find the cumulative distance along the path between the start point and the projection point.
-This is the 1D location corresponding to the vehicle's current reference time.
+\textbf{RG: this paragraph along with above figure should be merged with 2D to 1D s transformation section. in the figure please show a line for s coordinate as well}
+
+\textbf{RC: Only the first two sentences pertain to 1D <-> 2D. it's a broad overview paragraph that needs to come before the following paragraphs.}
+
+The general idea is to project the vehicle's current position onto the path, and then find the cumulative distance along the path between the start point and the projection point.
+The 2D vehicle position is thus transformed into a 1D location corresponding to the vehicle's current reference time.
 Once this reference time is obtained, it may be used to look up position, speed, acceleration, and jerk along the curve of the path over the future time window.
@@
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
-Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$, F may be approximated \textbf {as cubic if a spline ???????
-} is being constructed from knots which have a very high resolution.
-In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:jerkprofiles}, and it is now resampled at $\Delta t = 0.01 sec$.
+Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$, may be approximated as cubic using a spline.
+To get the knots, $s(t)$ is sampled at $\Delta t = 0.01 sec$.
 This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.
-Evaluating the $t(s_{veh})$ then yields the current reference time.
+Evaluating $t(s_{veh})$ then yields the current reference time.
 

### 2016-01-23 root.tex
removed bf comment to move paragraph cuz nah man

@@
 
-\textbf{RG: this paragraph along with above figure should be merged with 2D to 1D s transformation section. in the figure please show a line for s coordinate as well}
-
-\textbf{RC: Only the first two sentences pertain to 1D <-> 2D. it's a broad overview paragraph that needs to come before the following paragraphs.}
-
 The general idea is to project the vehicle's current position onto the path, and then find the cumulative distance along the path between the start point and the projection point.

### 2016-01-23 root.tex
minor edits in trajectory section

@@
 
-The path manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
+The Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
 Each of the states is described below.
@@
 Prescribed initial and end conditions for position, speed, and acceleration must be honored.
-All planning is longitudinal; that is, speed is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
-This means that the desired trajectory plan may be mapped into one dimension using curvilinear coordinates to significantly simplify the solution.
+All planning is longitudinal; that is, speed profile generation is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
+This means that the desired trajectory plan may be mapped onto one dimension using curvilinear coordinates to significantly simplify the solution.
 We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
@@
 
-The task of Trajectory Planning is to take a one dimensional set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
-This trajectory must be time optimal yet still fall within constraints on speed, acceleration, and jerk.
+The task of Trajectory Planning is to take a 1D set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
+The resulting trajectory must be time optimal and satisfy all the constraints on speed, acceleration, and jerk.
 Elevation changes will not be addressed here.
@@
 
-Our approach to fitting appropriate jerk profiles to complex sub-paths is to further break each sub-path into segments which have very simple shape.
+In this section we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
+We first break each sub-path into segments which have very simple shape.
 Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
 Each segment will be defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
-Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.
+Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
 A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1 ...
@@
 For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation \textbf{ RG: using centrifugal  force relationship ????} $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
 The value $\kappa_{k_P}$ denotes the path curvature at each point.
@@
     Algorithm to set speed boundary conditions for interior segments.
-    Having consolidated the waypoints in to a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
+    Having consolidated the waypoints into a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
     This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
@@
 
-There are many cases in which this solution will fail.
-For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
-We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$., and our algorithm judiciously chooses the appropriate profile.
+We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
+These 4 cases cover cases in which the 7-phase solution fails.
+For instance, if $L$ is too short or $v_m$ is too high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
 The other profiles are explained below.
@@
 It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
-To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used.
+To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used \textbf{RG: can specify solver name???} .
 
@@
   \subfigure[6-phase profile] {
-  \includegraphics[width=0.4\columnwidth]{graphics/6phase_v(s).png}
+  \includegraphics[width=0.43\columnwidth]{graphics/6phase_v(s).png}
   \label{fig:6phaseprofile}
@@
   \subfigure[4-phase profile] {
-  \includegraphics[width=0.4\columnwidth]{graphics/4phase_v(s).png}
+  \includegraphics[width=0.43\columnwidth]{graphics/4phase_v(s).png}
   \label{fig:4phaseprofile}
@@
   \subfigure[Reversed 4-phase profile] {
-  \includegraphics[width=0.4\columnwidth]{graphics/4Rphase_all_derivatives.png}
+  \includegraphics[width=0.43\columnwidth]{graphics/4Rphase_all_derivatives.png}
   \label{fig:4rphaseprofile}
@@
   \subfigure[3-phase profile] {
-  \includegraphics[width=0.4\columnwidth]{graphics/3phase_v(s).png}
+  \includegraphics[width=0.43\columnwidth]{graphics/3phase_v(s).png}
   \label{fig:3phaseprofile}
@@
   \end{tabular}
-  \caption{6, 4, Reversed 4 and 3-phase profiles, displayed as speed vs. position along path}
+  \caption{6, 4, Reversed 4 and 3-phase profiles, displayed as speed vs. position along path.}
 
@@
 The procedure for planning NORMAL mode trajectories is separate from that used to compute RSTOP trajectories.
-A 3-phase profile will be used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+A 3-phase profile is used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
 While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.

### 2016-01-23 root.tex
final submit version from me

@@
 Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider reactive behaviors such as responding to pedestrians.
-On the other hand, systems for collision avoidance plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
+Other systems for collision avoidance plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
 In this paper, we present an integrated trajectory generation and tracking system for urban environments that simultaneously considers constraints to maximize human comfort and real-time trajectory updates to react to pedestrians on the road.
@@
 
-When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
+When planning trajectories for an autonomous ground vehicle operating on roadways, the typical goal is to minimize navigation time.
 % "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
@@
 Some preliminary work on optimizing these constraints on urban roads has also been done by ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
-On the other hand, many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
-This approach also makes the assumption that alternative paths around pedestrians are available, which may be invalid in urban scenarios.
+
+Many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
+These approaches also make the assumption that alternative paths around obstacles are available, which may be invalid in urban scenarios.
 Planners which optimize human comfort by modifying the whole trajectory to seek obstacle-free space \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
-In this paper, we describe a trajectory generation system that satisfies these constraints for human comfort and is reactive to presence of pedestrians on the road.
+In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to presence of pedestrians on the road.
 We enforce hard upper limits on jerk without requiring that travel time be known before planning occurs.
@@
 %Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
-At stop signs the car should stop and wait for an appropriate amount of time.
+At stop sign, the autonomous car should stop and wait for an appropriate amount of time.
 It should also react to pedestrians - at intersections as well as on the road.
-Pedestrians on the sidewalks should be ignored by the autonomous car.
 Further, the pedestrians can be static or moving.
+Pedestrians on the sidewalks should be ignored by the car.
 The car should stop, slow down and go around pedestrians when appropriate.
@@
 Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path without requiring prior knowledge of travel time.
-Our method first segments the path using an arbitrary set of kinematic or dynamic constraints which may be functions of the desired path.
+Our method first segments the path using a set of kinematic or dynamic constraints which may be functions of the desired path.
 Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
@@
 % We are not doing control or tracking control. We're doing tracking.
-We present an online method of revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.
+We present an online method for revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.
 
@@
 
-Our system architecture is shown in Figure \ref{fig:addreact}.
-There are 4 primary input components: Navigation Filter, Pedestrian Detector, a Path Planner, and a Map which provides the location of stop signs.
+Our system architecture is shown in Figure \ref{fig:addreact}. We have three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.
+
+The high-level planner is referred to as the {\it Path Manager}, and performs decision-making as well as governing operation of the low-level planner.
+Its primary functions are:
+\begin{enumerate}
+  \item Stop the vehicle at stop signs and wait until the associated intersection is clear before continuing.
+  \item Reactively reduce speed or stop altogether for pedestrians which pose a collision risk, then continue when they are clear.
+  \item Govern the Trajectory Planner.
+\end{enumerate}
+The low-level planner is referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
+Its primary functions are:
+\begin{itemize}
+  \item Compute medium-range trajectories for a given path.
+  \item Revise trajectories to react to pedestrians.
+  \item Sample the current trajectory for input to the controller.
+\end{itemize}
+
+The Software Controller governs engine speed, braking, and steering.
+It takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
+This short-term trajectory is oriented in the vehicle body frame.
+This work focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
+
+Beyond these major components, our system has 4 primary input components: Navigation Filter, Pedestrian Detector, a Path Planner, and a Map which provides the location of stop signs.
 % nav filter
@@
 We assume that the path planner operates independently from the trajectory planner and tracker.
-We also assume that we have a high resolution digital map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
 The destination is manually input into the system by the passengers.
@@
 
-Our High Resolution Map database is used to search for stop signs and provide their locations.
+We also assume that we have a High Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
+Our map is used to search for location of stop signs.
 The plan waypoints corresponding to stop signs are then flagged for later processing.
 We also have an In-Out Algorithm implemented via polygon intersection that can be used to ignore “safe” pedestrians
-Pedestrians that are on sidewalks, off-path crosswalks, and behind planned stops as shown in Figure \ref{fig:intersect} can be considered ``safe'' and are ignored.
-This is done by searching the same map database used for path planning.
+Pedestrians that are on sidewalks, behind planned stops, and off-path crosswalks as shown in Figure \ref{fig:intersect} can be considered {\it safe} and are ignored.
+%This is done by searching the same map database used for path planning.
 Remaining pedestrians pose collision risk and the vehicle needs to reactively slow down or stop for them.
-This is accomplished by sending stop message with pedestrian locations to the path manager.
-
-Beyond these input subsystems, our system has three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.
-
-The high-level planner is referred to as the {\it Path Manager}, and performs decision-making as well as governing operation of the low-level planner.
-Its primary functions are:
-\begin{enumerate}
-  \item Stop the vehicle at stop signs and wait until the associated intersection is clear before continuing.
-  \item Reactively reduce speed or stop altogether for pedestrians which pose a collision risk, then continue when they are clear.
-  \item Govern the Trajectory Planner.
-\end{enumerate}
-The low-level planner is referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
-Its primary functions are:
-\begin{itemize}
-  \item Compute medium-range trajectories for a given path.
-  \item Revise trajectories to react to pedestrians.
-  \item Sample the current trajectory for input to the controller.
-\end{itemize}
-
-The Software Controller governs engine speed, braking, and steering.
-It takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
-This short-term trajectory is oriented in the vehicle body frame.
-This work focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
+%This is accomplished by sending stop message with pedestrian locations to the path manager.
 
@@
 \section{Path Manager} \label{sec:pathmanager}
-
-The Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
-Each of the states is described below.
-
-\begin{figure}[thpb]
-  \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
-  \caption{State Transition Diagram for ?????????? more detail????? }
-  \label{fig:st}
-\end{figure}
 
@@
 This means that the desired trajectory plan may be mapped onto one dimension using curvilinear coordinates to significantly simplify the solution.
-We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
+%We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
+We express the 2D positions of the car (specifically cornering locations and speeds) in 1D $s$ coordinate system as described in Sec. \ref{sec:trajectorytracking} to guide planning.
+
+The Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
+Each of the states are described below.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
+  \caption{State Transition Diagram for Path Manager. System transitions from NORMAL to PSTOP mode at each stop sign. In the presence of a pedestrian the system transitions to RSTOP.}
+  \label{fig:st}
+\end{figure}
 
 In the absence of pedestrians, the path manager cycles between NORMAL and PSTOP states.
-In NORMAL state, the path manager send the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
+In NORMAL state, the path manager sends the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
 In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
@@
 
-All detected pedestrian locations are compared against the high-resolution digital map, so that pedestrians that are not on the roadway or near an upcoming crosswalk are not considered.
-Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
-Pedestrians that are on a different subpath than the vehicle, and behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
+All detected pedestrian locations are compared against the high-resolution digital map, so that pedestrians that are off roadway are not considered.
+Then each pedestrian location is projected onto the 1D path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
+Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
 All the remaining pedestrians pose potential collision risks, however only the closest pedestrian is considered for RSTOP.
 
-The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no formal logic.
+The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no separate logic.
 They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
@@
 
-\begin{figure}[thpb]
+\begin{figure}[tb]
 \centering
@@
 When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
-The procedure for doing this is the same as the procedure for planning the RSTOP trajectory later with a few steps omitted, since the planning algorithm is extremely efficient.
+The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is extremely efficient.
 
@@
 The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
-The preceding quantities are depicted graphically in Fig. \ref{fig:react}.
+These quantities are depicted graphically in Fig. \ref{fig:react}.
 
@@
 The resulting trajectory must be time optimal and satisfy all the constraints on speed, acceleration, and jerk.
-Elevation changes will not be addressed here.
+%Elevation changes will not be addressed here.
 Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
@@
 In this section we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
-We first break each sub-path into segments which have very simple shape.
+We first break each sub-path into segments with simple shape.
 Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
-Each segment will be defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
+Each segment is defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
-A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1 ...
-N_s$ and can be solved individually in sequence.
+A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1, ..., N_s$ and can be solved individually in sequence.
 
@@
 \begin{enumerate} \label{asdf}
-  \item \emph{Ceiling calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
+  \item \emph{Ceiling Calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
 %$initial/final speeds, speed limit, peak acceleration and jerk 
@@
 \end{enumerate}
-After this is done, the process described in next section may be applied to each of the segments so that a plan of jerk vs. time may be generated.
+After these steps are completed, the process for fitting jerk profiles described in next section may be applied to each of the segments so that a plan of jerk vs. time may be generated.
 
-The ceiling calculation is a minimax problem.
+The {\it Ceiling Calculation} is a minimax problem.
 Each of the 9 scalars in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is a minimum of several maxima computed from an arbitrary set of constraints.
@@
 For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation \textbf{ RG: using centrifugal  force relationship ????} $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation 
+%\textbf{ RG: using centrifugal  force relationship ????} 
+$v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
 The value $\kappa_{k_P}$ denotes the path curvature at each point.
@@
 
-This information is used to choose the segment boundaries in the clustering step.
-Any number of clustering algorithms may be used here, and prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
-For the present example, we will choose a simple univariate heuristic which essentially identifies curves in the roadway using only the speed ceiling.
+This information is used to choose the segment boundaries in the {\it Clustering} step.
+Any number of clustering algorithms may be used here, and a prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
+We choose a simple univariate heuristic which essentially identifies curves in the roadway using only the speed ceiling.
 Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
@@
 
-Figure \ref{fig:1Dto2DandSegmentation} shows the result of this algorithm. 
+Figure \ref{fig:1Dto2DandSegmentation} shows the result of this Clustering algorithm. 
 We start with a sub-path which is similar in structure to the sub-path between the stop signs in Fig.
@@
 %initial/final speeds, speed limit, peak acceleration and jerk 
-must be calculated for each segment using the values corresponding to each of the waypoints which comprise the segment in question.
+must be calculated for each segment using the values corresponding to the waypoints comprising the segment.
 The minimum of scalar across all constituent points is chosen.
@@
 
-A jerk vs. time profile is fit to each segment using the algorithm in the next section.
+A jerk vs. time profile is fit to each segment using the algorithm described in the next section.
 Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
@@
 To introduce the process of computing a jerk profile for each segment, let us consider the case in which a vehicle needs to travel some arbitrary distance ahead.
-The vehicle is moving at some small initial speed and needs to accelerate to some traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+The vehicle is moving at an initial speed and needs to accelerate to a traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
 Figure \ref{fig:full7phasespec} shows this trajectory over time.
 
-\begin{figure}[thpb]
+\begin{figure}[tb]
   \centering
@@
 We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
-These 4 cases cover cases in which the 7-phase solution fails.
+These 4 profiles cover cases in which the 7-phase solution fails.
 For instance, if $L$ is too short or $v_m$ is too high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
-The other profiles are explained below.
-
+%The other profiles are explained below.
 All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
 It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
-To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used \textbf{RG: can specify solver name???} .
-
-For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
-This results in a negative time interval for one or more phases.
-In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
-Specifically, jerk may be increased while peak acceleration is decreased.
-This process is referred to herein as ``accel-jerk tuning''
-If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.
-
-To decide which profile to employ, a cascade approach is used.
-Analytic values of time intervals are evaluated for each profile by plugging in new values of $\mathbf{a}_m$ and $\mathbf{j}_m$ until all time intervals are positive and real.
-The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
-Each is described below.
+To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver is used 
+%\textbf{RG: can specify solver name???} .
+These 4 profiles are:
 
@@
 
-\begin{figure}[thpb]
+\begin{figure}[tb]
   \centering
@@
 % \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}
+For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
+This results in a negative time interval for one or more phases.
+In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
+Specifically, jerk may be increased while peak acceleration is decreased.
+This process is referred to herein as ``accel-jerk tuning''.
+If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.
+
+To decide which profile to employ, a cascade approach is used.
+Analytic values of time intervals are evaluated for each profile by plugging in new values of $\mathbf{a}_m$ and $\mathbf{j}_m$ until all time intervals are positive and real.
+The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
+Each is described below.
 
@@
   \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
-  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values
+  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values.
 \end{itemize}
@@
 The procedure for planning NORMAL mode trajectories is separate from that used to compute RSTOP trajectories.
-A 3-phase profile is used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
-While it would seem that the segment solving procedure described in Sec. \ref{sec:3phase} could be applied to reactive stopping, this is not the case.
+For RSTOP trajectory, a 3-phase profile is used, with end conditions $v_f=0, a_f=0$ and initial conditions set to the current state of the navigation filter.
+While it would seem that the 3-phase segment solving procedure described in Sec. 
+%\ref{sec:3phase} 
+\ref{sec:jerkprofiles} can be applied to reactive stopping, this is not the case.
 As was mentioned previously, only one of $[L, v_f]$ may be rigidly enforced for a 3-phase profile, but not both, since no closed-form solution exists for that case.
@@
 However, reactive stopping requires that $v_f$ be enforced, and that we solve for $L$.
-Reformulating a new 3-phase jerk profile accordingly is a simple matter for a symbolic solver.
+Reformulating a new 3-phase jerk profile accordingly is a simple matter for the symbolic solver.
 %%% This is the brief way of wording the preceding three sentences.
@@
 % This is depicted in Fig. \ref{fig:pathtostop}
-The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.
+The vehicle now begins sampling from this trajectory and executes a smooth reactive stop.
 
@@
 After planning has been performed, a reference signal must be regularly supplied to the control system in order for the vehicle to carry out the plan.
-This is known as tracking.
-The primary difficulty here is that the trajectory plan has been developed in a one dimensional position space using a reference time which will have unknown drift relative to real time.
+%This is known as tracking.
+The primary difficulty here is that the trajectory plan has been developed in a 1D position space using a reference time which has an unknown drift relative to real time.
 As such, we must find a way to look up the coordinates of the vehicle within the reference trajectory using information typically available for autonomous cars.
-A minimal set of such information would include position, velocity vector, heading, and acceleration in the body frame.
-After this, it is assumed that the vehicle controller will need samples of a small window into the future trajectory to pilot the vehicle.
+A minimal set of such information includes position, velocity vector, heading, and acceleration in the body frame.
+It is assumed that the vehicle controller will need samples of a small window into the future trajectory to pilot the vehicle.
 
@@
 \ref{fig:cartos}).
-Projections which lie outside of the link's two constituent path points are snapped to that which is the nearest.
+Projections which lie outside of the link's two constituent path points are snapped to the nearest one.
 The projection which is closest to the vehicle's actual position is then assumed to be the correct path projection $\mathbf{r}_p$.
@@
 
-Choosing an algorithm to obtain the current one dimensional position of the vehicle along the length of the path $s_{veh}$ (i.e., in the same one-dimensional coordinate system in which the planning was performed) is influenced by the form of the path plan.
-Assuming that the path is linear between waypoints reduces computational effort significantly.
+Choice of an algorithm to obtain the current 1D position of the vehicle along the length of the path $s_{veh}$ (i.e., in the same 1D $s$-coordinate system in which the planning was performed) is influenced by the form of the path plan.
+We assume that the path is linear between waypoints to significantly reduce the computational effort.
 So $s_{veh}$ at any epoch is the sum of the link lengths behind its perpendicular projection $\mathbf{r}_p$.
@@
 for speed, velocity and acceleration as well as reactively slow down or stop for pedestrians on
-the urban path.
+the urban path. We then explained how our trajectory plan is regularly supplied in real-time to the
+autonomous car controller.
 
@@
 We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
-they leave. Our system handles complex cases that were not explicitly programmed, gracefully, such as following 
+they leave. Our system gracefully handles complex cases that were not explicitly programmed, such as following 
 pedestrians if they continue walking on the road away from the car. Our system works with multiple 

### 2016-03-16 root.tex
update tex file with lots of corrections

@@
 In this paper, we address the problem of trajectory planning for urban roads while reacting to the presence of pedestrians.
-Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider reactive behaviors such as responding to pedestrians.
-Other systems for collision avoidance plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
-In this paper, we present an integrated trajectory generation and tracking system for urban environments that simultaneously considers constraints to maximize human comfort and real-time trajectory updates to react to pedestrians on the road.
+Past work limits jerk, velocities and acceleration for smooth trajectories, but does not consider reactive behaviors such as responding to pedestrians.
+Other systems for collision avoidance stop or plan paths around obstacles and pedestrians.
+In this paper, we present an integrated trajectory generation and tracking system for urban environments.
+Our system simultaneously considers constraints for smooth trajectory and updates the trajectory in real-time to react to pedestrians on the road.
 
-Given a desired path, we present a novel online method for planning longitudinal trajectories to follow the urban path with turns and intersections while honoring traffic regulations such as stop signs.
-Based on sensor data, our method updates the trajectory plan to safely avoid pedestrians on the road by slowing down or stopping.
-This is critical in urban environments, when modifying the vehicle's intended path is often not possible.
+We present a novel online method for planning longitudinal trajectories to follow an urban path with turns and intersections while honoring traffic regulations such as stop signs. 
+Our method updates the trajectory to safely avoid pedestrians on the road by slowing down or stopping.
+This is critical in urban environments, where modifying the vehicle's intended path is often not possible.
 Our method has closed form solutions, runs at 20 Hz, and is efficient and reliable for use in online planning.
@@
 
-When planning trajectories for an autonomous ground vehicle operating on roadways, the typical goal is to minimize navigation time.
+When planning trajectories for an autonomous ground vehicle operating on roadways, the typical goal is minimization of navigation time.
 % "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
-It is common to employ the path-velocity decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
+It is common to employ the path-trajectory decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
 First is {\it Path Planning}, which involves computation of the optimal path.
-The second task is {\it Trajectory Planning}, which involves the computation of the fastest velocity profile that is appropriate for the scenario.
+The second task is {\it Trajectory Planning}, which involves the computation of a velocity profile appropriate for the scenario.
 
-Computation of a trajectory plan for traveling from one point to another is often guided by
-passenger comfort. The set of possible paths to the destination is typically restricted 
+The set of possible paths to the destination is typically restricted
 by constraints such as traversability and legally available travel lanes, among others.
+Trajectory computation for traveling from one point to another is often guided by passenger comfort. 
 The intended speed, acceleration, and higher derivatives of position with respect to time are also subject to constraints.
@@
 A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
-Similar work has been performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013}, which sets the travel time a priori and optimally minimizes jerk but does not set hard limitations or control jerk directly.
-Some preliminary work on optimizing these constraints on urban roads has also been done by ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
+Similar work has been performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013}, which sets the travel time a priori and optimally minimizes without hard limitations or direct jerk control.
+Other work on optimizing these constraints on urban roads ~\cite{Rastelli14,Li15} doesn't consider reactive behaviors such as responding to pedestrians or other dynamic objects in the scene.
 
-Many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
-These approaches also make the assumption that alternative paths around obstacles are available, which may be invalid in urban scenarios.
-Planners which optimize human comfort by modifying the whole trajectory to seek obstacle-free space \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
+Many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that navigates urban roads.
+These approaches also make the assumption that alternative paths around obstacles are available, which may not always be valid.
+Planners that optimize human comfort by modifying the path seek obstacle-free space \cite{Villagra2012,Villagra2012a} and do not feature stopping as a well-integrated behavior.
 In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to presence of pedestrians on the road.
@@
 
-Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path without requiring prior knowledge of travel time.
-Our method first segments the path using a set of kinematic or dynamic constraints which may be functions of the desired path.
+We present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along the desired path without requiring prior knowledge of travel time.
+Our method segments the path using a set of kinematic or dynamic constraints.
 Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
-The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
+The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and within limits throughout the entire path.
 The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
@@
 In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
-We explain how we handle intersections, turns in the road and planned stops due to stop signs.
+We explain how we pass through intersections, turns in the road and planned stops due to stop signs.
 We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
@@
 
-The high-level planner is referred to as the {\it Path Manager}, and performs decision-making as well as governing operation of the low-level planner.
+The High Level Planner is referred to as the {\it Path Manager}, and performs decision-making as well as governs operation of the low-level planner.
 Its primary functions are:
@@
 \end{enumerate}
-The low-level planner is referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
+The Low Level Planner is also referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
 Its primary functions are:
@@
 Pedestrian detection is performed via a camera and LiDAR system based on \cite{Gepperth2013,Gepperth2014}.
-It computes a bounding box for each pedestrian as shown in Figure \ref{fig:ped}.
+The pedestrian detector outputs a bounding box for each pedestrian as shown in Figure \ref{fig:ped}.
 The pedestrian position is computed in the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
@@
 
-We assume that we have access to a path planner.
+We assume access to a path planner.
 For example, this can be a service similar to Google Maps except that instead of road level route, it provides lane level route. 
@@
   \caption{
-    A car turning right at the intersection only needs to consider pedestrians in the pedestrian crossings shown in yellow.
+    A car, with the blue line path, turning right at the intersection only needs to consider pedestrians in the zebra crossings shown in yellow.
     Pedestrians at the green crossings can be ignored unless they appear somewhere on the road.
@@
 
-We also assume that we have a High Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
+We also assume a High Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
 Our map is used to search for location of stop signs.
@@
 
-Many systems perform trajectory planning over a sliding window within the full path which progresses with the vehicle.
-We break the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call Trajectory Planner for each sub-path in the path.
+Path Manager performs trajectory planning over a sliding window within the full path which progresses with the vehicle.
+It breaks the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call the Trajectory Planner for each sub-path in the path.
 
 Prescribed initial and end conditions for position, speed, and acceleration must be honored.
-All planning is longitudinal; that is, speed profile generation is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
+All planning is longitudinal; that is, speed profile generation is the primary focus. We assume that the control subsystem ensures that the vehicle follows the prescribed path.
 This means that the desired trajectory plan may be mapped onto one dimension using curvilinear coordinates to significantly simplify the solution.
@@
 
-The Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
+Our Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
 Each of the states are described below.
@@
 In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
-When pedestrians are present, the transition from PSTOP to NORMAL is delayed until there are no pedestrians detected in the range denoted by $\Delta s_{resume}$.
+When pedestrians are present, the transition from PSTOP to NORMAL is delayed until there are no pedestrians detected in the range denoted by $\Delta s_{resume}$ as shown in Fig. \ref{fig:react}(c).
 
@@
 
-All detected pedestrian locations are compared against the high-resolution digital map, so that pedestrians that are off roadway are not considered.
-Then each pedestrian location is projected onto the 1D path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
+All detected pedestrian locations are compared against the high-resolution digital map. Pedestrians that are off roadway are not considered.
+Each pedestrian location is projected onto the 1D path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
 Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
@@
 They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
-It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.
+It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occur in the pedestrian detection algorithm.
 
@@
 When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
-The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is extremely efficient.
+???????????? The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is quite efficient. ?????????? not clear
 
 Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
-RSTOP is triggered when $\Delta s_{RSTOP} + \Delta s_{buf} >= \Delta s_{avail}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
+RSTOP is triggered when $\Delta s_{avail} <= \Delta s_{RSTOP} + \Delta s_{buf}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
 Planning and executing the trajectory for coming to a halt is detailed in Sec. \ref{sec:reactivestoptrajectory}.
@@
 We first break each sub-path into segments with simple shape.
-Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
+Then each segment can be compared to a library of possible jerk profiles to find a good match.
 Each segment is defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
-A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1, ..., N_s$ and can be solved individually in sequence.
+A sub-path is then a set of $N$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k}$, $\mathbf{a}_{m,k}$, and $\mathbf{j}_{m,k}$ where $k = 1, ..., N$ and can be solved individually in sequence.
 
@@
 \end{enumerate}
-After these steps are completed, the process for fitting jerk profiles described in next section may be applied to each of the segments so that a plan of jerk vs. time may be generated.
+After these steps are completed, the process for fitting jerk profiles described in next sub-section is applied to each of the segments to generate a plan of jerk vs. time.
 
@@
 %\textbf{ RG: using centrifugal  force relationship ????} 
-$v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
+$v_{m,w}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{w}}$ for $w = 1, ..., W$, where $W$ is the number of waypoints in the sub-path.
 The value $\kappa_{k_P}$ denotes the path curvature at each point.
-For each sub-path point, the value $v_{m,k_p}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+For each sub-path point, the value $v_{m,w}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
 
 This information is used to choose the segment boundaries in the {\it Clustering} step.
-Any number of clustering algorithms may be used here, and a prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
-We choose a simple univariate heuristic which essentially identifies curves in the roadway using only the speed ceiling.
-Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
-Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
+Any clustering algorithm may be used here, and a prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
+We choose a simple univariate heuristic which identifies curves in the roadway using only the speed ceiling.
+Each time $v_{m,w} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
+Each time $v_{m,w} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
 
@@
 The minimum of scalar across all constituent points is chosen.
-While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
+While not time optimal, it ensures that no constraints are violated while still keeping the number of total segments low.
 In the present example, this means that in the straight portions of the subpath $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
@@
 
-A jerk vs. time profile is fit to each segment using the algorithm described in the next section.
-Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
+A jerk vs. time profile is fit to each segment using the algorithm described in the next sub-section.
+Speed and acceleration must be continuous between each segment. Only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
 The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
@@
 For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
-Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments. 
-Primary goal of this process is the continuity of position, speed, and acceleration.
+Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments,
+with continuity of position, speed, and acceleration.
 
@@
       \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
-      \For{$i = 1, ..., N_s-1$}
+      \For{$i = 1, ..., N-1$}
         \If{$v_{m,i+1} > v_{m,i}$}
@@
     This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
+    ????????? define $v_p$ ???????????
   }
@@
 
-To summarize, in the last three sections we have described how a urban path is broken into sub-paths,
-and computation of a velocity and jerk profile for segments within sub-paths that honor constraints
+To summarize, in the last three sections we described how a urban path is broken into sub-paths. We 
+explained the computation of velocity and jerk profile for segments within sub-paths that honor constraints
 for speed, velocity and acceleration as well as reactively slow down or stop for pedestrians on
-the urban path. We then explained how our trajectory plan is regularly supplied in real-time to the
+the urban path. We then described how our trajectory plan is supplied regularly in real-time to the
 autonomous car controller.

### 2016-03-24 root.tex
improving english

@@
   % Reactive Path Planning with Pedestrains for Autonomous Driving
-  Reactive Trajectory Planning and Trajectory Tracking for Pedestrian-Aware Autonomous Driving in Urban Environments
+  Reactive Trajectory Planning and Tracking for Pedestrian-Aware Autonomous Driving in Urban Environments
 }
@@
 \begin{abstract}
-In this paper, we address the problem of trajectory planning for urban roads while reacting to the presence of pedestrians.
-Past work limits jerk, velocities and acceleration for smooth trajectories, but does not consider reactive behaviors such as responding to pedestrians.
-Other systems for collision avoidance stop or plan paths around obstacles and pedestrians.
+In this paper, we address the problem of trajectory planning for an autonomous car for urban roads while reacting to the presence of pedestrians.
+Past work limits jerk, velocities and acceleration for smooth trajectories, without considering reactive behaviors such as responding to pedestrians.
+Other systems based on collision avoidance, plan paths around obstacles and pedestrians in unstructured environments.
 In this paper, we present an integrated trajectory generation and tracking system for urban environments.
@@
 
-We present a novel online method for planning longitudinal trajectories to follow an urban path with turns and intersections while honoring traffic regulations such as stop signs. 
-Our method updates the trajectory to safely avoid pedestrians on the road by slowing down or stopping.
-This is critical in urban environments, where modifying the vehicle's intended path is often not possible.
+We present a novel online method for planning %longitudinal
+trajectories to follow a given urban path, while honoring traffic regulations such as stop signs at intersections. 
+We update the trajectory to safely avoid pedestrians on the road by slowing down or stopping.
 Our method has closed form solutions, runs at 20 Hz, and is efficient and reliable for use in online planning.
@@
 
-When planning trajectories for an autonomous ground vehicle operating on roadways, the typical goal is minimization of navigation time.
+%When planning trajectories for 
+An autonomous ground vehicle operating on roadways, typically plans for trajectories with the goal of minimizing navigation time.
 % "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
-It is common to employ the path-trajectory decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
+It is common, to employ the path-velocity decomposition to separate the twin tasks of Path and Trajectory Planning, so that they can be run sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-First is {\it Path Planning}, which involves computation of the optimal path.
-The second task is {\it Trajectory Planning}, which involves the computation of a velocity profile appropriate for the scenario.
+Here {\it Path Planning} involves computation of the optimal path and {\it Trajectory Planning} involves the computation of a velocity profile appropriate for the scenario.
 
 The set of possible paths to the destination is typically restricted
-by constraints such as traversability and legally available travel lanes, among others.
-Trajectory computation for traveling from one point to another is often guided by passenger comfort. 
-The intended speed, acceleration, and higher derivatives of position with respect to time are also subject to constraints.
-These include legal speed limits, engine dynamics, desired side-slip limits, traction availability, and rollover concerns.
+by constraints such as traversability and legally available travel lanes.
+Trajectory computation for traveling on the path is often guided by passenger comfort. 
+The intended speed, acceleration, and time derivatives of position are also subject to constraints.
+These include legal speed limits, engine dynamics, desired side-slip limits, traction availability, and rollover risk.
 
-For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort.
-A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
-Similar work has been performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013}, which sets the travel time a priori and optimally minimizes without hard limitations or direct jerk control.
-Other work on optimizing these constraints on urban roads ~\cite{Rastelli14,Li15} doesn't consider reactive behaviors such as responding to pedestrians or other dynamic objects in the scene.
+For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration and velocity values are prime contributors to ride discomfort.
+Lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
+Similar work performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013} sets the travel time a priori and optimally minimizes ???what?????
+without hard limitations or direct jerk control.
+Other work on optimizing these constraints on urban roads ~\cite{Rastelli14,Li15} doesn't consider reactive behaviors such as responding to pedestrians and other dynamic objects in the scene.
 
-Many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that navigates urban roads.
-These approaches also make the assumption that alternative paths around obstacles are available, which may not always be valid.
-Planners that optimize human comfort by modifying the path seek obstacle-free space \cite{Villagra2012,Villagra2012a} and do not feature stopping as a well-integrated behavior.
+Many systems in the literature plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13}.
+However, these systems do not work as part of an integrated system that navigates urban roads.
+These systems also make the assumption that alternative paths around obstacles are available. Such an assumption may not always be valid in an urban scenario.
+Planners that optimize human comfort by modifying the path to go around obstacles \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
 In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to presence of pedestrians on the road.
-We enforce hard upper limits on jerk without requiring that travel time be known before planning occurs.
+We enforce hard upper limits on jerk without requiring that travel time be known for planning.
 %Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
@@
 We present an autonomy framework for pedestrian-aware ground vehicles.
-Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
+We would like our autonomous car to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
 %Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
-At stop sign, the autonomous car should stop and wait for an appropriate amount of time.
-It should also react to pedestrians - at intersections as well as on the road.
+At stop signs, our autonomous car should stop and wait for an appropriate amount of time.
+The car should also stop, slow down for pedestrians when appropriate.
+It should react to pedestrians at the intersections as well as jay walking pedestrians.
 Further, the pedestrians can be static or moving.
 Pedestrians on the sidewalks should be ignored by the car.
-The car should stop, slow down and go around pedestrians when appropriate.
 
-We present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along the desired path without requiring prior knowledge of travel time.
+We present a novel method for planning longitudinal trajectories 
+%(the time derivatives of position in the body-forward direction) 
+along a desired path without requiring prior knowledge of travel time.
 Our method segments the path using a set of kinematic or dynamic constraints.
 Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
-The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and within limits throughout the entire path.
-The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
+We intelligently choose a jerk profile from a set of pre-solved profiles such that acceleration is continuous and within limits throughout the entire path.
+The resultant trajectory plan is then sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
+In urban environments, modifying the vehicle's intended path is often not possible.
 % We are not doing control or tracking control. We're doing tracking.
-We present an online method for revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.
+We revise the trajectory plan online to safely avoid pedestrians by slowing down or stopping. % when no alternative paths are available.
 
@@
 exhibit reactive behavior and stop for pedestrians that may appear on the road.
-In Section \ref{sec:trajectoryplanning}, we describe trajectory planning which takes a 2D path and generates a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
-We explain how we pass through intersections, turns in the road and planned stops due to stop signs.
+In Section \ref{sec:trajectoryplanning}, we describe trajectory planning. Starting from a 2D path, we generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
+We explain how we pass through intersections, turns in the road and planned stops at the stop signs.
 We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
-Once a trajectory plan is formulated, it must be utilized online to send the reference signal to the controllers that govern steering, braking, and engine speed.
+Once the trajectory plan is formulated, it is utilized online. We send the reference signal to the controllers that govern steering, braking, and engine speed.
 This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
@@
 The Software Controller governs engine speed, braking, and steering.
-It takes as an input a trajectory which is sampled at 10 Hz and spans 2 seconds, beginning at the moment of data transmission.
+It takes as an input a trajectory sampled at 10 Hz spanning 2 seconds, beginning at the moment of data transmission.
 This short-term trajectory is oriented in the vehicle body frame.
-This work focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
+This paper focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
 
-Beyond these major components, our system has 4 primary input components: Navigation Filter, Pedestrian Detector, a Path Planner, and a Map which provides the location of stop signs.
+Beyond these major components, our system has four primary input components: Navigation Filtering, Pedestrian Detection, a Path Planner, and a Map.
 % nav filter
-For navigation filtering we use an ADMA-G  commercial automotive GPS/INS system.
+For Navigation Filtering we use an ADMA-G  commercial automotive GPS/INS system.
 It receives RTK corrections via cell modem.
@@
 
-Pedestrian detection is performed via a camera and LiDAR system based on \cite{Gepperth2013,Gepperth2014}.
+Pedestrian Detection is performed via a camera and LiDAR system based on \cite{Gepperth2013,Gepperth2014}.
 The pedestrian detector outputs a bounding box for each pedestrian as shown in Figure \ref{fig:ped}.
@@
 
-We assume access to a path planner.
-For example, this can be a service similar to Google Maps except that instead of road level route, it provides lane level route. 
+We assume access to a lane level Path Planner. For example, this can be a service similar to Google Maps except that it provides a lane level route instead of a road level route. 
 We assume that the path planner operates independently from the trajectory planner and tracker.
-The destination is manually input into the system by the passengers.
-Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the destination.
+%The destination is manually input into the system by the passengers.
+Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the user specified destination.
 This plan is then output as a set of global waypoints.
@@
 We also assume a High Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
-Our map is used to search for location of stop signs.
+Our map provides the location of stop signs.
 The plan waypoints corresponding to stop signs are then flagged for later processing.
@@
 %This is done by searching the same map database used for path planning.
-Remaining pedestrians pose collision risk and the vehicle needs to reactively slow down or stop for them.
+Remaining pedestrians pose collision risk and our vehicle needs to reactively slow down or stop for them.
 %This is accomplished by sending stop message with pedestrian locations to the path manager.
@@
 
-Path Manager performs trajectory planning over a sliding window within the full path which progresses with the vehicle.
-It breaks the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and call the Trajectory Planner for each sub-path in the path.
+Path Manager performs trajectory planning over a sliding window within the full path with the sliding window moving with the vehicle.
+It breaks the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and calls the Trajectory Planner for each sub-path in the path.
 
@@
 Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
-All the remaining pedestrians pose potential collision risks, however only the closest pedestrian is considered for RSTOP.
+All the remaining pedestrians pose potential collision risks. However only the closest pedestrian is considered for RSTOP.
 

### 2016-03-29 root.tex
improve english
more english

@@
 We present a novel online method for planning %longitudinal
-trajectories to follow a given urban path, while honoring traffic regulations such as stop signs at intersections. 
+trajectories to follow a given urban path while honoring traffic regulations such as stop signs at intersections. 
 We update the trajectory to safely avoid pedestrians on the road by slowing down or stopping.
@@
 Planners that optimize human comfort by modifying the path to go around obstacles \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
-In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to presence of pedestrians on the road.
-We enforce hard upper limits on jerk without requiring that travel time be known for planning.
+In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to the presence of pedestrians on the road.
+We enforce hard upper limits on jerk without requiring that travel time is known for planning.
 %Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
@@
 %(the time derivatives of position in the body-forward direction) 
-along a desired path without requiring prior knowledge of travel time.
+along the desired path without requiring prior knowledge of travel time.
 Our method segments the path using a set of kinematic or dynamic constraints.
@@
 We also describe other subsystems that our system needs to work with, and planned stops.
-Section \ref{sec:pathmanager} describes the high level Path Manager that calls the trajectory planner as necessary to 
+Section \ref{sec:pathmanager} describes the high-level Path Manager that calls the trajectory planner as necessary to 
 exhibit reactive behavior and stop for pedestrians that may appear on the road.
@@
 
-The High Level Planner is referred to as the {\it Path Manager}, and performs decision-making as well as governs operation of the low-level planner.
+The High-Level Planner is referred to as the {\it Path Manager}, and performs decision-making as well as governs operation of the low-level planner.
 Its primary functions are:
@@
 \end{enumerate}
-The Low Level Planner is also referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
+The Low-Level Planner is also referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
 Its primary functions are:
@@
 % nav filter
-For Navigation Filtering we use an ADMA-G  commercial automotive GPS/INS system.
+For Navigation Filtering, we use an ADMA-G  commercial automotive GPS/INS system.
 It receives RTK corrections via cell modem.
@@
 The pedestrian detector outputs a bounding box for each pedestrian as shown in Figure \ref{fig:ped}.
-The pedestrian position is computed in the vehicle body frame, and subsequently transformed to global frame to be related to the planned path.
+The pedestrian position is computed in the vehicle body frame, and subsequently transformed to the global frame to be related to the planned path.
 
@@
 
-We also assume a High Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
+We also assume a High-Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
 Our map provides the location of stop signs.
@@
 
-Prescribed initial and end conditions for position, speed, and acceleration must be honored.
+Prescribed initial and end conditions for the position, speed, and acceleration must be honored.
 All planning is longitudinal; that is, speed profile generation is the primary focus. We assume that the control subsystem ensures that the vehicle follows the prescribed path.
@@
 Our Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
-Each of the states are described below.
+Each of the states is described below.
 
@@
 Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
-All the remaining pedestrians pose potential collision risks. However only the closest pedestrian is considered for RSTOP.
+All the remaining pedestrians pose potential collision risks. However, only the closest pedestrian is considered for RSTOP.
 
@@
 When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
-???????????? The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is quite efficient. ?????????? not clear
+??????????? The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is quite efficient. ?????????? not clear
 
@@
 %Elevation changes will not be addressed here.
-Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
+Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, a speed which is piecewise quadratic over time, and position which is piecewise cubic over time.
 
@@
 
-In this section we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
-We first break each sub-path into segments with simple shape.
+In this section, we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
+We first break each sub-path into segments with a simple shape.
 Then each segment can be compared to a library of possible jerk profiles to find a good match.
@@
 are found for every point in the given sub-path.
-  \item \emph{Clustering}, in which sub-path points are grouped sequentially and the segment boundaries are determined..
+  \item \emph{Clustering}, in which sub-path points are grouped sequentially and the segment boundaries are determined.
   \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
@@
   \subfigure[
-    Result of dividing a single sub-path into separate segments.
+    The result of dividing a single sub-path into separate segments.
     The clustering algorithm used here sets segments boundaries before and after curves.
@@
   \subfigure[
-    Example of how the sub-path's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
+    An example of how the sub-path's segment boundaries are determined, as well as how $v_m$ is consolidated to determine the speed ceiling for each segment.
     The x-axis corresponds to distance along the subpath above.
@@
   \caption{
-    7-phase constant jerk profile along curvilinear distance axis.
+    7-phase constant jerk profile along the curvilinear distance axis.
     All values needed to completely parameterize a single path segment for trajectory are labeled here.}
@@
 The middle travel period $\Delta t_4$ is removed, so that the vehicle increases speed and then immediately decreases speed. 
-The peak speed in this case has a solution, but cannot be manipulated directly.
+The peak speed, in this case, has a solution that cannot be manipulated directly.
 Without this known value, the problem becomes under-defined, so $\Delta t_2$ and $\Delta t_5$ each have two solutions: one negative and one positive.
@@
 
-Once the path manager triggers RSTOP, the trajectory is then recomputed with increasingly higher-magnitude values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
+Once the path manager triggers RSTOP, the trajectory is then recomputed with increasingly higher absolute values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
 Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
@@
   \includegraphics[width=0.8\columnwidth]{graphics/PathProjectionSlice.png}
-  \caption{Showing projection of current position of car shown by green triangle to s coordinate system.
+  \caption{Showing projection of the current position of the car shown by green triangle to s coordinate system.
 Each circle represents a node, and the lines between them are links.
@@
 
-Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity, and acceleration vectors) onto the path.
+Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity and acceleration vectors) onto the path.
 A simple approximation is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'', and is depicted in Fig.
@@
 
-To summarize, in the last three sections we described how a urban path is broken into sub-paths. We 
+To summarize, in the last three sections we described how an urban path is broken into sub-paths. We 
 explained the computation of velocity and jerk profile for segments within sub-paths that honor constraints
@@
 the Path Manager.
-We can travel along desired path, stop at stop signs, stop for pedestrians in roadway, and continue after 
+We can travel along a desired path, stop at stop signs, stop for pedestrians in the roadway, and continue after 
 they leave. Our system gracefully handles complex cases that were not explicitly programmed, such as following 
@@
 to the presence of pedestrians.
-Second, our trajectory planner has closed form solutions, and is suitable for online implementation
-on a car, unlike systems requiring solutions to non linear optimizations. Finally, we have confirmed the
+Second, our trajectory planner has closed form solution and is suitable for online implementation
+on a car, unlike systems requiring solutions to nonlinear optimizations. Finally, we have confirmed the
 efficiency and reliability of our trajectory planner on a test vehicle with over 100 hours of testing
@@
 
-In future, we would like to handle other roadway traffic, 
-coexisting with other vehicles. We would also like to perceive and stop for stop signs and red-lights,
+In future, we would like to handle other roadway traffic, perceiving and reacting to other vehicles. We would also like to perceive and stop for stop signs and red-lights,
 rather than using map information.

### 2016-04-09 root.tex
add slides, update some figures
remove some lines

@@
 
-We present an autonomy framework for pedestrian-aware ground vehicles.
 We would like our autonomous car to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
@@
 
-We present a novel method for planning longitudinal trajectories 
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=0.7\columnwidth]{graphics/its_car.png}
+  \caption{
+    Showing car used in our experiments.
+  }
+  \label{fig:addreact}
+\end{figure}
+
+We present an autonomy framework for pedestrian-aware ground vehicles
+with a novel method for planning longitudinal trajectories 
 %(the time derivatives of position in the body-forward direction) 
-along the desired path without requiring prior knowledge of travel time.
+along the desired path. %without requiring prior knowledge of travel time.
 Our method segments the path using a set of kinematic or dynamic constraints.
@@
 We intelligently choose a jerk profile from a set of pre-solved profiles such that acceleration is continuous and within limits throughout the entire path.
-The resultant trajectory plan is then sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
+The resultant trajectory plan is then sampled to provide a reference for real-time vehicle control.
+%, which is called trajectory tracking.
 In urban environments, modifying the vehicle's intended path is often not possible.
 % We are not doing control or tracking control. We're doing tracking.
-We revise the trajectory plan online to safely avoid pedestrians by slowing down or stopping. % when no alternative paths are available.
+%We revise the trajectory plan online to safely avoid pedestrians by slowing down or stopping. % when no alternative paths are available.
 
-We first describe the architecture of our system in Sec. \ref{sec:systemarchirecture}.
-We also describe other subsystems that our system needs to work with, and planned stops.
+Section \ref{sec:systemarchirecture} describes the architecture of our system and 
+other subsystems that interact with our system.
 Section \ref{sec:pathmanager} describes the high-level Path Manager that calls the trajectory planner as necessary to 
 exhibit reactive behavior and stop for pedestrians that may appear on the road.
-In Section \ref{sec:trajectoryplanning}, we describe trajectory planning. Starting from a 2D path, we generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
-We explain how we pass through intersections, turns in the road and planned stops at the stop signs.
+Section \ref{sec:trajectoryplanning} describe trajectory planning, where we start with a 2D path, and generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
+%We explain how we pass through intersections, turns in the road and planned stops at the stop signs.
 We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
-Once the trajectory plan is formulated, it is utilized online. We send the reference signal to the controllers that govern steering, braking, and engine speed.
-This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking}.
-We then conclude and discuss future extensions.
+Once the trajectory plan is formulated, we send the reference signal to the controllers that govern steering, braking, and engine speed.
+This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking},
+followed by conclusions and future extensions.
 % tracking
@@
   \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
+  \includegraphics[width=0.6\columnwidth]{graphics/stgraph.PNG}
   \caption{State Transition Diagram for Path Manager. System transitions from NORMAL to PSTOP mode at each stop sign. In the presence of a pedestrian the system transitions to RSTOP.}

### 2016-04-09 root.tex
remove some lines

@@
 exhibit reactive behavior and stop for pedestrians that may appear on the road.
-Section \ref{sec:trajectoryplanning} describe trajectory planning, where we start with a 2D path, and generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
+Section \ref{sec:trajectoryplanning} describes trajectory planning, where we generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
 %We explain how we pass through intersections, turns in the road and planned stops at the stop signs.
-We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
+%We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
 Once the trajectory plan is formulated, we send the reference signal to the controllers that govern steering, braking, and engine speed.
-This is the trajectory tracking problem, which is discussed in Section \ref{sec:trajectorytracking},
+This is the trajectory tracking and is discussed in Section \ref{sec:trajectorytracking},
 followed by conclusions and future extensions.

### 2016-04-14 root.tex
removed pdf, and made some reviewer changes

@@
   }
-  \label{fig:addreact}
+  \label{fig:car}
 \end{figure}
@@
 The pedestrian position is computed in the vehicle body frame, and subsequently transformed to the global frame to be related to the planned path.
+Pedestrian detection algorithms are still in heavy development.
+In order to make the algorithm robust to detection errors, it is designed under the assumption that pedestrian trajectory estimates and predictions are unavailable.
+Furthermore, it is assumed that no unique identifier is available to differentiate one pedestrian from the other in a single epoch, or to correlate the same pedestrian between epochs.
 
@@
 
-Path Manager performs trajectory planning over a sliding window within the full path with the sliding window moving with the vehicle.
-It breaks the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and calls the Trajectory Planner for each sub-path in the path.
+The Path Manager performs trajectory planning over a sliding window within the full path with the sliding window moving with the vehicle.
+As shown in Fig. \ref{fig:addreact}, a Path Planner passes paths into it as a 2D linestring.
+It then breaks the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and calls the Trajectory Planner for each sub-path in the path.
 
@@
 In the case where $\Delta t_2$ or $\Delta t_6$ are negative, accel-jerk tuning must be employed.
+The 7-phase profile can be expressed as $( j_1 > 0 , j_2 = 0 , j_3 < 0 , j_4 = 0 , j_5 < 0 , j_6 = 0 , j_7 > 0 )$.
 In the instance where $\Delta t_4$ is negative, the total segment length $L$ is too short to reach top speed of $v_m$, so a 6-phase profile must be used.
@@
 In the case where both solutions are negative, accel-jerk tuning is attempted. 
+The 6-phase profile can be expressed as $( j_1 > 0 , j_2 = 0 , j_3 < 0 , j_4 < 0 , j_5 = 0 , j_6 > 0 )$
 Should that yield no positive time intervals, one of the two 4-phase profiles then becomes appropriate, since $L$ is not large enough to accommodate accelerating past the end speed.
@@
 When $v_i < v_m \simeq v_f$, the vehicle should increase speed, then maintain that speed for the distance remaining in the segment.
+The 4-phase profile can be expressed as $( j_1 > 0, j_2 = 0 , j_3 < 0, j_4 = 0 )$
 
@@
 The initial speed is maintained for a period before applying the brakes.
+The reversed 4-phase profile can be expressed as $( j_1 = 0, j_2 < 0, j_3 = 0 , j_4 > 0 )$
 
@@
 When these discrepancies in the feasible $v_f$ for a segment arise, the corresponding value in $v_p$ is updated so that the new conditions may be accounted for in the subsequent segment.
+The 3-phase profile can be expressed as $(j_1 \pm 0 , j_2 = 0, j_3 \mp 0 )$
 
@@
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
-Each is described below.
+Each has been described above.
+The routines for fitting a profile to a given $\mathbf{b}, \mathbf{a}_m, \mathbf{j}_m$ to a segment using this cascade approach is referred to as a ``jerk library''.
 
@@
 in an autonomous car to follow an urban path with online updates to avoid pedestrians on the roadway by slowing down or 
-stopping. We built a robust stopping model as part of Trajectory Planner reliably integrated with 
-the Path Manager.
+stopping.
+We built a robust stopping model as part of Trajectory Planner reliably integrated with the Path Manager.
 We can travel along a desired path, stop at stop signs, stop for pedestrians in the roadway, and continue after 
@@
 pedestrians as we only need to consider the closest pedestrian ahead of us.
+
+\textbf{emphasize limited knowledge of pedestrians as a strength}
 

### 2016-04-14 root.tex
more reviewer comment changes

@@
 When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
-??????????? The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is quite efficient. ?????????? not clear
+This serves as a dynamic buffer between the pedestrian and the vehicle.
+In order to calculate $\Delta s_{RSTOP}$, a hypothetical trajectory plan is created (but not implemented).
+A 3-phase profile using the vehicle's current state as initial conditions and $v_f = 0$ is created, and the resultant distance then becomes the nominal required stopping distance.
 
@@
 These quantities are depicted graphically in Fig. \ref{fig:react}.
-
+Nominal parameter values are: $\Delta s_{buf} = 8.5 m$, $\Delta s_{replan} = 1.0 m$, $\Delta s_{resume} = 12.5 m$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
+For this work, they are set constant for all segments.
+However, this functionality enables behaviors such as more sedate driving in a school zone or when traversing small areas with lower friction coefficients, such as patches of ice or gravel detected via camera.
 A sub-path is then a set of $N$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k}$, $\mathbf{a}_{m,k}$, and $\mathbf{j}_{m,k}$ where $k = 1, ..., N$ and can be solved individually in sequence.
@@
 For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the sub-path.
+Any number of constraint sets may be used to compute the ceiling.
 
@@
 %\textbf{ RG: using centrifugal  force relationship ????} 
-$v_{m,w}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{w}}$ for $w = 1, ..., W$, where $W$ is the number of waypoints in the sub-path.
+$v_{m,w}(a_{B,y}) = \sqrt{ | a_{B,y}^{max} / \kappa_{w} | }$ for $w = 1, ..., W$, where $W$ is the number of waypoints in the sub-path.
 The value $\kappa_{k_P}$ denotes the path curvature at each point.

### 2016-04-14 maps_ITSPlanner.cpp

@@ -1,0 +1,593 @@
+#include "maps_ITSPlanner.h"	// Includes the header of this component
+
+
+MAPS_BEGIN_INPUTS_DEFINITION(MAPSITSPlanner)
+
+  /// - it is encouraged that this input be run through the maps_PathPreprocessor 
+  ///   block first. (or before DLCManager)
+  /// - format:
+  ///   [x1,y1,v1,x2,y2,v2,...,xn,yn,vn]
+  ///   v could really be anything, but as of v0.1, only 0/-1 is supported.
+  ///     -1: unspecified speed
+  ///     0: stop sign
+  /// the path that the vehicle will follow, including stop signs
+  MAPS_INPUT("master_path",  MAPS::FilterFloat64, MAPS::FifoReader)
+
+  /// The kinematic state of the vehicle (at the rear axle)
+  /// [x,y,theta,vBx,aBx]
+  /// x: x-position in HRI linear coordinates (m)
+  /// y: y-position in HRI linear coordinates (m)
+  /// theta: heading in ISO convention (rad)
+  /// vBx: forward speed of the vehicle (body x axis) (m/s)
+  /// aBx: forward acceleration of the vehicle (body x axis) (m/s^2)
+  MAPS_INPUT("state",        MAPS::FilterFloat64, MAPS::LastOrNextReader)
+
+  /// pedestrians on the roadway
+  /// - this can come from either the hsun pedestrian detection block, or from the 
+  ///   pedestrian detection simulator block in this same package.
+  /// - The only reliable piece of information we have available is position, so that's
+  ///   the only piece of information from the RealObject that we use.
+  /// - will stop for these pedestrians at the appropriate juncture, unless they 
+  ///   disappear first
+  /// - an empty vector of pedestrians indicates that none are detected. It is
+  ///   critical that an empty real object vector be published here when there
+  ///   are no pedestrians.
+  MAPS_INPUT("rstop_peds", MAPS::FilterRealObjects, MAPS::LastOrNextReader)
+
+MAPS_END_INPUTS_DEFINITION
+
+
+MAPS_BEGIN_OUTPUTS_DEFINITION(MAPSITSPlanner)
+
+  /// This is the 1D trajectory simulated at time step specified by "sim_time_step"
+  /// it is compressed into a double vector.
+  /// There are 5 fields: t, s, v, a, j. Each field has the same number of points
+  /// The format is as follows:
[…]

### 2016-04-14 root.tex
fixed question mark places

@@
 Lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
-Similar work performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013} sets the travel time a priori and optimally minimizes ???what?????
-without hard limitations or direct jerk control.
+Similar work performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013} sets the travel time a priori and optimally minimizes jerk without hard limitations or direct jerk control.
 Other work on optimizing these constraints on urban roads ~\cite{Rastelli14,Li15} doesn't consider reactive behaviors such as responding to pedestrians and other dynamic objects in the scene.
@@
     This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
-    ????????? define $v_p$ ???????????
   }

### 2016-04-16 root.tex
removed ref to pedestrians in conclusion
heavy changes to path segmentation section

@@
 
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=0.3\columnwidth]{graphics/IntersectionCrosswalks.png}
-  \caption{
-    A car, with the blue line path, turning right at the intersection only needs to consider pedestrians in the zebra crossings shown in yellow.
-    Pedestrians at the green crossings can be ignored unless they appear somewhere on the road.
-  }
-  \label{fig:intersect}
-\end{figure}
+% \begin{figure}[tb]
+%   \centering
+%   \includegraphics[width=0.3\columnwidth]{graphics/IntersectionCrosswalks.png}
+%   \caption{
+%     A car, with the blue line path, turning right at the intersection only needs to consider pedestrians in the zebra crossings shown in yellow.
+%     Pedestrians at the green crossings can be ignored unless they appear somewhere on the road.
+%   }
+%   \label{fig:intersect}
+% \end{figure}
 
@@
 All planning is longitudinal; that is, speed profile generation is the primary focus. We assume that the control subsystem ensures that the vehicle follows the prescribed path.
-This means that the desired trajectory plan may be mapped onto one dimension using curvilinear coordinates to significantly simplify the solution.
-%We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
-We express the 2D positions of the car (specifically cornering locations and speeds) in 1D $s$ coordinate system as described in Sec. \ref{sec:trajectorytracking} to guide planning.
+This allows all planning to be done in a Frenet (rather than Cartesian) frame for simplicity.
+The Frenet axis which runs along the length of the planned path is denoted by $s$, and all planning is performed in the $s$-axis.
+Differences along this axis ($\Delta s$) correspond to distances the car will actually travel.
+Mapping back to Cartesian coordinates (necessary for trajectory tracking) is well-discussed in the literature.
 
@@
 All detected pedestrian locations are compared against the high-resolution digital map. Pedestrians that are off roadway are not considered.
-Each pedestrian location is projected onto the 1D path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
+Each pedestrian location is perpendicularly projected onto the path to determine how much travel distance separates them from the car.
 Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
@@
 
-The task of Trajectory Planning is to take a 1D set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
-The resulting trajectory must be time optimal and satisfy all the constraints on speed, acceleration, and jerk.
+The task of Trajectory Planning is to take a set of scalar waypoints in the $s$-frame from the Path Manager, and compute a series of constant jerk intervals for longitudinal motion along the $s$-axis
+The resulting trajectory should minimize time as much as possible, and satisfy all the constraints on speed, acceleration, and jerk.
 %Elevation changes will not be addressed here.
@@
 In this section, we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
-We first break each sub-path into segments with a simple shape.
+We first break each sub-path into segments where the constraint values can be considered constant.
 Then each segment can be compared to a library of possible jerk profiles to find a good match.
 Each segment is defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
-Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
+Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ are defined for each segment.
 For this work, they are set constant for all segments.
@@
 \end{enumerate}
-After these steps are completed, the process for fitting jerk profiles described in next sub-section is applied to each of the segments to generate a plan of jerk vs. time.
+% After these steps are completed, the process for fitting jerk profiles described in next sub-section is applied to each of the segments to generate a plan of jerk vs. time.
 
@@
 For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation 
-%\textbf{ RG: using centrifugal  force relationship ????} 
-$v_{m,w}(a_{B,y}) = \sqrt{ | a_{B,y}^{max} / \kappa_{w} | }$ for $w = 1, ..., W$, where $W$ is the number of waypoints in the sub-path.
-The value $\kappa_{k_P}$ denotes the path curvature at each point.
-For each sub-path point, the value $v_{m,w}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation :
+\begin{equation}
+  v_{m,w}(a_{B,y}) = \sqrt{ | a_{B,y}^{max} / \kappa_{w} | } ~,~ w = 1, ..., W
+\end{equation}
+where $W$ is the number of waypoints in the sub-path, and $\kappa_{k_P}$ denotes the path curvature at each point.
+For each sub-path point, the value $v_{m,w}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration $a^{max}_{B,y}$.
 
@@
 Any clustering algorithm may be used here, and a prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
-We choose a simple univariate heuristic which identifies curves in the roadway using only the speed ceiling.
+We choose a simple univariate heuristic which effectively identifies curves in the roadway using only the speed ceiling.
 Each time $v_{m,w} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
@@
 Figure \ref{fig:1Dto2DandSegmentation} shows the result of this Clustering algorithm. 
-We start with a sub-path which is similar in structure to the sub-path between the stop signs in Fig.
-\ref{fig:subpathdivision}.
-This sub-path is broken up into five segments as shown in Fig.
-\ref{fig:1Dto2DandSegmentationa}, three shown in blue corresponding to straight line part of the sub-path and two shown in red corresponding to turns in the sub-path.
-As shown in Fig \ref{fig:1Dto2DandSegmentationb} we use the speed limit of 11.1 m/sec for the straight line segments and 3.0 m/sec for the turn segments to determine the segment boundaries.
+We start with a sub-path which is similar in structure to the green sub-path in Fig. \ref{fig:subpathdivision}.
+This sub-path is broken up into five segments as shown in Fig. \ref{fig:1Dto2DandSegmentationa}.
+The three segments shown in blue correspond to more straight portions, and the two shown in red correspond to more curved portions.
 
@@
     The longitudinal speed required to adhere to lateral acceleration limits is shown in red.
-    The resultant speed ceiling is shown in blue.
+    The resultant speed ceiling ($v_m$) is shown in blue.
   ]{
@@
 
-% need better notation here.
 In the consolidation step, a single set of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
-%initial/final speeds, speed limit, peak acceleration and jerk 
 must be calculated for each segment using the values corresponding to the waypoints comprising the segment.
-The minimum of scalar across all constituent points is chosen.
-While not time optimal, it ensures that no constraints are violated while still keeping the number of total segments low.
-In the present example, this means that in the straight portions of the subpath $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
-As a result, many of the curves will have a constant speed.
+The principle task is to find a value of $v_m$ for each segment using the values from segmentation, $SL$ and $v_{m,w}(a_{B,y}^{max})$. 
+The value of the speed ceiling for segment $k$ is chosen to be:
+\begin{equation}
+  v_{m,k} = min \left ( \mathbf{SL}_k ,  \mathbf{v}_{m}(a_{B,y}^{max}) |_k \right )
+\end{equation}
+where $\mathbf{SL}_k$ is the vector containing the values of the legal speed limit at every waypoint in segment $k$.
+The vector $\mathbf{v}_{m}(a_{B,y}^{max}) |_k$ contains the values of the speed limit from lateral acceleration at every waypoint in segment $k$.
+% The minimum of scalar across all constituent points is chosen.
+While not time optimal, choosing the minimum of these values over the entire segment ensures that no constraints are violated, and as a result, many of the curves will have a constant speed.
 For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
+Once the values $v_{m,k}$ are set for every segment in the subpath, boundary conditions must be set to ensure that speed is continuous between segments.
+Acceleration continuity will be addressed later.
+The algorithm in Fig. \ref{alg:segmentspeedboundaryconditions} is used to compute the vector $\mathbf{v}_p$ and then each segment $k$ can be solved using:
+\begin{equation}
+  v_{i,k} = v_{f,k-1} = v_{p,k}
+\end{equation}
 
-A jerk vs. time profile is fit to each segment using the algorithm described in the next sub-section.
-Speed and acceleration must be continuous between each segment. Only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
 The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
@@
 For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
-Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments,
-with continuity of position, speed, and acceleration.
 
@@
 Now that all segments within the subpath have been parameterized, they may be solved sequentially.
+A jerk vs. time profile is fit to each segment using the algorithm described in the next sub-section.
 
@@
 % After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
-The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
+The path must now be truncated for mapping back to Cartesian coordinates.
 Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$.
@@
 %This is known as tracking.
-The primary difficulty here is that the trajectory plan has been developed in a 1D position space using a reference time which has an unknown drift relative to real time.
+The primary difficulty here is that the trajectory plan has been developed in the $s$-frame using a reference time which has an unknown drift relative to real time.
 As such, we must find a way to look up the coordinates of the vehicle within the reference trajectory using information typically available for autonomous cars.
@@
 The general idea is to project the vehicle's current position onto the path, and then find the cumulative distance along the path between the start point and the projection point.
-The 2D vehicle position is thus transformed into a 1D location corresponding to the vehicle's current reference time.
+The Cartesian vehicle position is thus mapped onto the path at a Frenet position which corresponds to the vehicle's current reference time.
 Once this reference time is obtained, it may be used to look up position, speed, acceleration, and jerk along the curve of the path over the future time window.
@@
 
-Choice of an algorithm to obtain the current 1D position of the vehicle along the length of the path $s_{veh}$ (i.e., in the same 1D $s$-coordinate system in which the planning was performed) is influenced by the form of the path plan.
+Choice of an algorithm to obtain the current position of the vehicle along the length of the path $s_{veh}$ is influenced by the form of the path plan.
 We assume that the path is linear between waypoints to significantly reduce the computational effort.
@@
 
-\textbf{emphasize limited knowledge of pedestrians as a strength}
-
 We make three contributions in this paper. First, we present an integrated trajectory planner that

### 2016-04-19 root.tex
many edits, make figures smaller

@@
 \usepackage{subfigure}
+\usepackage{amssymb}
 \usepackage[noadjust]{cite}
@@
 \begin{abstract}
-In this paper, we address the problem of trajectory planning for an autonomous car for urban roads while reacting to the presence of pedestrians.
+In this paper, we address the problem of trajectory planning for an autonomous car for urban roads while simultaneously reacting to the presence of pedestrians.
 Past work limits jerk, velocities and acceleration for smooth trajectories, without considering reactive behaviors such as responding to pedestrians.
 Other systems based on collision avoidance, plan paths around obstacles and pedestrians in unstructured environments.
-In this paper, we present an integrated trajectory generation and tracking system for urban environments.
+In this paper, we present an integrated trajectory generation and tracking system. % for urban environments. 
 Our system simultaneously considers constraints for smooth trajectory and updates the trajectory in real-time to react to pedestrians on the road.
@@
 % "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
-It is common, to employ the path-velocity decomposition to separate the twin tasks of Path and Trajectory Planning, so that they can be run sequentially or in parallel.
+It is common to employ the path-velocity decomposition to separate the twin tasks of Path and Trajectory Planning, so that these tasks can be run sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-Here {\it Path Planning} involves computation of the optimal path and {\it Trajectory Planning} involves the computation of a velocity profile appropriate for the scenario.
+{\it Path Planning} involves computation of the optimal path and {\it Trajectory Planning} involves the computation of a velocity profile appropriate for the scenario.
 
-The set of possible paths to the destination is typically restricted
-by constraints such as traversability and legally available travel lanes.
 Trajectory computation for traveling on the path is often guided by passenger comfort. 
-The intended speed, acceleration, and time derivatives of position are also subject to constraints.
-These include legal speed limits, engine dynamics, desired side-slip limits, traction availability, and rollover risk.
-
 For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration and velocity values are prime contributors to ride discomfort.
+The intended speed, acceleration, and time derivatives of position are also subject to constraints
+due to legal speed limits, engine dynamics, desired side-slip limits, traction availability, and rollover risk.
 Lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
@@
 
+The set of possible paths to the destination is typically restricted
+by constraints such as available travel lanes, and obstacles such as pedestrians.
 Many systems in the literature plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13}. 
-However, these systems do not work as part of an integrated system that navigates urban roads.
-These systems also make the assumption that alternative paths around obstacles are available. Such an assumption may not always be valid in an urban scenario.
+%However, these systems do not work as part of an integrated system that navigates urban roads.
+These systems make an assumption that alternative paths around obstacles are available. Such an assumption may not always be valid in an urban scenario.
 Planners that optimize human comfort by modifying the path to go around obstacles \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
 In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to the presence of pedestrians on the road.
-We enforce hard upper limits on jerk without requiring that travel time is known for planning.
+We enforce hard upper limits on jerk without requiring that travel time be known for planning.
 %Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
@@
 At stop signs, our autonomous car should stop and wait for an appropriate amount of time.
-The car should also stop, slow down for pedestrians when appropriate.
-It should react to pedestrians at the intersections as well as jay walking pedestrians.
+The car should also stop, slow down for pedestrians %when appropriate.It should react to pedestrians 
+at the intersections as well as jay walking pedestrians.
 Further, the pedestrians can be static or moving.
@@
   \centering
-  \includegraphics[width=0.7\columnwidth]{graphics/its_car.png}
+  \includegraphics[width=0.6\columnwidth]{graphics/its_car.png}
   \caption{
@@
 
-Section \ref{sec:systemarchirecture} describes the architecture of our system and 
-other subsystems that interact with our system.
-Section \ref{sec:pathmanager} describes the high-level Path Manager that calls the trajectory planner as necessary to 
-exhibit reactive behavior and stop for pedestrians that may appear on the road.
-Section \ref{sec:trajectoryplanning} describes trajectory planning, where we generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
+Section \ref{sec:systemarchirecture} describes the architecture of our system and interacting subsystems.
+Section \ref{sec:pathmanager} describes the high-level Path Manager that calls the trajectory planner to 
+exhibit reactive behavior and stop for pedestrians on the road.
+Section \ref{sec:trajectoryplanning} describes trajectory planning to generate a trajectory that satisfies maximum velocity, acceleration and jerk constraints. 
 %We explain how we pass through intersections, turns in the road and planned stops at the stop signs.
@@
   \caption{
-    Primary system components. The {\it Path Manager} and {\it Trajectory Planner} components (in blue) are presented herein.
+    Primary system components. The {\it Path Manager} and {\it Trajectory Planner} components (in green) are presented herein.
   }
@@
 The pedestrian position is computed in the vehicle body frame, and subsequently transformed to the global frame to be related to the planned path.
-Pedestrian detection algorithms are still in heavy development.
-In order to make the algorithm robust to detection errors, it is designed under the assumption that pedestrian trajectory estimates and predictions are unavailable.
-Furthermore, it is assumed that no unique identifier is available to differentiate one pedestrian from the other in a single epoch, or to correlate the same pedestrian between epochs.
+%Pedestrian detection algorithms are still in heavy development.
+%In order to make the algorithm robust to detection errors, it is designed under the assumption that pedestrian trajectory estimates and predictions are unavailable. Furthermore, 
+It is assumed that no unique identifier is available to differentiate one pedestrian from the other in a single epoch, or to correlate the same pedestrian between epochs.
 
@@
 We also assume a High-Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
-Our map provides the location of stop signs.
-The plan waypoints corresponding to stop signs are then flagged for later processing.
+This map provides the location of stop signs.
+%The plan waypoints corresponding to stop signs are then flagged for later processing.
 We also have an In-Out Algorithm implemented via polygon intersection that can be used to ignore “safe” pedestrians
-Pedestrians that are on sidewalks, behind planned stops, and off-path crosswalks as shown in Figure \ref{fig:intersect} can be considered {\it safe} and are ignored.
+Pedestrians that are on sidewalks, behind planned stops, and off-path crosswalks.
+% as shown in Figure \ref{fig:intersect} can be considered {\it safe} and are ignored.
 %This is done by searching the same map database used for path planning.
-Remaining pedestrians pose collision risk and our vehicle needs to reactively slow down or stop for them.
+%Remaining pedestrians pose collision risk and our vehicle needs to reactively slow down or stop for them.
 %This is accomplished by sending stop message with pedestrian locations to the path manager.
@@
 The Path Manager performs trajectory planning over a sliding window within the full path with the sliding window moving with the vehicle.
-As shown in Fig. \ref{fig:addreact}, a Path Planner passes paths into it as a 2D linestring.
-It then breaks the path into sub-paths at each stop location which is known a priori, as shown in Fig. \ref{fig:subpathdivision}, and calls the Trajectory Planner for each sub-path in the path.
+%As shown in Fig. \ref{fig:addreact}, a Path Planner passes paths into it as a 2D linestring.
+It breaks the path planner provided path into sub-paths at each stop location which are known a priori, as shown in Fig.
+\ref{fig:subpathdivision}.
+It then calls the Trajectory Planner for each sub-path.
 
@@
   \centering
-  \includegraphics[width=0.6\columnwidth]{graphics/stgraph.PNG}
+  \includegraphics[width=0.9\columnwidth]{graphics/state_transition.PNG}
   \caption{State Transition Diagram for Path Manager. System transitions from NORMAL to PSTOP mode at each stop sign. In the presence of a pedestrian the system transitions to RSTOP.}
@@
 In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
-When pedestrians are present, the transition from PSTOP to NORMAL is delayed until there are no pedestrians detected in the range denoted by $\Delta s_{resume}$ as shown in Fig. \ref{fig:react}(c).
+When pedestrians are present, the transition from PSTOP to NORMAL is delayed until no pedestrians are detected in the range denoted by $\Delta s_{resume}$ as shown in Fig. \ref{fig:react}(c).
 
@@
 All detected pedestrian locations are compared against the high-resolution digital map. Pedestrians that are off roadway are not considered.
-Each pedestrian location is perpendicularly projected onto the path to determine how much travel distance separates them from the car.
+Each pedestrian location is perpendicularly projected onto the path to determine the travel distance separating them from the car.
 Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
@@
 They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
-It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occur in the pedestrian detection algorithm.
+It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure no tracking errors in the pedestrian detection algorithm.
 
@@
 \centering
-  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
+  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=0.8\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
   \\
-  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Replan.png}}
+  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=0.9\columnwidth]{graphics/RSTOP_Replan.png}}
   \\
-  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
+  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=0.7\columnwidth]{graphics/RSTOP_Resume.png}}
   \caption{Graphical representation of static and dynamic thresholds used to trigger state transitions.}
@@
 Note that a static buffer distance, $\Delta s_{buf}$ is added to ensure safety and comfort.
-RSTOP is triggered when $\Delta s_{avail} <= \Delta s_{RSTOP} + \Delta s_{buf}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
+RSTOP is triggered when $\Delta s_{avail} \leqslant \Delta s_{RSTOP} + \Delta s_{buf}$ and the actual available stopping distance $\Delta s_{avail} = s_{ped} - s_{veh}$ is computed.
 Planning and executing the trajectory for coming to a halt is detailed in Sec. \ref{sec:reactivestoptrajectory}.
@@
 
-The task of Trajectory Planning is to take a set of scalar waypoints in the $s$-frame from the Path Manager, and compute a series of constant jerk intervals for longitudinal motion along the $s$-axis
-The resulting trajectory should minimize time as much as possible, and satisfy all the constraints on speed, acceleration, and jerk.
+The task of Trajectory Planning is to take a set of scalar waypoints in the frenet frame from the Path Manager, and compute a series of constant jerk intervals for longitudinal motion.% along the $s$-axis
+The resulting trajectory should minimize time, and satisfy all the constraints on speed, acceleration, and jerk.
 %Elevation changes will not be addressed here.
@@
 Then each segment can be compared to a library of possible jerk profiles to find a good match.
-Each segment is defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
+Each segment is defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, with parameters initial speed, initial acceleration, total distance, speed limit, and the final speed, respectively.
 Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ are defined for each segment.
@@
 For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the sub-path.
-Any number of constraint sets may be used to compute the ceiling.
+%Any number of constraint sets may be used to compute the ceiling.
 
 % How we actually did this.
-For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
+For the purposes of developing a simple example application, two constraints are used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
 Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation :
@@
 \end{equation}
-where $W$ is the number of waypoints in the sub-path, and $\kappa_{k_P}$ denotes the path curvature at each point.
+where $W$ is the number of waypoints in the sub-path, and $\kappa_{w}$ denotes the path curvature at each point.
 For each sub-path point, the value $v_{m,w}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration $a^{max}_{B,y}$.
@@
 This information is used to choose the segment boundaries in the {\it Clustering} step.
-Any clustering algorithm may be used here, and a prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
+%Any clustering algorithm may be used here, and 
+A prudent choice of segment boundaries is arguably a critical piece of the trajectory planning process.
 We choose a simple univariate heuristic which effectively identifies curves in the roadway using only the speed ceiling.
-Each time $v_{m,w} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
-Each time $v_{m,w} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.
+Each time $v_{m,w} < SL$ becomes true, a curve is begun, and a segment boundary is drawn.
+Each time $v_{m,w} = SL$ is again true, the current curve is ended, and another segment boundary is drawn.
 
 Figure \ref{fig:1Dto2DandSegmentation} shows the result of this Clustering algorithm. 
-We start with a sub-path which is similar in structure to the green sub-path in Fig. \ref{fig:subpathdivision}.
+We start with a sub-path similar in structure to the green sub-path in Fig. \ref{fig:subpathdivision}.
 This sub-path is broken up into five segments as shown in Fig. \ref{fig:1Dto2DandSegmentationa}.
@@
 \end{equation}
-where $\mathbf{SL}_k$ is the vector containing the values of the legal speed limit at every waypoint in segment $k$.
-The vector $\mathbf{v}_{m}(a_{B,y}^{max}) |_k$ contains the values of the speed limit from lateral acceleration at every waypoint in segment $k$.
+where $\mathbf{SL}_k$ is the vector containing the values of the legal speed limit at every way-point in segment $k$.
+The vector $\mathbf{v}_{m}(a_{B,y}^{max}) |_k$ contains the values of the speed limit from lateral acceleration at every way-point in segment $k$.
 % The minimum of scalar across all constituent points is chosen.
-While not time optimal, choosing the minimum of these values over the entire segment ensures that no constraints are violated, and as a result, many of the curves will have a constant speed.
+While not time optimal, choosing the minimum of these values over the entire segment ensures that no constraints are violated. As a result, many of the curves have a constant speed.
 For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
@@
 The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
-Since the final acceleration is to be set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
+Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
 For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
@@
 
-\subsection{Fitting Jerk Profiles to Each Segment} \label{sec:jerkprofiles}
+\subsection{Fitting Jerk Profiles to each Segment} \label{sec:jerkprofiles}
 
-To introduce the process of computing a jerk profile for each segment, let us consider the case in which a vehicle needs to travel some arbitrary distance ahead.
-The vehicle is moving at an initial speed and needs to accelerate to a traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
+To introduce the process of computing a jerk profile for each segment, we consider the case where a vehicle needs to travel an arbitrary distance ahead.
+Starting from an initial speed, the vehicle needs to accelerate to a traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
 Figure \ref{fig:full7phasespec} shows this trajectory over time.
@@
   \centering
-  \includegraphics[width=1.0\columnwidth]{graphics/Full7PhaseSpecVertical.png}
+  \includegraphics[width=0.8\columnwidth]{graphics/Full7PhaseSpecVertical.png}
   \caption{
@@
 \begin{equation}
-  a^{+,max}_m >= a^+_m > 0 > a^-_m >= a^{-,max}_m
+  a^{+,max}_m \geqslant a^+_m > 0 > a^-_m \geqslant a^{-,max}_m
   \label{eq:am}
@@
 \begin{equation}
-  j^{+,max}_m >= j^+_m > 0 > j^-_m >= j^{-,max}_m
+  j^{+,max}_m \geqslant j^+_m > 0 > j^-_m \geqslant j^{-,max}_m
   \label{eq:jm}
@@
 
-We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
-These 4 profiles cover cases in which the 7-phase solution fails.
+The 7 phase solution is time optimal as it drives the car at the speed limit.
+However, sometimes the distance is too small to get up-to the speed limit, or the boundary conditions are different, such as going from a specific speed to a stop.
+We define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
+%These 4 profiles cover cases in which the 7-phase solution fails.
 For instance, if $L$ is too short or $v_m$ is too high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
@@
 
-Once the path manager triggers RSTOP, the trajectory is then recomputed with increasingly higher absolute values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} <= \Delta s_{avail}$.
+Once the path manager triggers RSTOP, the trajectory is then recomputed with increasingly higher absolute values of $a_m^-$ and $j_m^-$ until $\Delta s_{RSTOP} \leqslant \Delta s_{avail}$.
 Should one of the values $a_m^-$ or $j_m^-$ required to fulfill this condition exceed the limits $a_m^{-,max}$ or $j_m^{-,max}$, then limit braking parameters are set such that $a_m^-=a_m^{-,max}$ and $j_m^-=j_m^{-,max}$, and the user is alerted.
@@
 However, the time at which the trajectory plan must be referenced is not always known.
-The solution to this problem is known as trajectory tracking, and will be discussed in the next section.
+The solution to this problem is known as trajectory tracking, and is discussed in the next section.
 
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{graphics/PathProjectionSlice.png}
+  \includegraphics[width=0.65\columnwidth]{graphics/PathProjectionSlice.png}
   \caption{Showing projection of the current position of the car shown by green triangle to s coordinate system.
@@
 Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity and acceleration vectors) onto the path.
-A simple approximation is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a ``link'', and is depicted in Fig.
+A simple approximation is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a {\it link}, and is depicted in Fig.
 \ref{fig:cartos}).
@@
 
-To summarize, in the last three sections we described how an urban path is broken into sub-paths. We 
+To summarize, in the last three sections we described breaking an urban path into sub-paths. We 
 explained the computation of velocity and jerk profile for segments within sub-paths that honor constraints

### 2016-04-20 root.tex
small sentence addition in section 4 for clarity
figures at top and bottom only

@@
 \centering
-  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=0.8\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
+  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=0.75\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
   \\
-  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=0.9\columnwidth]{graphics/RSTOP_Replan.png}}
+  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=0.85\columnwidth]{graphics/RSTOP_Replan.png}}
   \\
@@
 For calculating $\Delta s_{RSTOP}$, the same set of values from normal planning is used and the distance traversed in the resultant trajectory is reported as the current value of $\Delta s_{RSTOP}$.
+This is the procedure for computing the dynamic, speed-dependent distance required to stop at any epoch.
 

### 2016-04-20 root.tex
remove B for body frame

@@
 % How we actually did this.
-For the purposes of developing a simple example application, two constraints are used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
+For the purposes of developing a simple example application, two constraints are used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{y}^{max}$. 
 Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation :
 \begin{equation}
-  v_{m,w}(a_{B,y}) = \sqrt{ | a_{B,y}^{max} / \kappa_{w} | } ~,~ w = 1, ..., W
+  v_{m,w}(a_{y}) = \sqrt{ | a_{y}^{max} / \kappa_{w} | } ~,~ w = 1, ..., W
 \end{equation}
 where $W$ is the number of waypoints in the sub-path, and $\kappa_{w}$ denotes the path curvature at each point.
-For each sub-path point, the value $v_{m,w}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration $a^{max}_{B,y}$.
+For each sub-path point, the value $v_{m,w}(a_{y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration $a^{max}_{y}$.
 
@@
   \label{fig:1Dto2DandSegmentationb}
-  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling.png}
+  \includegraphics[width=0.8\columnwidth]{graphics/speed_ceiling2.PNG}
   }
@@
 must be calculated for each segment using the values corresponding to the waypoints comprising the segment.
-The principle task is to find a value of $v_m$ for each segment using the values from segmentation, $SL$ and $v_{m,w}(a_{B,y}^{max})$. 
+The principle task is to find a value of $v_m$ for each segment using the values from segmentation, $SL$ and $v_{m,w}(a_{y}^{max})$. 
 The value of the speed ceiling for segment $k$ is chosen to be:
 \begin{equation}
-  v_{m,k} = min \left ( \mathbf{SL}_k ,  \mathbf{v}_{m}(a_{B,y}^{max}) |_k \right )
+  v_{m,k} = min \left ( \mathbf{SL}_k ,  \mathbf{v}_{m}(a_{y}^{max}) |_k \right )
 \end{equation}
 where $\mathbf{SL}_k$ is the vector containing the values of the legal speed limit at every way-point in segment $k$.
-The vector $\mathbf{v}_{m}(a_{B,y}^{max}) |_k$ contains the values of the speed limit from lateral acceleration at every way-point in segment $k$.
+The vector $\mathbf{v}_{m}(a_{y}^{max}) |_k$ contains the values of the speed limit from lateral acceleration at every way-point in segment $k$.
 % The minimum of scalar across all constituent points is chosen.
@@
 % To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
-% Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
-% Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
-% Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
-% The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.
+% Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{x}(t)$ to set the derivatives.
+% Speed along the path $v_{x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{x}(t)$ used to set the derivatives at every knot.
+% Acceleration $a_{x}(t)$ is piecewise linear, so lookup is a trivial matter.
+% The same applies to $j_{x}(t)$, which may be implemented with a simple table.
 
@@
 
-Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to look up $v_{B,x}(t_{veh})$, $a_{B,x}(t_{veh})$, and $j_{B,x}(t_{veh})$.
+Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to look up $v_{x}(t_{veh})$, $a_{x}(t_{veh})$, and $j_{x}(t_{veh})$.
 Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.

### 2016-04-20 scripts/results_plot_final.py
results plot added

@@ -1,0 +1,186 @@
+#!/usr/bin/env python
+import sys, os
+import pyqtgraph as pg
+from scipy.signal import filtfilt, savgol_filter
+import numpy as np
+import pyqtgraph.exporters
+
+
+slice_cmp_filen = '/media/rgcofield/Handvaska/GoogleDrive/GAVLab/IVS-2016/Data/Export_ITS_20160419_142308_ITS_vehicle_planning_1__ITSPlanner_1_slice_cmp.csv'
+slice_s_filen = '/media/rgcofield/Handvaska/GoogleDrive/GAVLab/IVS-2016/Data/Export_ITS_20160419_142308_ITS_vehicle_planning_1__ITSPlanner_1_slice_s.csv'
+state_filen = 'Data/Export_ITS_20160419_142308_vehicle_state_2__Kinematic_state_vectorizer_outputFloat.csv'
+
+
+this_dir = os.path.dirname(os.path.abspath(__file__))
+data_dir = os.path.join(this_dir, 'Data')
+
+
+def rtm_time_str_to_float(s):
+  return float(s.split(':')[0])*60. + float(s.split(':')[1])
+
+def smooth(y, box_pts):
+  """moving average"""
+  box = np.ones(box_pts)/box_pts
+  y_smooth = np.convolve(y, box, mode='same')
+  return y_smooth
+
+# slice_cmp:
+#  RTMTime, reftime, x, y, heading, curvature, speed, accel, jerk &&&& repeat
+# slice_time = []
+slice_v = []
+slice_a = []
+for line in open(slice_cmp_filen, 'r'):
+  data = line.split(';')
+  # slice_time.append(rtm_time_str_to_float(data[0]))
+  slice_v.append(float(data[6]))
+  slice_a.append(float(data[7]))
+
+# slice_s
+veh_s = []
+for line in open(slice_s_filen, 'r'):
+  data = line.split(';')
+  veh_s.append(float(data[1]))
+
+# veh state:
+#  RTMTime, x, y, speed, accel, heading ????
+# veh_time = []
+veh_v = []
+veh_a = []
+for line in open(state_filen):
+  data = line.split(';')
+  if len(data) < 3:
+    continue
+  # veh_time.append(rtm_time_str_to_float(data[0]))
[…]

### 2016-04-20 root.tex
results plot added
add caption
add results section

@@
 This is the trajectory tracking and is discussed in Section \ref{sec:trajectorytracking},
-followed by conclusions and future extensions.
+followed by results and conclusions.
 % tracking
@@
 
-Our Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
-Each of the states is described below.
-
 \begin{figure}[tb]
@@
 
+Our Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
+%Each of the states is described below.
 In the absence of pedestrians, the path manager cycles between NORMAL and PSTOP states.
@@
   \centering
-  \includegraphics[width=0.5\columnwidth]{graphics/Subpaths.png}
+  \includegraphics[width=0.4\columnwidth]{graphics/Subpaths.png}
   \caption{
@@
 The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no separate logic.
-They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
-It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure no tracking errors in the pedestrian detection algorithm.
+They follow the same state transition, and since all trajectory plans are smooth, riders are rarely aware that any change took place.
+There is a brief waiting period before resuming NORMAL mode from RSTOP to ensure no tracking errors in pedestrian detection.
 
 \begin{figure}[tb]
-\centering
+%\captionsetup[subfigure]{width=\columnwidth}
+\centering%\captionsetup{width=\columnwidth}
   \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=0.75\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
@@
 While not time optimal, choosing the minimum of these values over the entire segment ensures that no constraints are violated. As a result, many of the curves have a constant speed.
-For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
-Once the values $v_{m,k}$ are set for every segment in the subpath, boundary conditions must be set to ensure that speed is continuous between segments.
-Acceleration continuity will be addressed later.
-The algorithm in Fig. \ref{alg:segmentspeedboundaryconditions} is used to compute the vector $\mathbf{v}_p$ and then each segment $k$ can be solved using:
+For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.
+Once $v_{m,k}$ are set for every segment in the subpath, boundary conditions are set to ensure that speed is continuous between segments. Acceleration continuity is addressed later.
+The algorithm in Fig. \ref{alg:segmentspeedboundaryconditions} is used to compute the vector $\mathbf{v}_p$ and each segment $k$ is solved using:
 \begin{equation}
@@
 The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
-Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
+Since the final acceleration is set to zero for all profiles, each subsequent segment begins with $a_i = 0$.
 For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
@@
 
-Now that all segments within the subpath have been parameterized, they may be solved sequentially.
+Now that all segments within the subpath have been parameterized, they are solved sequentially.
 A jerk vs. time profile is fit to each segment using the algorithm described in the next sub-section.
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{graphics/Full7PhaseSpecVertical.png}
+  \includegraphics[width=0.7\columnwidth]{graphics/Full7PhaseSpecVertical.png}
   \caption{
@@
 \ref{eq:am} and \ref{eq:jm}.
-In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.
+In the simplest case, one value for each of the quantities expressed in these relations is chosen for the entire path.
 
@@
 However, sometimes the distance is too small to get up-to the speed limit, or the boundary conditions are different, such as going from a specific speed to a stop.
-We define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
+We define four other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
 %These 4 profiles cover cases in which the 7-phase solution fails.
@@
 All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
-It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
+Traveling backward (i.e., in reverse gear) is not supported in this work.
 To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver is used 
 %\textbf{RG: can specify solver name???} .
-These 4 profiles are:
+The five jerk profiles in our {\it Jerk Library} are:
 
@@
 The 7-phase profile can be expressed as $( j_1 > 0 , j_2 = 0 , j_3 < 0 , j_4 = 0 , j_5 < 0 , j_6 = 0 , j_7 > 0 )$.
-In the instance where $\Delta t_4$ is negative, the total segment length $L$ is too short to reach top speed of $v_m$, so a 6-phase profile must be used.
+When $\Delta t_4$ is negative, the total segment length $L$ is too short to reach top speed of $v_m$, so a 6-phase profile must be used.
 
@@
 Should that yield no positive time intervals, one of the two 4-phase profiles then becomes appropriate, since $L$ is not large enough to accommodate accelerating past the end speed.
-If $v_i > v_f$, then a reversed 4-phase profile is attempted.
-However, if $v_i > v_f$, then a regular 4-phase profile is attempted.
+If $v_i > v_f$, then a 4-phase profile is attempted. Else 
+if $v_i > v_f$, then a reversed 4-phase profile is attempted.
 
@@
 % \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}
-For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
-This results in a negative time interval for one or more phases.
+For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ is not feasible, % for a desired speed at the end of the phase.
+resulting in a negative time interval for one or more phases.
 In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
 Specifically, jerk may be increased while peak acceleration is decreased.
-This process is referred to herein as ``accel-jerk tuning''.
-If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.
+This process is referred to herein as {\it accel-jerk tuning}.
+If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a different jerk profile must be chosen.
 
@@
 The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
-Each has been described above.
-The routines for fitting a profile to a given $\mathbf{b}, \mathbf{a}_m, \mathbf{j}_m$ to a segment using this cascade approach is referred to as a ``jerk library''.
+%Each has been described above.
+The routines for fitting a profile to a given $\mathbf{b}, \mathbf{a}_m, \mathbf{j}_m$ to a segment using this cascade approach is referred to as a {\it Jerk Library}.
 
@@
 Now a unified reference trajectory must be constructed by stitching these together.
-
 For each segment with $M$ jerk phases, the following values are computed, which correspond to the instants at which jerk changes:
@@
 
-However, the time at which the trajectory plan must be referenced is not always known.
+The time at which the trajectory plan must be referenced is not always known.
 The solution to this problem is known as trajectory tracking, and is discussed in the next section.
@@
 
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=0.65\columnwidth]{graphics/results_final.png}
+  \caption{Planned speed and longitudinal acceleration vs. measured values}
+
+  \label{fig:results}
+\end{figure}
+
+\section{Results} \label{sec:results}
+
+Fig.
+\ref{fig:results} shows a planned and actual velocity and acceleration for the second sub-path between the two stop signs in Fig 5.
+Both velocity and acceleration track quite well.
+However the real acceleration is quite noisy.
+We have high jerk and acceleration to compensate for significant damping effects and time constants in the controller.
+%So we need high error to get the vehicle to move.
+Our controller was tuned for interstate driving, so its behavior at low urban speeds has not been optimized leading to noisy acceleration.
+
 \section{Summary} \label{sec:summary}
 
-In this paper, we have presented a novel real-time method for planning longitudinal trajectories 
-in an autonomous car to follow an urban path with online updates to avoid pedestrians on the roadway by slowing down or 
-stopping.
+In this paper, we have presented a novel real-time method for planning longitudinal trajectories in an autonomous car to follow an urban path with online updates to avoid pedestrians on the roadway by slowing down or stopping.
 We built a robust stopping model as part of Trajectory Planner reliably integrated with the Path Manager.
-We can travel along a desired path, stop at stop signs, stop for pedestrians in the roadway, and continue after 
-they leave. Our system gracefully handles complex cases that were not explicitly programmed, such as following 
-pedestrians if they continue walking on the road away from the car. Our system works with multiple 
-pedestrians as we only need to consider the closest pedestrian ahead of us.
+We can travel along a desired path, stop at stop signs, stop for pedestrians in the roadway, and continue after they leave.
+Our system gracefully handles complex cases that were not explicitly programmed, such as following pedestrians if they continue walking on the road away from the car. 
+Our system works with multiple pedestrians as we only need to consider the closest pedestrian ahead of us.
 

### 2016-04-21 root.tex
updated results section, added data files

@@
 Fig.
-\ref{fig:results} shows a planned and actual velocity and acceleration for the second sub-path between the two stop signs in Fig 5.
-Both velocity and acceleration track quite well.
-However the real acceleration is quite noisy.
-We have high jerk and acceleration to compensate for significant damping effects and time constants in the controller.
+\ref{fig:results} shows planned speed and acceleration for the second sub-path between the two stop signs in Fig \ref{fig:subpathdivision} (highlighted in green).
+Overlayed on top are measured values of both quantities, both of which track quite well.
+Note that they are not smoothed or bias-corrected, thus the acceleration is quite noisy.
+Relatively high jerk and acceleration are used to compensate for significant damping effects and time constants in the controller.
 %So we need high error to get the vehicle to move.
-Our controller was tuned for interstate driving, so its behavior at low urban speeds has not been optimized leading to noisy acceleration.
+The controller used in this work was originally tuned for interstate driving, so its behavior at low urban speeds has not been optimized leading to noisy acceleration.
 

### 2016-04-21 scripts/results_plot_final.py

@@ -162,6 +162,15 @@ for p1 in fig_2_1, fig_2_2:
       if isinstance(single_item, pg.graphicsItems.LabelItem.LabelItem):
           single_item.setText(single_item.text, **legendLabelStyle)
 
+## RMS speed error
+
+speed_percent_error = []
+
+for k in range(len(veh_v2)):
+  # # if slice_v2[k] > 0:
+  # speed_percent_error.append( abs(veh_v2[k]-slice_v2[k])/slice_v2[k] )
+rms_speed_percent_error = np.sqrt(np.mean(np.square(speed_percent_error))) * 100
+print rms_speed_percent_error
 
 # win3 = pg.GraphicsWindow()
 # fig_3_1 = win3.addPlot(row=0, col=0)

### 2016-04-21 root.tex
minor grammatical edits

@@
 %When planning trajectories for 
-An autonomous ground vehicle operating on roadways, typically plans for trajectories with the goal of minimizing navigation time.
+An autonomous ground vehicle operating on roadways typically plans for trajectories with the goal of minimizing navigation time.
 % "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
@@
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-{\it Path Planning} involves computation of the optimal path and {\it Trajectory Planning} involves the computation of a velocity profile appropriate for the scenario.
+{\it Path Planning} involves computation of the optimal path and {\it Trajectory Planning} involves the computation of a time-parameterized velocity profile appropriate for the scenario.
 
@@
 
-We present an autonomy framework for pedestrian-aware ground vehicles
-with a novel method for planning longitudinal trajectories 
+We present a deterministic autonomy framework for pedestrian-aware ground vehicles with a novel method for planning longitudinal trajectories 
 %(the time derivatives of position in the body-forward direction) 
@@
 Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
-We intelligently choose a jerk profile from a set of pre-solved profiles such that acceleration is continuous and within limits throughout the entire path.
+We intelligently choose a jerk profile from a set of analytically solved profiles such that acceleration is continuous and within limits throughout the entire path.
 The resultant trajectory plan is then sampled to provide a reference for real-time vehicle control.
@@
 Its primary functions are:
-\begin{enumerate}
+\begin{itemize}
   \item Stop the vehicle at stop signs and wait until the associated intersection is clear before continuing.
@@
   \item Govern the Trajectory Planner.
-\end{enumerate}
+\end{itemize}
 The Low-Level Planner is also referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
@@
 
-The Path Manager performs trajectory planning over a sliding window within the full path with the sliding window moving with the vehicle.
+The Path Manager performs trajectory planning over an incremental window within the full path.
 %As shown in Fig. \ref{fig:addreact}, a Path Planner passes paths into it as a 2D linestring.
-It breaks the path planner provided path into sub-paths at each stop location which are known a priori, as shown in Fig.
+This is done by breaking the planned path into smaller sub-paths at each stop location which are known a priori, as shown in Fig.
 \ref{fig:subpathdivision}.
-It then calls the Trajectory Planner for each sub-path.
+Trajectory planning is then performed for each sub-path incrementally as the vehicle progresses.
 
@@
 This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
-$\Delta s_{replan}$ value is typically set to 1 meter.
+% $\Delta s_{replan}$ value is typically set to 1 meter.
 Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$.
@@
 
-The task of Trajectory Planning is to take a set of scalar waypoints in the frenet frame from the Path Manager, and compute a series of constant jerk intervals for longitudinal motion.% along the $s$-axis
+The task of Trajectory Planning is to take a set of scalar waypoints in the frenet frame from the Path Manager, and compute a series of constant jerk intervals for longitudinal motion.
+% along the $s$-axis
 The resulting trajectory should minimize time, and satisfy all the constraints on speed, acceleration, and jerk.
@@
 
-To introduce the process of computing a jerk profile for each segment, we consider the case where a vehicle needs to travel an arbitrary distance ahead.
+To introduce the process of computing a jerk profile for each segment, consider the case where a vehicle needs to travel an arbitrary distance ahead.
 Starting from an initial speed, the vehicle needs to accelerate to a traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
@@
 A profile with $M$ constant jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
-As we show later, $\mathbf{a}_m$, and $\mathbf{j}_m$ must sometimes be modified to obtain physically possible time intervals.
+As we show later, $\mathbf{a}_m$, and $\mathbf{j}_m$ must sometimes be modified to obtain non-negative possible time intervals.
 To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs.
@@
 The 7 phase solution is time optimal as it drives the car at the speed limit.
-However, sometimes the distance is too small to get up-to the speed limit, or the boundary conditions are different, such as going from a specific speed to a stop.
+However, sometimes the distance is too small to reach the speed limit, or the boundary conditions are different, such as going from a specific speed to a stop.
 We define four other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile.
@@
 Should that yield no positive time intervals, one of the two 4-phase profiles then becomes appropriate, since $L$ is not large enough to accommodate accelerating past the end speed.
-If $v_i > v_f$, then a 4-phase profile is attempted. Else 
-if $v_i > v_f$, then a reversed 4-phase profile is attempted.
+If $v_i > v_f$, then a 4-phase profile is attempted.
+However, if $v_i > v_f$, then a reversed 4-phase profile is attempted.
 
@@
 Several algorithms exist to find the optimal projection of the target vehicle's present kinematic state (position, velocity and acceleration vectors) onto the path.
+Such procedures are referred to as \emph{map matching} in the literature.
 A simple approximation is to perpendicularly project the vehicle's position onto a series of lines running between each pair of adjacent points in the path (this will be referred to as a {\it link}, and is depicted in Fig.
@@
 \ref{fig:results} shows planned speed and acceleration for the second sub-path between the two stop signs in Fig \ref{fig:subpathdivision} (highlighted in green).
-Overlayed on top are measured values of both quantities, both of which track quite well.
+Overlayed on top are measured values of both quantities for a single autonomous driving run, both of which track quite well.
 Note that they are not smoothed or bias-corrected, thus the acceleration is quite noisy.

### 2016-04-21 scripts/results_plot_final.py

@@ -167,8 +167,11 @@ for p1 in fig_2_1, fig_2_2:
 speed_percent_error = []
 
 for k in range(len(veh_v2)):
-  # # if slice_v2[k] > 0:
-  # speed_percent_error.append( abs(veh_v2[k]-slice_v2[k])/slice_v2[k] )
+  if slice_v2[k] > 0:
+    speed_percent_error.append( abs(veh_v2[k]-slice_v2[k])/slice_v2[k] )
+  else:
+    speed_percent_error.append(0)
+pg.plot(veh_s2, speed_percent_error)
 rms_speed_percent_error = np.sqrt(np.mean(np.square(speed_percent_error))) * 100
 print rms_speed_percent_error
 

### 2016-04-21 root.tex
put RMS speed error value in

@@
 
-\section{Results} \label{sec:results}
+   \label{sec:results}
 
@@
 \ref{fig:results} shows planned speed and acceleration for the second sub-path between the two stop signs in Fig \ref{fig:subpathdivision} (highlighted in green).
-Overlayed on top are measured values of both quantities for a single autonomous driving run, both of which track quite well.
+Overlayed on top are measured values of both quantities for a single autonomous driving run, both of which track quite well (the RMS speed error is $0.55 ~m/s$, though this is largely dependent upon the controller) .
 Note that they are not smoothed or bias-corrected, thus the acceleration is quite noisy.
@@
 %So we need high error to get the vehicle to move.
-The controller used in this work was originally tuned for interstate driving, so its behavior at low urban speeds has not been optimized leading to noisy acceleration.
+The controller used in this work was originally tuned for interstate driving, so its behavior at low urban speeds has not been optimized leading to significant steady state acceleration errors.
 

### 2016-04-21 scripts/results_plot_final.py

@@ -95,23 +95,23 @@ for line in open(state_filen):
 #     slice_s: line 1538
 #     state: line 1554
 
-slice_v1 = slice_v[:412]
-slice_a1 = slice_a[:412]
-veh_s1 = veh_s[:412]
-veh_v1 = veh_v[16:428]
-veh_a1 = veh_a[16:428]
+slice_v1 = np.array(slice_v[:412])
+slice_a1 = np.array(slice_a[:412])
+veh_s1 = np.array(veh_s[:412])
+veh_v1 = np.array(veh_v[16:428])
+veh_a1 = np.array(veh_a[16:428])
 
-slice_v2 = slice_v[413:1538]
-slice_a2 = slice_a[413:1538]
-veh_s2 = veh_s[413:1538]
-veh_v2 = veh_v[428:1553]
-veh_a2 = veh_a[428:1553]
+slice_v2 = np.array(slice_v[413:1538])
+slice_a2 = np.array(slice_a[413:1538])
+veh_s2 = np.array(veh_s[413:1538])
+veh_v2 = np.array(veh_v[428:1553])
+veh_a2 = np.array(veh_a[428:1553])
 
-slice_v3 = slice_v[1538:2076]
-slice_a3 = slice_a[1538:2076]
-veh_s3 = veh_s[1538:2076]
-veh_v3 = veh_v[1553:]
-veh_a3 = veh_a[1553:]
+slice_v3 = np.array(slice_v[1538:2076])
+slice_a3 = np.array(slice_a[1538:2076])
+veh_s3 = np.array(veh_s[1538:2076])
+veh_v3 = np.array(veh_v[1553:])
+veh_a3 = np.array(veh_a[1553:])
 
 
 def quit(*args):
@@ -164,16 +164,19 @@ for p1 in fig_2_1, fig_2_2:
 
 ## RMS speed error
 
-speed_percent_error = []
+# speed_percent_error = []
+# error_range = range(50, len(veh_v2)-50)
+# for k in error_range:
+#   if slice_v2[k] > 0:
+#     speed_percent_error.append( abs(veh_v2[k]-slice_v2[k])/slice_v2[k] )
+#   else:
+#     speed_percent_error.append(0)
+# rms_speed_percent_error = np.sqrt(np.mean(np.square(speed_percent_error))) * 100
+# print rms_speed_percent_error
+# pg.plot(veh_s2[error_range], speed_percent_error)
 
-for k in range(len(veh_v2)):
[…]

### 2016-04-25 root.tex
more edits

@@
 \begin{abstract}
-In this paper, we address the problem of trajectory planning for an autonomous car for urban roads while simultaneously reacting to the presence of pedestrians.
+In this paper, we address the problem of trajectory planning while simultaneously reacting to the presence of pedestrians for an autonomous car on urban roads.
 Past work limits jerk, velocities and acceleration for smooth trajectories, without considering reactive behaviors such as responding to pedestrians.
@@
 In this paper, we present an integrated trajectory generation and tracking system. % for urban environments. 
-Our system simultaneously considers constraints for smooth trajectory and updates the trajectory in real-time to react to pedestrians on the road.
+Our system simultaneously considers both parameter and reactive constraints for smooth trajectory and updates it in real-time.
 
@@
 %When planning trajectories for 
-An autonomous ground vehicle operating on roadways typically plans for trajectories with the goal of minimizing navigation time.
+An autonomous ground vehicle operating on roadways typically plans trajectories with the goal of minimizing navigation time.
 % "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
-It is common to employ the path-velocity decomposition to separate the twin tasks of Path and Trajectory Planning, so that these tasks can be run sequentially or in parallel.
+It is common to employ a path-velocity decomposition to separate the twin tasks of Path and Trajectory Planning, so that these tasks can be run sequentially or in parallel.
 % (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
-{\it Path Planning} involves computation of the optimal path and {\it Trajectory Planning} involves the computation of a time-parameterized velocity profile appropriate for the scenario.
+{\it Path Planning} computes the optimal path while {\it Trajectory Planning} computes a time-parameterized velocity profile appropriate for the scenario.
 
@@
 The set of possible paths to the destination is typically restricted
-by constraints such as available travel lanes, and obstacles such as pedestrians.
+by constraints such as available travel lanes, and obstacles including pedestrians.
 Many systems in the literature plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13}. 
@@
 
-We would like our autonomous car to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
+Our autonomous car should travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
 %Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
-At stop signs, our autonomous car should stop and wait for an appropriate amount of time.
+At stop signs, it should stop and wait for an appropriate amount of time.
 The car should also stop, slow down for pedestrians %when appropriate.It should react to pedestrians 
@@
 
-We present a deterministic autonomy framework for pedestrian-aware ground vehicles with a novel method for planning longitudinal trajectories 
+In this paper, we present a deterministic autonomy framework for pedestrian-aware ground vehicles with a novel method for planning longitudinal trajectories. 
 %(the time derivatives of position in the body-forward direction) 
-along the desired path. %without requiring prior knowledge of travel time.
-Our method segments the path using a set of kinematic or dynamic constraints.
+%along the desired path. %without requiring prior knowledge of travel time.
+Our method segments the path using both kinematic and dynamic constraints.
 Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
 We intelligently choose a jerk profile from a set of analytically solved profiles such that acceleration is continuous and within limits throughout the entire path.
-The resultant trajectory plan is then sampled to provide a reference for real-time vehicle control.
+The resultant trajectory plan is sampled to provide a reference for real-time vehicle control.
 %, which is called trajectory tracking.
@@
 %We then describe extensions to trajectory planner that handle reactivity to pedestrians in the environment. 
-Once the trajectory plan is formulated, we send the reference signal to the controllers that govern steering, braking, and engine speed.
-This is the trajectory tracking and is discussed in Section \ref{sec:trajectorytracking},
+Once the trajectory is formulated, we send the reference signal to the controllers that govern steering, braking, and engine speed.
+This is trajectory tracking and is discussed in Section \ref{sec:trajectorytracking},
 followed by results and conclusions.
@@
 
-Our system architecture is shown in Figure \ref{fig:addreact}. We have three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.
-
+Our system architecture is shown in Figure \ref{fig:addreact}.
+We have three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.
 The High-Level Planner is referred to as the {\it Path Manager}, and performs decision-making as well as governs operation of the low-level planner.
@@
 \end{itemize}
+
 The Low-Level Planner is also referred to as the {\it Trajectory Planner}, and is responsible for quantifying the actions dictated by the Path Manager.
@@
 This short-term trajectory is oriented in the vehicle body frame.
-This paper focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} \& Sec. \ref{sec:trajectorytracking}).
+This paper focuses on the Path Manager (Sec. \ref{sec:pathmanager}) and the Trajectory Planner (Sec. \ref{sec:trajectoryplanning} and Sec. \ref{sec:trajectorytracking}).
 
@@
 Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the user specified destination.
-This plan is then output as a set of global waypoints.
+This plan is output as a set of global waypoints.
 
@@
 In the absence of pedestrians, the path manager cycles between NORMAL and PSTOP states.
-In NORMAL state, the path manager sends the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
+In NORMAL state, the path manager sends the first subpath to the trajectory planner, which then plans and executes it, coming to a stop at the end of the subpath.
 In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
@@
 
-All detected pedestrian locations are compared against the high-resolution digital map. Pedestrians that are off roadway are not considered.
+All detected pedestrian locations are compared against the high-resolution digital map. Off road pedestrians are not considered.
 Each pedestrian location is perpendicularly projected onto the path to determine the travel distance separating them from the car.
 Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
-All the remaining pedestrians pose potential collision risks. However, only the closest pedestrian is considered for RSTOP.
+All the remaining pedestrians pose potential collision risk. However, only the closest pedestrian is considered for RSTOP.
 
 The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no separate logic.
-They follow the same state transition, and since all trajectory plans are smooth, riders are rarely aware that any change took place.
+They follow the same state transition, and since all trajectory plans are smooth, riders are rarely aware of the change.
 There is a brief waiting period before resuming NORMAL mode from RSTOP to ensure no tracking errors in pedestrian detection.
@@
 For the purposes of developing a simple example application, two constraints are used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{y}^{max}$. 
-Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation :
+Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation:
 \begin{equation}
@@
   \centering
-  \includegraphics[width=0.65\columnwidth]{graphics/results_final.png}
+  \includegraphics[width=0.7\columnwidth]{graphics/results_final.png}
   \caption{Planned speed and longitudinal acceleration vs. measured values}
@@
 We can travel along a desired path, stop at stop signs, stop for pedestrians in the roadway, and continue after they leave.
-Our system gracefully handles complex cases that were not explicitly programmed, such as following pedestrians if they continue walking on the road away from the car. 
-Our system works with multiple pedestrians as we only need to consider the closest pedestrian ahead of us.
+Our system gracefully handles complex cases that were not explicitly programmed, such as follow pedestrians if they continue walking on the road away from the car. 
+Our system works with multiple pedestrians as we only react to the closest pedestrian ahead of us.
 
@@
 Second, our trajectory planner has closed form solution and is suitable for online implementation
-on a car, unlike systems requiring solutions to nonlinear optimizations. Finally, we have confirmed the
+on a car, unlike systems requiring solutions to nonlinear optimizations. Finally, we have confirm the
 efficiency and reliability of our trajectory planner on a test vehicle with over 100 hours of testing
